---
title: "Flood Impact Analysis Report"
author: "Rik Lubbers"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)

file_path <- "C:/Users/rikib/Downloads/flood_impact_data.xlsx"

areas <- read_excel(file_path, sheet = "Areas")
respondents <- read_excel(file_path, sheet = "Respondents")
flood_experience <- read_excel(file_path, sheet = "FloodExperience")
healthcare_accessibility <- read_excel(file_path, sheet = "HealthcareAccessibility")
coping_mechanisms <- read_excel(file_path, sheet = "CopingMechanisms")
suggestions <- read_excel(file_path, sheet = "Suggestions")

final_merged_data <- respondents %>%
  left_join(areas, by = "area_id") %>%
  left_join(flood_experience %>% rename(flood_frequency = frequency_per_year), by = "respondent_id") %>%
  left_join(healthcare_accessibility, by = "respondent_id") %>%
  left_join(coping_mechanisms, by = "respondent_id") %>%
  left_join(suggestions, by = "respondent_id")

final_merged_data <- final_merged_data %>%
  mutate(
    area_name = as.factor(area_name),
    flood_frequency = as.factor(flood_frequency),
    severity = as.factor(severity),
    normal_travel_time_min = as.numeric(normal_travel_time_min),
    flood_travel_time_min = as.numeric(flood_travel_time_min),
    travel_time_diff = flood_travel_time_min - normal_travel_time_min
  )
```

# Exploratory Data Analysis

## Flood Frequency and Severity

The following plots show the distribution of flood frequency among respondents.

```{r}
table(final_merged_data$flood_frequency)
```

## Severity

Every respondent experienced floodings to a severe extent

```{r}
summary(final_merged_data$severity)
```

## Duration

The duration of the floods is shown in the following histogram, with one outlier removed (3600 hours extra)

```{r}
final_merged_data <- final_merged_data %>% filter(duration_hours < 3600)

ggplot(final_merged_data, aes(x = duration_hours)) +
  geom_histogram(binwidth = 10, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Flood Duration",
       x = "Flood Duration (hours)",
       y = "Frequency")
```

## Causes 

The causes of flooding are mostly due to poor and blocked drainage

```{r}
table(final_merged_data$causes)
```

Making travelling impossible during these events

```{r}	
table(final_merged_data$travel_difficulty)
```

# Healthcare access

Most people made use of the public facilities and some private. Private was preferred sometimes as public was not able to treat patients due to lack of medicines and other resources.

```{r}
ggplot(final_merged_data, aes(x = public_or_private
)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Distribution of Healthcare Access",
       x = "Healthcare Access",
       y = "Count")
```

## Distribution of Travel Time Difference

The distribution of travel time differences highlights the delays caused by flooding. Everyone thought it was more difficult to access healthcare during floodings. And people mostly walked with some using a boda (and one a taxi)

```{r}
summary(final_merged_data$harder_to_access)

summary(final_merged_data$transport_type)

# First removed an outlier
final_merged_data <- final_merged_data %>% filter(travel_time_diff < 300)

summary(final_merged_data[, c("normal_travel_time_min", "flood_travel_time_min", "travel_time_diff")])

summary(final_merged_data$water_height_cm)
```

```{r}
ggplot(final_merged_data, aes(x = travel_time_diff)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Travel Time Difference",
       x = "Travel Time Difference (minutes)",
       y = "Frequency")
```

## Travel Time Difference by Area

Understanding the differences in travel time across areas provides insight into localized impacts:

```{r}
ggplot(final_merged_data, aes(x = area_name, y = travel_time_diff)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Travel Time Difference by Area",
       x = "Area",
       y = "Travel Time Difference (minutes)")
```

Healthcare access was limited while everyones experiences more health issues during floodings

```{r}
summary(final_merged_data$health_problems)
summary(final_merged_data$problem_types)
```

## Coping and suggestions

Most people coped by staying at home and building barriers and clearing rubbish. The suggestions were mostly about improving the drainage system, roads and removing rubbish from drainage.

# Interpretation of Results

- **Flood Frequency**: Respondents experience frequent flooding, with most experiencing it multiple times a year.
- **Health impacts**: Many experience health problems during floodings.
- **Travel Time Differences**: The histogram shows that flooding significantly increases travel times for most respondents, with some extreme delays observed.
- **Area Variability**: Boxplots reveal that certain areas experience greater travel time delays, which could be due to infrastructure issues or geographic challenges.
- **Healthcare Access**: Respondents primarily rely on public healthcare facilities, but some opt for private care when public facilities are unable to provide treatment.
- **Coping Mechanisms**: Respondents employ a variety of coping strategies during floods, such as staying at home and clearing debris. These strategies can help mitigate the impacts of flooding.
- **Suggestions for Improvement**: Respondents suggest improving drainage systems, roads, and waste management to reduce the frequency and severity of floods.


---
