# Block Averaging of Population Raster Data

## Overview

This QMD file provides a script for performing block averaging of population raster data using R. Block averaging involves aggregating high-resolution population data into larger blocks to create a lower-resolution representation. This process can be useful for downsampling data or preparing it for analysis at a coarser scale.


# Load libraries
```{r}
# Set up renv for package management
if (!requireNamespace("renv", quietly = TRUE)) {
    install.packages("renv")
}

# Load the renv package
library(renv)

# Check if renv is already active, activate if not
if (!renv::status()$synchronized) {
    renv::activate()
}

# To restore the environment (for a collaborator)
renv::restore()

# Load the required libraries
library(terra)
```

# Load data
```{r}
# Read the population TIFF file
pop_raster_total <- rast("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Uganda/data/rPopulation/20240308161107/raw/UG_total.tif")

## load the shapefile of the study area
study_area <- vect("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Uganda/data/vBorders/20240308161055/processed/20240308162043/vBorders_Processed.shp")
```

# Transform the projection of the population raster
```{r}
# Transform the CRS of the study area to the population raster for faster processing
study_area <- project(study_area, crs(pop_raster_total))
```

# Crop, disaggregate and mask the data to the extent of the study area
```{r}
# Crop the population raster to the extent of the study area first due to computational limitations
## Crop the data to the extent of the study area with 2km extra on all sides
study_area_buffer <- buffer(study_area, width = 2000)

pop_raster_total <- crop(pop_raster_total, study_area_buffer)

# disaggregate population data
pop_total_10m <- disagg(pop_raster_total, c(10, 10))

## Mask the population raster to the extent of the study area
pop_total_10m <- mask(pop_total_10m, study_area)

# Project the population raster to the same projection as the original study area
study_area <- vect("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Uganda/data/vBorders/20240308161055/processed/20240308162043/vBorders_Processed.shp")

pop_total_10m <- project(pop_total_10m, crs(study_area))
```

# Save the data to new TIFF files
```{r}
## Write the disaggregated data to a new TIFF file
writeRaster(pop_total_10m, "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Uganda/data/rPopulation/20240308161107/processed/UG_total_10m.tif", overwrite = TRUE)

# also save it to the zToAccessMod folder
writeRaster(pop_total_10m, "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Uganda/data/zToAccessMod/20240319094916/UG_total_10m.tif", overwrite = TRUE)
```
