# Block Averaging of Population Raster Data

## Overview

This QMD file provides a script for performing block averaging of population raster data using R. Block averaging involves aggregating high-resolution population data into larger blocks to create a lower-resolution representation. This process can be useful for downsampling data or preparing it for analysis at a coarser scale.


# Load libraries
```{r}
# Set up renv for package management
if (!requireNamespace("renv", quietly = TRUE)) {
    install.packages("renv")
} else {
    message("renv is already installed.")
}

# Load the renv package
library(renv)

# Check if renv is already active, activate if not
if (!renv::status()$synchronized) {
    renv::activate()
}

# To restore the environment (for a collaborator)
renv::restore()

# Load the required libraries
library(terra)
```

# Load data
```{r}
# Read the population TIFF file
pop_raster_total <- rast(here("Data/processed/Worldpop_density/worldpop_density_kampala.tif"))

## load the shapefile of the study area
study_area <- vect(here("Data/processed/Adm_boundaries/parishes_kampala.shp"))
```

# Transform the projection of the population raster
```{r}
# Transform the CRS of the study area to the population raster for faster processing
study_area <- project(study_area, crs(pop_raster_total))
```

# disaggregate the data to 10x10m resolution
```{r}
# disaggregate population data
pop_total_10m <- disagg(pop_raster_total, c(10, 10))

# Adjust the values in the 10x10m cells
# Divide each cell by the number of new cells (10x10 grid = 100 cells)
pop_total_10m <- pop_total_10m / 100
```

# Save the data to new TIFF files
```{r}
# Save it to the zToAccessMod folder
writeRaster(pop_total_10m, here("AccessMod/InAccessMod/Uganda/data/zToAccessMod/20250108115609/pop_total_10m.tif"), overwrite = TRUE)
```
