# Merged Landcovers

This code is used to create merged landcover layers for each flood extent in the AccessMod analysis. It takes flood extent shapefiles and merges them with a base landcover raster to create new landcover layers.

## Step 1: Seperate the flood extents

The code first separates the flood extents based on the unique values of rainfall and duration. It loads the full flood extent data and defines the unique values for threshold, rainfall, and duration. Then, it iterates through each combination of threshold, rainfall, and duration and filters the flood extent data based on the current combination. If the filtered data is not empty, it creates a sub-folder for the flood extent and writes the filtered data to a shapefile in the sub-folder.

## Step 2: Load and name the flood extents

The code then loads all the flood extents from the sub-folders and reads them into a list. It names the flood extents in the list based on their sub-folder name. It also removes the other variable names from the flood extents.

## Step 3: Merge landcover layers for each flood extent

The code creates merged landcover layers for each flood extent. It initializes an empty list to store the merged landcover layers. It loads a base landcover raster and sets its projection. Then, it iterates through each flood extent in the flood list and performs the following steps:

1. Loads the landcover raster and creates a grid with the same extent and resolution.
2. Resamples the landcover raster to the grid resolution.
3. Transforms the flood extent polygons to the same projection as the landcover raster.
4. Rasterizes the flood extent polygons using the grid as a template.
5. Sets the background value of the rasterized flood extent to NA.
6. Merges the rasterized flood extent with the landcover raster.
7. Writes the merged landcover layer to a file.

The resulting merged landcover layers are saved in the "Merged landcovers" folder.

# R code
## Load libraries
```{r}
# Set up renv for package management
if (!requireNamespace("renv", quietly = TRUE)) {
    install.packages("renv")
} else {
    message("renv is already installed.")
}

# Load the renv package
library(renv)

# Check if renv is already active, activate if not
if (!renv::status()$synchronized) {
    renv::activate()
}

# To restore the environment (for a collaborator)
renv::restore()

# Load the required libraries
library(sf)
library(dplyr)
library(foreach)
library(snow)
library(terra)
library(raster)
library(ggplot2)
library(gridExtra)
library(tidyterra)
library(cowplot)
library(patchwork)
library(Spatial.align)
library(ggspatial)
library(here)
```

# Create merged landcover layer for AccessMod
## First seperate the flood extents based on the unique values of threshold, rainfall, and duration
```{r}
# Flood(x1)x(x2)x(x3), MLC = merged land cover, x1 = rainfall (mm), x2 = rain duration (h), x3 = flood depth (m)
# Load full data
flood_full <- st_read(here("AccessMod/InAccessMod/Uganda/data/vNaturalPolygons/20250107104606/processed/20250107121314"))

# Define unique values for threshold, rainfall, and duration
rainfall_values <- c(20, 40, 60, 80, 100)
duration_values <- c(3600, 10800, 21600)

# Seperate each flood extent based on the unique values of rainfall and duration
for (rainfall in rainfall_values) {
    for (duration in duration_values) {
        # Filter data based on current combination
        filtered_data <- flood_full[flood_full$rainfall == rainfall & flood_full$duration == duration, ]
        
        # Check if the filtered data is not empty
        if (nrow(filtered_data) > 0) {
            # Define sub-folder name
            subfolder_name <- paste0("flood_", rainfall, "_", duration)
            
            # Define the path directly inside Flood_extents
            dir_path <- here("Data", "raw", "Flood_extents", subfolder_name)
            
            # Create sub-folder if it doesn't exist
            if (!dir.exists(dir_path)) {
                dir.create(dir_path, recursive = TRUE)
            }
            
            # Define shapefile name
            shapefile_name <- paste0(subfolder_name, ".shp")
            
            # Write filtered data to shapefile in the sub-folder
            st_write(filtered_data, file.path(dir_path, shapefile_name), append = FALSE)
        }
    }
}

# Load all flood_extents from subfolders
flood_list <- list.files(path = here("Data", "raw", "Flood_extents"), pattern = "\\.shp$", full.names = TRUE, recursive = TRUE) %>%
    lapply(st_read)

# Name the flood extents in flood_list based on their subfolder name
flood_extents_names <- basename(dirname(list.files(path = here("Data", "raw", "Flood_extents"), pattern = "\\.shp$", full.names = TRUE, recursive = TRUE)))

for (i in 1:length(flood_list)) {
    flood_list[[i]]$Scenario <- gsub("flood_", "", flood_extents_names[i])
}

# Assign names to flood list
names(flood_list) <- flood_extents_names

# Remove other variable names from the flood extents
flood_list <- lapply(flood_list, function(x) {
    x$rainfall <- NULL
    x$duration <- NULL
    return(x)
})

```

## Merge landcover layers for each flood extent
```{r}
# Create merged landcover list
merged_landcovers <- list()

# Base MLC file path
MLC0 <- rast("C:/Users/rikib/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Merged landcovers/raster_land_cover_merged_MLC0_major_roads_10m/raster_land_cover_merged_MLC0_major_roads_10m.img")

# Load landcover raster to set projection
landcover <- rast("C:/Users/rikib/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Uganda/data/zToAccessMod/20240319094916/landcover_10m.tif")

crs(landcover)

crs(MLC0) <- crs(landcover)

## Iterate through each flood extent in the flood list and create merged landcover layers for each flood scenario
for (i in 1:length(flood_list)) {
    landcov <- MLC0 # load landcover raster
    flood_extent <- flood_list[[i]] %>% # load water polygons
        mutate(class = case_when(
            threshold == 0.1 ~ 1,
            threshold == 0.2 ~ 2,
            threshold == 0.3 ~ 3,
            TRUE ~ NA_real_
        )) %>% # assign classes based on threshold values
        as(., "Spatial") %>% # convert to sp object
        vect() # convert to terra object
    flood_extent_rast <- terra::rasterize(flood_extent, landcov, "class", touches = TRUE) # rasterize water polygons
    values(flood_extent_rast)[values(flood_extent_rast) == 0] <- NA # make background value NA
    landcover_merge <- terra::merge(flood_extent_rast, landcov) # merge landcover and water raster
    writeRaster(landcover_merge, paste0("C:/Users/rikib/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Merged landcovers/MLC_", flood_extents_names[i], ".tif"), overwrite = T) # write merged landcover to file
}
```

# Prepare merged landcover layers for plotting
```{r}	
# Load all merged landcovers from the Merged landcovers folder
merged_landcovers <- list.files(path = "C:/Users/rikib/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Merged landcovers", pattern = "\\.tif$", full.names = TRUE, recursive = TRUE) %>%
    lapply(terra::rast)

# Name the MLC extents in merged_landcovers based on their subfolder name
merged_landcovers_names <- list.files(path = "C:/Users/rikib/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Merged landcovers", pattern = "\\.tif$", full.names = F, recursive = TRUE)

merged_landcovers_names <- gsub("C:/Users/rikib/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Merged landcovers/MLC_", "", merged_landcovers_names)

# Load all merged landcovers as separate objects
for (i in 1:length(merged_landcovers)) {
    assign(gsub(".tif", "", merged_landcovers_names[i]), merged_landcovers[[i]])
}

# Define legend colours and names
legend_colour <- data.frame(
    value = c(1, 2, 3, 10, 20, 30, 40, 50, 60, 80, 90, 1000, 1001, 1002, 1003, 1004, 1005, 1006, 1007, 1008, 1009),
    color = c("#169FFB", "#3C3CCF", "#170AA6", "#A8E16E", "#FFB400", "#FFFF1B", "#D7B5D8", "#FF0000", "#30123B", "#0312DC", "#15C44C", "#DE0004", "#CF5052", "#CE0FA4", "#CA0FA4", "#336CE6", "#336CE6", "#34EF05", "#34EF05", "#333333", "#333333")
)

legend_names <- c(
    "Flood 10cm",
    "Flood 20cm",
    "Flood 30cm",
    "Tree Cover",
    "Shrubs",
    "Grass Land",
    "Agriculture",
    "Urban",
    "Bare",
    "Water Bodies",
    "Herbaceous wetland",
    "Trunk",
    "Trunk link",
    "Primary",
    "Primary Link",
    "Motorway",
    "Motorway Link",
    "Secondary",
    "Secondary link",
    "Tertiary",
    "Tertiary link"
)

# Create list of MLC names
MLC_names <- c(MLC_flood_20_3600, MLC_flood_20_10800, MLC_flood_20_21600, MLC_flood_40_3600, MLC_flood_40_10800, MLC_flood_40_21600, MLC_flood_60_3600, MLC_flood_60_10800, MLC_flood_60_21600, MLC_flood_80_3600, MLC_flood_80_10800, MLC_flood_80_21600, MLC_flood_100_3600, MLC_flood_100_10800, MLC_flood_100_21600)

# Create a data frame to define levels
levels_df <- data.frame(
  ID = legend_colour$value,   # Values in raster
  Category = legend_names     # Corresponding category names
)

# Assign levels to the factorized raster
MLC_names <- lapply(MLC_names, function(r) {
  levels(r) <- levels_df
  r
})

MLC_names <- lapply(MLC_names, function(r) {
  # Convert each raster to a factor
  as.factor(r)
})

names(MLC_names) <- c("20 mm / 1hr", "20 mm / 3hr", "20 mm / 6hr", "40 mm / 1hr", "40 mm / 3hr", "40 mm / 6hr", "60 mm / 1hr", "60 mm / 3hr", "60 mm / 6hr", "80 mm / 1hr", "80 mm / 3hr", "80 mm / 6hr", "100 mm / 1hr", "100 mm / 3hr", "100 mm / 6hr")

# Define color_mapping
color_mapping <- setNames(legend_colour$color, legend_colour$value)
```


# Load and prepare Health Facilities
```{r}
health_facilities <- vect("C:/Users/rikib/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/Facilities/suhsharan_health_facilities_UG/UG_HF_all.shp")

# Align the health facilities data to MLC
health_facilities <- align(health_facilities, MLC_flood_20_3600, show_plot = FALSE)
```	

# Plot
```{r}
# Define the output folder
output_folder <- "C:/Users/rikib/OneDrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Figures/MLC_floodings"  # Specify the folder to save plots
dir.create(output_folder, showWarnings = FALSE, recursive = TRUE)  # Create the folder automatically if it doesn't exist

# Generate and save plots
plots <- lapply(seq_along(MLC_names), function(i) {
    # Generate the plot
    plot <- autoplot(MLC_names[[i]]) +
        geom_sf(data = health_facilities, color = "black", size = 5, shape = 21, fill = "purple") +  # Add health facilities
        scale_fill_manual(
            values = setNames(legend_colour$color, legend_colour$value),  # Map colors to values
            labels = legend_names,  # Use correct legend names
            name = "Landcover Type",
            na.translate = FALSE,
            na.value = "transparent"
        ) +
        # Add scale bar
        annotation_scale(
            location = "bl", 
            width_hint = 0.3,
            text_cex = 2,  # Increase text size
            pad_y = unit(11, "cm"),  # Move scale bar up
            pad_x = unit(1, "cm")   # Move scale bar right
        ) +
        # Add north arrow
        annotation_north_arrow(
            location = "tr", 
            which_north = "true",
            style = north_arrow_fancy_orienteering,  # Fancy orienteering style
            height = unit(4, "cm"),  # Increase size
            width = unit(4, "cm")
        ) +
        # Add source annotation
        annotate(
            "text", 
            x = Inf, y = -Inf, 
            label = "Source: Authors", 
            hjust = 1.1, vjust = -0.5,  # Adjust position
            size = 5, fontface = "italic"
        ) +
        labs(title = names(MLC_names)[i]) +  # Use object name as title
        theme_minimal() +
        theme(
            axis.text = element_blank(),
            axis.title = element_blank(),
            legend.position = c(0.25, 0.13),  # Bottom-left corner (normalized coordinates)
            legend.title = element_text(size = 20, face = "bold"),
            legend.text = element_text(size = 18),
            plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
            legend.background = element_rect(fill = "white", color = "black", linewidth = 0.5),  # Add box around legend
            panel.border = element_rect(color = "black", fill = NA, linewidth = 1)  # Add border around the plot
        )
    
    # Rename the title to replace `/` with `X` and remove extra spaces
    original_title <- names(MLC_names)[i]
    renamed_title <- gsub(" / ", "X", trimws(original_title))  # Replace " / " with "X" and trim spaces
    
    # Save the plot as a file
    file_name <- paste0(output_folder, "/", renamed_title, ".png")  # Construct file name
    ggsave(filename = file_name, plot = plot, width = 24, height = 16, dpi = 1000)  # Save the plot as PNG
    
    return(plot)  # Return the plot object if needed
})

# Split plots into three groups
plots_group1 <- plots[1:6]
plots_group2 <- plots[7:12]
plots_group3 <- plots[13:15]

# Combine plots with a single legend for each group
combined_plot1 <- wrap_plots(plots_group1, ncol = 3) + 
  plot_layout(guides = "collect") +  # Combine all legends into one
  plot_annotation(title = "Merged Landcover Scenarios (1-6)") &  # Optional title 
    theme(legend.position = "bottom")

combined_plot2 <- wrap_plots(plots_group2, ncol = 3) + 
  plot_layout(guides = "collect") +  # Combine all legends into one
  plot_annotation(title = "Merged Landcover Scenarios (7-12)") &  # Optional title 
    theme(legend.position = "bottom")

combined_plot3 <- wrap_plots(plots_group3, ncol = 3) + 
  plot_layout(guides = "collect") +  # Combine all legends into one
  plot_annotation(title = "Merged Landcover Scenarios (13-15)")  &  # Optional title 
    theme(legend.position = "bottom")

# Print combined plots
combined_plot1
combined_plot2
combined_plot3

```
