# Merged Landcovers

This code is used to create merged landcover layers for each flood extent in the AccessMod analysis. It takes flood extent shapefiles and merges them with a base landcover raster to create new landcover layers.

## Step 1: Seperate the flood extents

The code first separates the flood extents based on the unique values of rainfall and duration. It loads the full flood extent data and defines the unique values for threshold, rainfall, and duration. Then, it iterates through each combination of threshold, rainfall, and duration and filters the flood extent data based on the current combination. If the filtered data is not empty, it creates a sub-folder for the flood extent and writes the filtered data to a shapefile in the sub-folder.

## Step 2: Load and name the flood extents

The code then loads all the flood extents from the sub-folders and reads them into a list. It names the flood extents in the list based on their sub-folder name. It also removes the other variable names from the flood extents.

## Step 3: Merge landcover layers for each flood extent

The code creates merged landcover layers for each flood extent. It initializes an empty list to store the merged landcover layers. It loads a base landcover raster and sets its projection. Then, it iterates through each flood extent in the flood list and performs the following steps:

1. Loads the landcover raster and creates a grid with the same extent and resolution.
2. Resamples the landcover raster to the grid resolution.
3. Transforms the flood extent polygons to the same projection as the landcover raster.
4. Rasterizes the flood extent polygons using the grid as a template.
5. Sets the background value of the rasterized flood extent to NA.
6. Merges the rasterized flood extent with the landcover raster.
7. Writes the merged landcover layer to a file.

The resulting merged landcover layers are saved in the "Merged landcovers" folder.

# R code
## Load libraries
```{r}
# Set up renv for package management
if (!requireNamespace("renv", quietly = TRUE)) {
    install.packages("renv")
} else {
    message("renv is already installed.")
}

# Load the renv package
library(renv)

# Check if renv is already active, activate if not
if (!renv::status()$synchronized) {
    renv::activate()
}

# To restore the environment (for a collaborator)
renv::restore()

# Load the required libraries
library(sf)
library(dplyr)
library(foreach)
library(snow)
library(terra)
library(raster)
library(ggplot2)
library(gridExtra)
library(tidyterra)
library(cowplot)
library(patchwork)
library(Spatial.align)
library(ggspatial)
library(here)
```

# Create merged landcover layer for AccessMod
## First seperate the flood extents based on the unique values of threshold, rainfall, and duration
```{r}
# Flood(x1)x(x2)x(x3), MLC = merged land cover, x1 = rainfall (mm), x2 = rain duration (h), x3 = flood depth (m)
# Load full data
flood_full <- st_read(here("AccessMod/InAccessMod/Uganda/data/vNaturalPolygons/20250107104606/processed/20250107121314"))

# Define unique values for threshold, rainfall, and duration
rainfall_values <- c(20, 40, 60, 80, 100)
duration_values <- c(3600, 10800, 21600)

# Seperate each flood extent based on the unique values of rainfall and duration
for (rainfall in rainfall_values) {
    for (duration in duration_values) {
        # Filter data based on current combination
        filtered_data <- flood_full[flood_full$rainfall == rainfall & flood_full$duration == duration, ]
        
        # Check if the filtered data is not empty
        if (nrow(filtered_data) > 0) {
            # Define sub-folder name
            subfolder_name <- paste0("flood_", rainfall, "_", duration)
            
            # Define the path directly inside Flood_extents
            dir_path <- here("Data", "raw", "Flood_extents", subfolder_name)
            
            # Create sub-folder if it doesn't exist
            if (!dir.exists(dir_path)) {
                dir.create(dir_path, recursive = TRUE)
            }
            
            # Define shapefile name
            shapefile_name <- paste0(subfolder_name, ".shp")
            
            # Write filtered data to shapefile in the sub-folder
            st_write(filtered_data, file.path(dir_path, shapefile_name), append = FALSE)
        }
    }
}

# Load all flood_extents from subfolders
flood_list <- list.files(path = here("Data", "raw", "Flood_extents"), pattern = "\\.shp$", full.names = TRUE, recursive = TRUE) %>%
    lapply(st_read)

# Name the flood extents in flood_list based on their subfolder name
flood_extents_names <- basename(dirname(list.files(path = here("Data", "raw", "Flood_extents"), pattern = "\\.shp$", full.names = TRUE, recursive = TRUE)))

for (i in 1:length(flood_list)) {
    flood_list[[i]]$Scenario <- gsub("flood_", "", flood_extents_names[i])
}

# Assign names to flood list
names(flood_list) <- flood_extents_names

# Remove other variable names from the flood extents
flood_list <- lapply(flood_list, function(x) {
    x$rainfall <- NULL
    x$duration <- NULL
    return(x)
})

```

## Merge landcover layers for each flood extent
```{r}
# Create merged landcover list
merged_landcovers <- list()

# Base MLC file path
MLC0 <- rast(here("Data/processed/Merged_landcovers/raster_land_cover_merged_MLC0.img"))

# Load landcover raster to set projection
landcover <- rast(here("AccessMod/InAccessMod/Uganda/data/zToAccessMod/20250108115609/UG_Landcover_kamp.tif"))

crs(MLC0) <- crs(landcover)

## Iterate through each flood extent in the flood list and create merged landcover layers for each flood scenario
for (i in 1:length(flood_list)) {
    landcov <- MLC0 # Load base landcover raster
    flood_extent <- flood_list[[i]] %>% # Load water polygons
        mutate(class = case_when(
            threshold == 0.1 ~ 1, # Assign classes based on thresholds
            threshold == 0.2 ~ 2,
            threshold == 0.3 ~ 3,
            TRUE ~ NA_real_
        )) %>% 
        as(., "Spatial") %>% # Convert to sp object
        vect() # Convert to terra object

    # Rasterize flood extent
    flood_extent_rast <- terra::rasterize(flood_extent, landcov, "class", touches = TRUE)

    # Preserve NA values from the base raster to prevent creating artificial "bridges"
    flood_extent_rast <- mask(flood_extent_rast, landcov, inverse = FALSE) 

    # Merge landcover and water raster while preserving NAs
    landcover_merge <- terra::cover(flood_extent_rast, landcov) 

    # Write the merged raster to file
    writeRaster(landcover_merge, here("Data", "processed", "Merged_landcovers", paste0("MLC_", flood_extents_names[i], ".tif")), overwrite = TRUE)
}
```

# Prepare merged landcover layers for plotting
```{r}	
# Load all merged landcovers from the Merged landcovers folder
merged_landcovers <- list.files(path = here("Data", "processed", "Merged_landcovers"), pattern = "\\.tif$", full.names = TRUE, recursive = TRUE) %>%
    lapply(terra::rast)

# Name the MLC extents in merged_landcovers based on their subfolder name
merged_landcovers_names <- list.files(path = here("Data", "processed", "Merged_landcovers"), pattern = "\\.tif$", full.names = FALSE, recursive = TRUE)

# Order the merged landcovers from low to high (20 to 100 mm)
# Extract flood depths and durations from file names
flood_depths <- as.numeric(sub(".*_flood_([0-9]+)_.*", "\\1", merged_landcovers_names))
durations <- as.numeric(sub(".*_([0-9]+)\\.tif$", "\\1", merged_landcovers_names))

# Replace NA for non-matching files (e.g., "MLC_Flood_extents_full.tif") with a large number
flood_depths[is.na(flood_depths)] <- Inf
durations[is.na(durations)] <- Inf

# Reorder the file names based on flood depth and then by duration
merged_landcovers_names <- merged_landcovers_names[order(flood_depths, durations, merged_landcovers_names)]

# Remove "MLC_Flood_extents_full.tif" and NA from the list
merged_landcovers_names <- merged_landcovers_names[!grepl("full", merged_landcovers_names) & !is.na(merged_landcovers_names)]

# Reorder the merged_landcovers list based on the same rules
merged_landcovers <- merged_landcovers[order(flood_depths, durations)]

# Load all merged landcovers as separate objects
for (i in 1:length(merged_landcovers)) {
    assign(gsub(".tif", "", merged_landcovers_names[i]), merged_landcovers[[i]])
}

# Define legend colours and names
legend_colour <- data.frame(
    value = c(1, 2, 3, 10, 20, 30, 40, 50, 60, 80, 90, 1001, 1002, 1003, 1004, 1005),
    color = c("#169FFB", "#3C3CCF", "#170AA6", "#A8E16E", "#FFB400", "#FFFF1B", "#D7B5D8", "#FF0000", "#30123B", "#0312DC", "#15C44C", "#DE0004", "#CE0FA4", "#336CE6", "#34EF05", "#333333")
)

legend_names <- c(
    "Flood 10cm",
    "Flood 20cm",
    "Flood 30cm",
    "Tree Cover",
    "Shrubs",
    "Grass Land",
    "Agriculture",
    "Urban",
    "Bare",
    "Water Bodies",
    "Herbaceous wetland",
    "Trunk",
    "Primary",
    "Motorway",
    "Secondary",
    "Tertiary"
)

# Create a data frame to define levels
levels_df <- data.frame(
  ID = legend_colour$value,   # Values in raster
  Category = legend_names     # Corresponding category names
)

# Assign levels to the factorized raster
merged_landcovers <- lapply(merged_landcovers, function(r) {
    levels(r) <- levels_df
    r
})

merged_landcovers <- lapply(merged_landcovers, function(r) {
    # Convert each raster to a factor
    as.factor(r)
})

names(merged_landcovers) <- c("20 mm / 1hr", "20 mm / 3hr", "20 mm / 6hr", "40 mm / 1hr", "40 mm / 3hr", "40 mm / 6hr", "60 mm / 1hr", "60 mm / 3hr", "60 mm / 6hr", "80 mm / 1hr", "80 mm / 3hr", "80 mm / 6hr", "100 mm / 1hr", "100 mm / 3hr", "100 mm / 6hr")

# Define color_mapping
color_mapping <- setNames(legend_colour$color, legend_colour$value)
```


# Load and prepare Health Facilities
```{r}
# Load health facilities from the processed folder
hf_all <- vect(here("Data", "processed", "Facilities", "facilities_all_kampala.shp"))
hf_hosp <- vect(here("Data", "processed", "Facilities", "facilities_hosp_kampala.shp"))
```	

# Plot
```{r}
# Define the output folder
output_folder <- here("Figures/MLC_floodings") # Specify the folder to save plots
dir.create(output_folder, showWarnings = FALSE, recursive = TRUE)  # Create the folder automatically if it doesn't exist

# Generate and save plots
plots <- lapply(seq_along(merged_landcovers), function(i) {
    # Generate the plot
    plot <- autoplot(merged_landcovers[[i]]) +
        geom_sf(data = hf_all, color = "black", size = 5, shape = 21, fill = "purple") +
        scale_fill_manual(
            values = setNames(legend_colour$color, legend_colour$value),
            labels = legend_names,
            name = "Landcover Type",
            na.translate = FALSE,
            na.value = "transparent"
        ) +
        annotation_scale(location = "bl", width_hint = 0.3, text_cex = 2, pad_y = unit(11, "cm"), pad_x = unit(1, "cm")) +
        annotation_north_arrow(location = "tr", which_north = "true", style = north_arrow_fancy_orienteering, height = unit(4, "cm"), width = unit(4, "cm")) +
        annotate("text", x = Inf, y = -Inf, label = "Source: Authors", hjust = 1.1, vjust = -0.5, size = 5, fontface = "italic") +
        labs(title = names(merged_landcovers)[i]) +
        theme_minimal() +
        theme(
            axis.text = element_blank(),
            axis.title = element_blank(),
            legend.position = c(0.25, 0.13),
            legend.title = element_text(size = 20, face = "bold"),
            legend.text = element_text(size = 18),
            plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
            legend.background = element_rect(fill = "white", color = "black", linewidth = 0.5),
            panel.border = element_rect(color = "black", fill = NA, linewidth = 1)
        )

    # Save the plot
    file_name <- paste0(output_folder, "/", gsub(" / ", "X", trimws(names(merged_landcovers)[i])), ".png")
    ggsave(filename = file_name, plot = plot, width = 24, height = 16, dpi = 1000)

    return(plot)
})

# Split plots into three groups
plots_group1 <- plots[1:6]
plots_group2 <- plots[7:12]
plots_group3 <- plots[13:15]

# Combine plots with a single legend for each group
combined_plot1 <- wrap_plots(plots_group1, ncol = 3) + 
  plot_layout(guides = "collect") +  # Combine all legends into one
  plot_annotation(title = "Merged Landcover Scenarios (1-6)") &  # Optional title 
    theme(legend.position = "bottom")

combined_plot2 <- wrap_plots(plots_group2, ncol = 3) + 
  plot_layout(guides = "collect") +  # Combine all legends into one
  plot_annotation(title = "Merged Landcover Scenarios (7-12)") &  # Optional title 
    theme(legend.position = "bottom")

combined_plot3 <- wrap_plots(plots_group3, ncol = 3) + 
  plot_layout(guides = "collect") +  # Combine all legends into one
  plot_annotation(title = "Merged Landcover Scenarios (13-15)")  &  # Optional title 
    theme(legend.position = "bottom")

# Print combined plots
combined_plot1
combined_plot2
combined_plot3

```
