
```{r}
# Set up renv for package management
if (!requireNamespace("renv", quietly = TRUE)) {
    install.packages("renv")
} else {
    message("renv is already installed.")
}

# Load the renv package
library(renv)

# Check if renv is already active, activate if not
if (!renv::status()$synchronized) {
    renv::activate()
}

# To restore the environment (for a collaborator)
renv::restore()

# Load the required libraries
library(terra)
library(Spatial.align)
library(patchwork)
library(sf)
library(biscale)
library(classInt)
library(here)
library(ggplot2)
library(tidyterra)
```

# Weighted Delta Travel Time Calculation for scenarios with All Facilities
```{r}
# Load delta travel time and population distribution
## For delta travel time go to script 03.3 maps_tt_and_delta_TT
### After doing so make a list using the following code:

delta_tt_list_all <- mget(ls(pattern = "delta_tt_all"))

## Load population distribution
pop_density_raster <- rast(here("AccessMod", "InAccessMod", "Uganda", "data", "zToAccessMod", "20250108115609", "pop_total_10m.tif"))

# Load the SpatVector for parish boundaries
parish_boundaries_all <- vect(here("Data", "processed", "Adm_boundaries", "parishes_kampala.shp"))

# Step 1: Iterate through each delta_tt_raster in delta_tt_list_all and multiply by population density raster
# Create weighted delta TT list for all facilities
weighted_delta_tt_list_all <- lapply(delta_tt_list_all, function(delta_tt_raster) {
    delta_tt_raster * pop_density_raster
})
names(weighted_delta_tt_list_all) <- sapply(names(weighted_delta_tt_list_all), function(name) {
    parts <- strsplit(name, "_")[[1]]
    paste(tail(parts, 2), collapse = "_")
})

# Step 2: Zonal stats for all facilities
for (i in seq_along(weighted_delta_tt_list_all)) {
    sum_weighted_delta_tt <- zonal(weighted_delta_tt_list_all[[i]], parish_boundaries_all, fun = "sum", na.rm = TRUE)
    original_name <- names(weighted_delta_tt_list_all)[i]
    parish_boundaries_all[[paste0("Weighted_delta_tt_", original_name)]] <- sum_weighted_delta_tt[, 1]
}

# Step 3: Sum population per parish for all facilities
parish_boundaries_all$Population_Sum <- zonal(pop_density_raster, parish_boundaries_all, fun = "sum", na.rm = TRUE)

# Step 4: Calculate average weighted delta TT
for (i in seq_along(weighted_delta_tt_list_all)) {
    original_name <- names(weighted_delta_tt_list_all)[i]
    parish_boundaries_all[[paste0("Average_Weighted_Delta_TT_", original_name)]] <-
        parish_boundaries_all[[paste0("Weighted_delta_tt_", original_name)]] / parish_boundaries_all$Population_Sum
}

```

# Create a bivariate map of accessibility reduction and population sum for all facilities
```{r}	
# Step 5: Plot bivariate map for ALL FACILITIES
# STEP: Prepare sf object for ALL FACILITIES
parish_boundaries_all_sf <- st_as_sf(parish_boundaries_all)

# STEP: Make negative values positive for visualization
for (col in names(parish_boundaries_all_sf)[grepl("Average_Weighted_Delta_TT_", names(parish_boundaries_all_sf))]) {
    parish_boundaries_all_sf[[col]] <- abs(parish_boundaries_all_sf[[col]])
}

# STEP: Extract and reorder scenario columns
scenario_columns_all <- grep("Average_Weighted_Delta_TT_", names(parish_boundaries_all_sf), value = TRUE)
scenario_columns_ordered_all <- scenario_columns_all[order(
    as.numeric(sub(".*_(\\d+)_.*", "\\1", scenario_columns_all)),
    as.numeric(sub(".*_(\\d+)$", "\\1", sub(".*_\\d+_", "", scenario_columns_all)))
)]

# STEP: Create consistent population bin
parish_boundaries_all_sf$Population_bin <- factor(cut(
    parish_boundaries_all_sf$Population_Sum,
    breaks = c(0, 10000, 30000, Inf),
    include.lowest = TRUE,
    labels = c("Low", "Medium", "High")
), levels = c("High", "Medium", "Low"))

# STEP: Define bivariate legend labels
labels_all <- list(
    bi_x = c("0-30", "30-60", "60+"),
    bi_y = c("30000+", "10000-30000", "0-10000")
)

# STEP: Create plots for each scenario
plot_list_all <- list()

for (i in seq_along(scenario_columns_ordered_all)) {
    scenario_col <- scenario_columns_ordered_all[i]

    # Extract mm and hour values
    numbers <- sub(".*_(\\d+)_(\\d+)$", "\\1 \\2", scenario_col)
    mm <- unlist(strsplit(numbers, " "))[1]
    hours <- as.numeric(unlist(strsplit(numbers, " "))[2]) / 3600
    title_text <- paste(mm, "mm x", hours, "hour")

    # Create weighted bin
    parish_boundaries_all_sf$Weighted_bin <- factor(cut(
        parish_boundaries_all_sf[[scenario_col]],
        breaks = c(0, 30, 60, Inf),
        include.lowest = TRUE,
        labels = c("0-30", "30-60", "60+")
    ), levels = c("0-30", "30-60", "60+"))

    # Create bivariate classification
    parish_boundaries_bi_all <- bi_class(
        parish_boundaries_all_sf,
        x = "Weighted_bin",
        y = "Population_bin",
        dim = 3
    )

    # Create plot
    p <- ggplot() +
        geom_sf(data = parish_boundaries_bi_all, aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
        bi_scale_fill(pal = "DkViolet", dim = 3, flip_axes = FALSE) +
        labs(title = title_text) +
        bi_theme() +
        theme(
            plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
            legend.position = "none"
        )

    plot_list_all[[i]] <- p
}

# STEP: Create shared legend
legend_all <- bi_legend(
    pal = "DkViolet",
    dim = 3,
    xlab = "Travel Time Increase (min.)",
    ylab = "Population Sum",
    size = 30,
    breaks = labels_all,
    flip_axes = FALSE,
    arrows = FALSE
)

# STEP: Combine all plots
weighted_delta_TT_all <- wrap_plots(plot_list_all, ncol = 3) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom") &
    plot_annotation(
        title = "Average Population Weighted Delta Travel Time Scenarios (All Facilities)",
        theme = theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
        )
    )

# STEP: Display plot
print(weighted_delta_TT_all)

# OPTIONAL: Save the combined plot and legend
# ggsave(here("Figures", "Weighted_delta_TT", "weighted_delta_TT_all.png"), plot = weighted_delta_TT_all, width = 10, height = 15, dpi = 300)
# ggsave(here("Figures", "Weighted_delta_TT", "legend_all.png"), plot = legend_all, width = 5, height = 5, dpi = 300)
```

# Highlight parishes with the highest decrease in travel time for each scenario for all facilities
```{r}	
# Format scenario titles (e.g., "100_21600" → "100 mm x 6 hours")
format_scenario_title <- function(scenario) {
    numbers <- sub(".*_(\\d+)_(\\d+)$", "\\1 \\2", scenario)
    numbers_split <- unlist(strsplit(numbers, " "))
    hours <- as.numeric(numbers_split[2]) / 3600
    paste(numbers_split[1], "mm x", hours, "hours")
}

# Scenario column names for ALL facilities
scenario_columns_ordered_all <- c(
    "Average_Weighted_Delta_TT_20_3600", "Average_Weighted_Delta_TT_20_10800", "Average_Weighted_Delta_TT_20_21600",
    "Average_Weighted_Delta_TT_40_3600", "Average_Weighted_Delta_TT_40_10800", "Average_Weighted_Delta_TT_40_21600",
    "Average_Weighted_Delta_TT_60_3600", "Average_Weighted_Delta_TT_60_10800", "Average_Weighted_Delta_TT_60_21600",
    "Average_Weighted_Delta_TT_80_3600", "Average_Weighted_Delta_TT_80_10800", "Average_Weighted_Delta_TT_80_21600",
    "Average_Weighted_Delta_TT_100_3600", "Average_Weighted_Delta_TT_100_10800", "Average_Weighted_Delta_TT_100_21600"
)

# Convert to data frame for analysis
parish_boundaries_all_df <- as.data.frame(parish_boundaries_all)

# Identify top 10 parishes for each scenario
top_parishes_list_all <- list()

for (scenario_col in scenario_columns_ordered_all) {
    if (scenario_col %in% names(parish_boundaries_all_df)) {
        travel_times <- parish_boundaries_all_df[[scenario_col]]
        if (!any(is.na(travel_times))) {
            top_parishes <- parish_boundaries_all_df[order(travel_times), ][1:10, ]
            top_parishes_list_all[[scenario_col]] <- data.frame(
                Parish_Name = top_parishes$PName2016,
                Travel_Time_Increase = top_parishes[[scenario_col]],
                Scenario = scenario_col
            )
        } else {
            warning(paste("Scenario column", scenario_col, "contains NAs. Skipping."))
        }
    } else {
        warning(paste("Scenario column", scenario_col, "not found in all dataset."))
    }
}

# Print each scenario's top 10
for (scenario in names(top_parishes_list_all)) {
    cat("\n\n============================\n")
    cat("Top 10 Parishes for Scenario:", format_scenario_title(scenario), "\n")
    cat("============================\n\n")
    print(top_parishes_list_all[[scenario]][, c("Parish_Name", "Travel_Time_Increase")])
    cat("\n")
}

# Create frequency table across all scenarios
parish_frequency_all <- table(unlist(lapply(top_parishes_list_all, function(df) df$Parish_Name)))
parish_frequency_all_df <- as.data.frame(parish_frequency_all)
colnames(parish_frequency_all_df) <- c("Parish_Name", "Frequency")
parish_frequency_all_df$Frequency <- as.numeric(parish_frequency_all_df$Frequency)
parish_frequency_all_df <- parish_frequency_all_df[order(-parish_frequency_all_df$Frequency), ]
top_10_parish_frequency_all <- head(parish_frequency_all_df, 10)

cat("\n\nTop 10 Parishes by Frequency in Top 10 Lists (All Facilities):\n")
print(top_10_parish_frequency_all)

# Highlight top 10 on map
parish_boundaries_all$Top_10 <- ifelse(parish_boundaries_all$PName2016 %in% top_10_parish_frequency_all$Parish_Name, "Top 10", "Not in Top 10")

# Centroids
parish_centroids_all <- centroids(parish_boundaries_all)
centroids_coords_all <- geom(parish_centroids_all)
parish_centroids_all_df <- as.data.frame(parish_centroids_all) %>%
    mutate(PName2016 = parish_boundaries_all$PName2016,
           Top_10 = parish_boundaries_all$Top_10) %>%
    cbind(centroids_coords_all[, c("x", "y")])

# Plot
highlighted_plot_all <- ggplot() +
    geom_spatvector(data = parish_boundaries_all, aes(fill = Top_10)) +
    geom_label(
        data = parish_centroids_all_df %>% filter(Top_10 == "Top 10"),
        aes(label = PName2016, x = x, y = y),
        color = "black", fill = "white", size = 3,
        label.padding = unit(0.2, "lines"), label.size = 0.25,
        hjust = 0.5, vjust = -0.5
    ) +
    scale_fill_manual(values = c("Top 10" = "red", "Not in Top 10" = "gray"), name = "Top 10 Parishes") +
    theme_minimal() +
    ggtitle("Top 10 Parishes with Lowest Travel Time Increase (All Facilities)") +
    theme(
        legend.position = "bottom",
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid.major = element_line(color = "gray80", size = 0.5),
        panel.grid.minor = element_line(color = "gray90", size = 0.25),
        strip.background = element_rect(fill = "white")
    )

# Display
print(highlighted_plot_all)
```

# Densitiy plots of the average weighted delta travel time for each scenario across all parishes for all facilities 
```{r}
# Define columns for ALL FACILITIES
scenario_columns_ordered_all <- c(
    "Average_Weighted_Delta_TT_20_3600", "Average_Weighted_Delta_TT_20_10800", "Average_Weighted_Delta_TT_20_21600",
    "Average_Weighted_Delta_TT_40_3600", "Average_Weighted_Delta_TT_40_10800", "Average_Weighted_Delta_TT_40_21600",
    "Average_Weighted_Delta_TT_60_3600", "Average_Weighted_Delta_TT_60_10800", "Average_Weighted_Delta_TT_60_21600",
    "Average_Weighted_Delta_TT_80_3600", "Average_Weighted_Delta_TT_80_10800", "Average_Weighted_Delta_TT_80_21600",
    "Average_Weighted_Delta_TT_100_3600", "Average_Weighted_Delta_TT_100_10800", "Average_Weighted_Delta_TT_100_21600"
)

# Compute global axis limits across all scenarios
x_limits_all <- range(unlist(parish_boundaries_all_df[scenario_columns_ordered_all]), na.rm = TRUE)
y_limits_all <- c(0, max(sapply(scenario_columns_ordered_all, function(col) {
    max(density(parish_boundaries_all_df[[col]], na.rm = TRUE)$y)
})))

# Generate density plots for each scenario
density_plot_list_all <- list()

for (i in seq_along(scenario_columns_ordered_all)) {
    scenario_col <- scenario_columns_ordered_all[i]
    if (scenario_col %in% names(parish_boundaries_all_df)) {
        p <- ggplot(parish_boundaries_all_df, aes(x = .data[[scenario_col]])) + 
            geom_density(fill = "blue", color = "black", alpha = 0.7) +
            xlim(x_limits_all) +
            ylim(y_limits_all) +
            theme_minimal() +
            ggtitle(format_scenario_title(scenario_col)) +
            xlab("Average Weighted Delta Travel Time") +
            ylab("Density") +
            theme(
                plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
                axis.title = element_text(size = 12),
                axis.text = element_text(size = 10)
            )
        density_plot_list_all[[i]] <- p
    }
}

# Combine all density plots into one grid
final_density_plot_all <- wrap_plots(density_plot_list_all, ncol = 3) + 
    plot_layout(guides = "collect") &
    plot_annotation(
        title = "Density Plots of Average Weighted Delta Travel Time Scenarios (All Facilities)",
        theme = theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
        )
    )

print(final_density_plot_all)

# Generate overlayed density plots by rainfall duration (1h, 3h, 6h)
overlayed_plot_list_all <- list()
hours_list <- c(3600, 10800, 21600)
color_palette <- c("blue", "green", "orange", "purple", "red")

for (hour in hours_list) {
    scenarios_for_hour <- grep(paste0("_", hour), scenario_columns_ordered_all, value = TRUE)
    combined_df <- data.frame()
    
    for (scenario_col in scenarios_for_hour) {
        mm_value <- sub(".*_(\\d+)_\\d+$", "\\1", scenario_col)
        temp_df <- data.frame(
            mm = as.numeric(mm_value),
            travel_times = parish_boundaries_all_df[[scenario_col]]
        )
        combined_df <- rbind(combined_df, temp_df)
    }
    
    p <- ggplot(combined_df, aes(x = travel_times, color = as.factor(mm))) + 
        geom_density(fill = NA, alpha = 0.7) +
        theme_minimal() +
        ggtitle(paste(hour / 3600, "hour(s)")) +
        xlab("Average Weighted Delta Travel Time") +
        ylab("Density") +
        scale_color_manual(values = color_palette, name = "mm Values", breaks = c(20, 40, 60, 80, 100)) +
        theme(
            plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
            axis.title = element_text(size = 12),
            axis.text = element_text(size = 10)
        )
    overlayed_plot_list_all[[as.character(hour)]] <- p
}

final_overlayed_plot_all <- wrap_plots(overlayed_plot_list_all, ncol = 1) +
    plot_annotation(
        title = "Overlayed Density Plots for Different mm by Duration (All Facilities)",
        theme = theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
        )
    )

print(final_overlayed_plot_all)
```

########################################################
Hospiital only scenarios
########################################################

# Weighted Delta Travel Time Calculation for scenarios with Hospitals only
# Load delta travel time and population distribution
```{r}
## For delta travel time go to script 05. maps_tt_and_delta_TT
### After doing so make a list using the following code:
delta_tt_list_hosp <- mget(ls(pattern = "delta_tt_hosp"))

## Load population distribution
pop_density_raster <- rast(here("AccessMod", "InAccessMod", "Uganda", "data", "zToAccessMod", "20250108115609", "pop_total_10m.tif"))

# Load the SpatVector for parish boundaries
# Load parish boundaries and rename
parish_boundaries_hosp <- vect(here("Data", "processed", "Adm_boundaries", "parishes_kampala.shp"))

# Create weighted delta TT list for hospital-only
weighted_delta_tt_list_hosp <- lapply(delta_tt_list_hosp, function(delta_tt_raster) {
    delta_tt_raster * pop_density_raster
})
names(weighted_delta_tt_list_hosp) <- sapply(names(weighted_delta_tt_list_hosp), function(name) {
    parts <- strsplit(name, "_")[[1]]
    paste(tail(parts, 2), collapse = "_")
})

# Step 2: Zonal stats for hospital-only
for (i in seq_along(weighted_delta_tt_list_hosp)) {
    sum_weighted_delta_tt <- zonal(weighted_delta_tt_list_hosp[[i]], parish_boundaries_hosp, fun = "sum", na.rm = TRUE)
    original_name <- names(weighted_delta_tt_list_hosp)[i]
    parish_boundaries_hosp[[paste0("Weighted_delta_tt_", original_name)]] <- sum_weighted_delta_tt[, 1]
}

# Step 3: Sum population per parish for hospital-only
parish_boundaries_hosp$Population_Sum <- zonal(pop_density_raster, parish_boundaries_hosp, fun = "sum", na.rm = TRUE)

# Step 4: Calculate average weighted delta TT
for (i in seq_along(weighted_delta_tt_list_hosp)) {
    original_name <- names(weighted_delta_tt_list_hosp)[i]
    parish_boundaries_hosp[[paste0("Average_Weighted_Delta_TT_", original_name)]] <-
        parish_boundaries_hosp[[paste0("Weighted_delta_tt_", original_name)]] / parish_boundaries_hosp$Population_Sum
}

```

# Create a bivariate map of accessibility reduction and population sum for hospital only scenarios
```{r}
# STEP: Convert to sf
parish_boundaries_hosp_sf <- st_as_sf(parish_boundaries_hosp)

# STEP: Make negative values positive for plotting
for (col in names(parish_boundaries_hosp_sf)[grepl("Average_Weighted_Delta_TT_", names(parish_boundaries_hosp_sf))]) {
    parish_boundaries_hosp_sf[[col]] <- abs(parish_boundaries_hosp_sf[[col]])
}

# STEP: Extract and order scenario columns
scenario_columns_hosp <- grep("Average_Weighted_Delta_TT_", names(parish_boundaries_hosp_sf), value = TRUE)
scenario_columns_ordered_hosp <- scenario_columns_hosp[order(
    as.numeric(sub(".*_(\\d+)_.*", "\\1", scenario_columns_hosp)),
    as.numeric(sub(".*_(\\d+)$", "\\1", sub(".*_\\d+_", "", scenario_columns_hosp)))
)]

# STEP: Create population bin
parish_boundaries_hosp_sf$Population_bin <- factor(cut(
    parish_boundaries_hosp_sf$Population_Sum,
    breaks = c(0, 10000, 30000, Inf),
    include.lowest = TRUE,
    labels = c("Low", "Medium", "High")
), levels = c("High", "Medium", "Low"))

# STEP: Define legend labels
labels_hosp <- list(
    bi_x = c("0-30", "30-60", "60+"),
    bi_y = c("30000+", "10000-30000", "0-10000")
)

# STEP: Initialize list of plots
plot_list_hosp <- list()

# STEP: Loop over each scenario and create the plots
for (i in seq_along(scenario_columns_ordered_hosp)) {
    scenario_col <- scenario_columns_ordered_hosp[i]

    # Extract mm and hours
    mm <- sub(".*_(\\d+)_\\d+$", "\\1", scenario_col)
    hr <- as.numeric(sub(".*_(\\d+)$", "\\1", scenario_col)) / 3600
    title_text <- paste(mm, "mm x", hr, "hour")

    # Classify into bins
    parish_boundaries_hosp_sf$Weighted_bin <- factor(cut(
        parish_boundaries_hosp_sf[[scenario_col]],
        breaks = c(0, 30, 60, Inf),
        include.lowest = TRUE,
        labels = c("0-30", "30-60", "60+")
    ), levels = c("0-30", "30-60", "60+"))

    # Bivariate classification
    parish_boundaries_bi_hosp <- bi_class(
        parish_boundaries_hosp_sf,
        x = "Weighted_bin",
        y = "Population_bin",
        dim = 3
    )

    # Plot
    p <- ggplot() +
        geom_sf(data = parish_boundaries_bi_hosp, aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
        bi_scale_fill(pal = "DkViolet", dim = 3, flip_axes = FALSE) +
        labs(title = title_text) +
        bi_theme() +
        theme(
            plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
            legend.position = "none"
        )

    plot_list_hosp[[i]] <- p
}

# STEP: Create shared legend
legend_hosp <- bi_legend(
    pal = "DkViolet",
    dim = 3,
    xlab = "Travel Time Increase (min.)",
    ylab = "Population Sum",
    size = 30,
    breaks = labels_hosp,
    flip_axes = FALSE,
    arrows = FALSE
)

# STEP: Combine and display
weighted_delta_TT_hosp <- wrap_plots(plot_list_hosp, ncol = 3) +
    plot_layout(guides = "collect") & 
    theme(legend.position = "bottom") & 
    plot_annotation(
        title = "Average Population Weighted Delta Travel Time Scenarios (Hospitals Only)",
        theme = theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
        )
    )

# Display the combined plot
print(weighted_delta_TT_hosp)
```

# Highlight parishes with the highest decrease in travel time for each scenario for all facilities
```{r}
# Function to format scenario titles (e.g., "100_21600" -> "100 mm x 6 hours")
format_scenario_title <- function(scenario) {
    numbers <- sub(".*_(\\d+)_(\\d+)$", "\\1 \\2", scenario)
    numbers_split <- unlist(strsplit(numbers, " "))
    hours <- as.numeric(numbers_split[2]) / 3600
    formatted_title <- paste(numbers_split[1], "mm x", hours, "hours")
    return(formatted_title)
}

# Use the hospital-specific SpatVector
parish_boundaries_hosp_df <- as.data.frame(parish_boundaries_hosp)

# Extract relevant columns
scenario_columns_ordered_hosp <- grep("Average_Weighted_Delta_TT_", names(parish_boundaries_hosp_df), value = TRUE)

# Initialize empty list for results
top_parishes_list_hosp <- list()

# Loop through each scenario and extract top 10 parishes
for (i in seq_along(scenario_columns_ordered_hosp)) {
    scenario_col <- scenario_columns_ordered_hosp[i]

    if (scenario_col %in% names(parish_boundaries_hosp_df)) {
        travel_times <- parish_boundaries_hosp_df[[scenario_col]]
        if (!any(is.na(travel_times))) {
            top_parishes <- parish_boundaries_hosp_df[order(travel_times), ][1:10, ]
            top_parishes_list_hosp[[scenario_col]] <- data.frame(
                Parish_Name = top_parishes$PName2016,
                Travel_Time_Increase = top_parishes[[scenario_col]],
                Scenario = scenario_col
            )
        }
    }
}

# Print top 10 per scenario
for (scenario in names(top_parishes_list_hosp)) {
    formatted_title <- format_scenario_title(scenario)
    cat("\n\n============================\n")
    cat("Top 10 Parishes for Scenario:", formatted_title, "\n")
    cat("============================\n\n")
    print(top_parishes_list_hosp[[scenario]][, c("Parish_Name", "Travel_Time_Increase")])
    cat("\n")
}

# Frequency count across scenarios
parish_frequency_hosp <- table(unlist(lapply(top_parishes_list_hosp, function(df) df$Parish_Name)))
parish_frequency_df_hosp <- as.data.frame(parish_frequency_hosp)
colnames(parish_frequency_df_hosp) <- c("Parish_Name", "Frequency")
parish_frequency_df_hosp$Frequency <- as.numeric(parish_frequency_df_hosp$Frequency)
parish_frequency_df_hosp <- parish_frequency_df_hosp[order(-parish_frequency_df_hosp$Frequency), ]
top_10_parish_frequency_hosp <- head(parish_frequency_df_hosp, 10)

cat("\n\nTop 10 Parishes by Frequency in Top 10 Lists (Hospitals Only):\n")
print(top_10_parish_frequency_hosp)

# Highlight top 10 in SpatVector
parish_boundaries_hosp$Top_10 <- ifelse(parish_boundaries_hosp$PName2016 %in% top_10_parish_frequency_hosp$Parish_Name, "Top 10", "Not in Top 10")

# Centroids and coordinates
parish_centroids_hosp <- centroids(parish_boundaries_hosp)
centroids_coords_hosp <- geom(parish_centroids_hosp)

# Add attributes to centroid df
parish_centroids_df_hosp <- as.data.frame(parish_centroids_hosp) %>%
    mutate(
        PName2016 = parish_boundaries_hosp$PName2016,
        Top_10 = parish_boundaries_hosp$Top_10,
        x = centroids_coords_hosp[, "x"],
        y = centroids_coords_hosp[, "y"]
    )

# Color palette
top_10_palette <- c("Top 10" = "red", "Not in Top 10" = "gray")

# Plot
highlighted_plot_hosp <- ggplot() +
    geom_spatvector(data = parish_boundaries_hosp, aes(fill = Top_10)) +
    geom_label(
        data = parish_centroids_df_hosp %>% filter(Top_10 == "Top 10"),
        aes(label = PName2016, x = x, y = y),
        color = "black",
        fill = "white",
        size = 3,
        label.padding = unit(0.2, "lines"),
        label.size = 0.25
    ) +
    scale_fill_manual(values = top_10_palette, name = "Top 10 Parishes") +
    theme_minimal() +
    ggtitle("Top 10 Parishes with Highest Travel Time Increase (Hospitals Only)") +
    theme(
        legend.position = "bottom",
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid.major = element_line(color = "gray80", size = 0.5),
        panel.grid.minor = element_line(color = "gray90", size = 0.25),
        strip.background = element_rect(fill = "white")
    )

# Show plot
print(highlighted_plot_hosp)
```

# Densitiy plots of the average weighted delta travel time for each scenario across all parishes for all facilities 
```{r}
# Define the list of columns for hospital-only delta TT scenarios
scenario_columns_ordered_hosp <- c(
    "Average_Weighted_Delta_TT_20_3600", "Average_Weighted_Delta_TT_20_10800", "Average_Weighted_Delta_TT_20_21600",
    "Average_Weighted_Delta_TT_40_3600", "Average_Weighted_Delta_TT_40_10800", "Average_Weighted_Delta_TT_40_21600",
    "Average_Weighted_Delta_TT_60_3600", "Average_Weighted_Delta_TT_60_10800", "Average_Weighted_Delta_TT_60_21600",
    "Average_Weighted_Delta_TT_80_3600", "Average_Weighted_Delta_TT_80_10800", "Average_Weighted_Delta_TT_80_21600",
    "Average_Weighted_Delta_TT_100_3600", "Average_Weighted_Delta_TT_100_10800", "Average_Weighted_Delta_TT_100_21600"
)

# Ensure you're using the correct hospital-specific data frame
x_limits <- range(unlist(parish_boundaries_hosp_df[scenario_columns_ordered_hosp]), na.rm = TRUE)

y_limits <- c(0, max(sapply(scenario_columns_ordered_hosp, function(col) {
    max(density(parish_boundaries_hosp_df[[col]], na.rm = TRUE)$y)
})))

# Create density plots for each scenario
density_plot_list_hosp <- list()

for (i in seq_along(scenario_columns_ordered_hosp)) {
    scenario_col <- scenario_columns_ordered_hosp[i]
    if (scenario_col %in% names(parish_boundaries_hosp_df)) {
        p <- ggplot(parish_boundaries_hosp_df, aes(x = .data[[scenario_col]])) +
            geom_density(fill = "blue", color = "black", alpha = 0.7) +
            xlim(x_limits) +
            ylim(y_limits) +
            theme_minimal() +
            ggtitle(format_scenario_title(scenario_col)) +
            xlab("Average Weighted Delta Travel Time") +
            ylab("Density") +
            theme(
                plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
                axis.title = element_text(size = 12),
                axis.text = element_text(size = 10)
            )
        density_plot_list_hosp[[i]] <- p
    }
}

# Combine into one figure
final_density_plot_hosp <- wrap_plots(density_plot_list_hosp, ncol = 3) +
    plot_layout(guides = "collect") &
    plot_annotation(
        title = "Density Plots of Average Weighted Delta Travel Time Scenarios for Hospital Only per Parish",
        theme = theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
        )
    )

# Display
print(final_density_plot_hosp)

# Overlayed density plots: same hour, varying mm
overlayed_plot_list_hosp <- list()
hours_list <- c(3600, 10800, 21600)
color_palette <- c("blue", "green", "orange", "purple", "red")

for (hour in hours_list) {
    scenarios_for_hour <- grep(paste0("_", hour), scenario_columns_ordered_hosp, value = TRUE)
    combined_df <- data.frame()

    for (scenario_col in scenarios_for_hour) {
        mm_value <- sub(".*_(\\d+)_\\d+$", "\\1", scenario_col)
        temp_df <- data.frame(
            mm = as.numeric(mm_value),
            travel_times = parish_boundaries_hosp_df[[scenario_col]]
        )
        combined_df <- rbind(combined_df, temp_df)
    }

    p <- ggplot(combined_df, aes(x = travel_times, color = as.factor(mm))) +
        geom_density(fill = NA, alpha = 0.7) +
        theme_minimal() +
        ggtitle(paste(hour / 3600, "hour(s)")) +
        xlab("Average Weighted Delta Travel Time") +
        ylab("Density") +
        scale_color_manual(values = color_palette, name = "mm Values", breaks = c(20, 40, 60, 80, 100)) +
        theme(
            plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
            axis.title = element_text(size = 12),
            axis.text = element_text(size = 10)
        )

    overlayed_plot_list_hosp[[as.character(hour)]] <- p
}

# Combine overlayed plots
final_overlayed_plot_hosp <- wrap_plots(overlayed_plot_list_hosp, ncol = 1) +
    plot_annotation(
        title = "Overlayed Density Plots for Different mm by Hours for Hospital Only",
        theme = theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
        )
    )

# Display
print(final_overlayed_plot_hosp)
```