# Loading packages

```{r package loading}
library(tidyverse)
library(haven)
library(lattice)
library(data.table)
library(knitr)
library(survey)
 options(survey.lonely.psu="adjust")
library(expss)
library(ggplot2)
library(labelled)
library(table1)
library(readxl)
library(GGally)
library(kableExtra)
library(xtable)
library(modelsummary)
library(performance)
library(finalfit)
library(naniar)
library(rms)
library(openxlsx)
library(rcompanion)
library(questionr)
library(flextable)
library(MASS)
library(car)
library(cowplot)
library(heatmaply)
library(esquisse)
library(visreg)
library(forcats) 
library(mosaic)
library(DT)
library(mice)
library(gtsummary)
library(styler)
library(languageserver)
library(httpgd)
library(plotly)
library(sf)
```

# Data handling

#### Loading data and creating subset

```{r Dota load and subsetting}
# Load the children dataset from a .SAV file 

#depending on computer C:/Users/Rik/OneDrive/PhD_Undernutrition_Uganda_Physical_Accessbility/1_Study_1_Physical_access/Data/DATA/Uganda Full/DHS Clusters/UGKR7BSV_Children/UGKR7BFL.SAV   

DHS_children_sub <- read_sav("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/1_Study_1_Physical_access/Data/DATA/Uganda Full/DHS Clusters/UGKR7BSV_Children/UGKR7BFL.SAV")


# Filter out rows with children who are no longer alive (B5 == 1)
DHS_children_sub = subset(DHS_children_sub, B5 == 1)

# Keep only children with anthropometric measurements (non-missing HW5 values)
DHS_children_sub = DHS_children_sub %>% drop_na(HW5)

Kamp = subset(DHS_children_sub, V024 == 0)

DHS_loc <- st_read("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/1_Study_1_Physical_access/Data/DATA/Uganda Full/DHS Clusters/UGGE7AFL/UGGE7AFL.shp")

# Only select 
DHS_loc <- subset(DHS_loc, ADM1NAME == "KAMPALA")


Kamp <- Kamp %>%
  group_by(V001) %>%
  summarise(Count = n()) %>%
  mutate(V001 = reorder(V001, Count)) # Reorder V001 based on Count

# Rename V001 to DHSCUST
names(Kamp)[names(Kamp) == "V001"] <- "DHSCLUST"

DHS_loc <- merge(DHS_loc, Kamp, by = "DHSCLUST", all.x = TRUE)

# Read the shapefile
parishes <- st_read("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/Processed/Adm_boundaries_parishes/Parishes_Kamp_UTM36N.shp")

# Reproject the parishes shapefile to the same projection as the DHS location data
parishes <- st_transform(parishes, st_crs(DHS_loc))

map = ggplot() +
  geom_sf(data = parishes, fill = "lightgray", color = "black") +  # Add the shapefile as a base layer
  geom_point(data = DHS_loc, aes(x = LONGNUM, y = LATNUM, size = Count, color = Count), alpha = 0.7) +
  labs(title = "Distribution of Children per Cluster in Kampala", x = "Longitude", y = "Latitude") +
  scale_color_gradient(low = "blue", high = "red") +
  scale_size(range = c(2, 10), name = "Count") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        legend.position = "right")

# Summary statistics of the count variable
summary(Kamp$Count)

# Create a bar plot of the cluster distribution
ggplot(Kamp, aes(x = DHSCLUST, y = Count)) +
  geom_bar(stat = "identity", fill = "steelblue") + # Use 'identity' to use Count values directly
  labs(title = "Distribution of Clusters in Kampala", x = "Cluster", y = "Count") +
  theme_bw() + # Use a black and white theme
  theme(plot.title = element_text(size = 14, face = "bold"), # Increase title size and make it bold
    axis.title = element_text(size = 12), # Increase axis label size
    axis.text = element_text(size = 10), # Increase axis tick label size
    panel.grid.major = element_blank(), # Remove major grid lines
    panel.grid.minor = element_blank(), # Remove minor grid lines
    panel.border = element_blank(), # Remove plot border
    legend.position = "none") + # Remove legend
  scale_y_continuous(breaks = seq(0, max(Kamp$Count), by = 1)) + # Add more y ticks
  geom_hline(yintercept = seq(0, max(Kamp$Count), by = 1), linetype = "dotted", color = "gray") # Add dotted lines


