# InAccessmod.qmd Documentation

## Overview
The `InAccessmod.qmd` file contains the data preparation code for the Accessmod tool. This tool is used for analyzing physical accessibility in the context of undernutrition in Uganda.

## Purpose
The purpose of this file is to preprocess the input data required by the Accessmod tool. It performs various data cleaning, transformation, and aggregation tasks to ensure that the data is in the appropriate format for analysis.

## Usage
To use this file, follow these steps:
1. Ensure that the required input data files are available.
2. Open the `InAccessmod.qmd` file in a compatible code editor or environment.
3. Review the code and make any necessary modifications based on your specific data requirements.
4. Execute the code to preprocess the data.

## Dependencies
This file may have dependencies on other code files or libraries. Ensure that all required dependencies are installed and accessible before running the code.

## Output
The output of this file is the preprocessed data that can be used as input for the Accessmod tool. The specific format and structure of the output data will depend on the requirements of the tool.

## Notes
- It is recommended to review the code and documentation of the Accessmod tool to understand the expected input data format and any additional preprocessing steps that may be required.

## References
- [Accessmod Tool Documentation](https://www.accessmod.org/)

# Install the inAccessMod package from GitHub & load the library
```{r}
if (!require("devtools")) install.packages("devtools")

devtools::install_github("unige-geohealth/inAccessMod", build_vignettes = TRUE, dependencies = TRUE)

library(inAccessMod)
library(sf)
library(dplyr)
library(doParallel)
library(foreach)
library(snow)
library(terra)
library(raster)
library(ggplot2)
library(gridExtra)
```

# Initialize the project for Uganda
```{r}
mainPath <- "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod"

#initiate_project(mainPath) # Only run this line once to create the project folder

country <- "Uganda"
```

# Download and process the required input data for AccessMod
## Administrative boundaries
```{r}
# Parish boundaries
inputPath_adm = "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/Processed/Adm_boundaries_parishes"
copy_input(mainPath, country, inputPath_adm)
```

## Population data
```{r}
#Total population/women of reproductive age (WRA) population/under 5 population
inputPath_pop <- "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/WorldPop_density/Inaccessmod"
copy_input(mainPath, country, inputPath_pop)
```

## Digital Elevation Model (DEM)
```{r}
inputPath_DEM <- "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/DEM/Uganda_SRTM30m"
copy_input(mainPath, country, inputPath_DEM)
```

## Land cover data
```{r}
inputPath_LC <- "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/Landcover/terrascope_download_20240305_144418/ESA_WorldCover_10m_2021_v200_N00E030_Map"
copy_input(mainPath, country, inputPath_LC)
```

## Health facilities data
```{r}
# Health facilities all/ hospitals only
inputPath_HF <- "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/Facilities/suhsharan_health_facilities_UG"
copy_input(mainPath, country, inputPath_HF)
```

## OpenStreetMap (OSM) data
```{r}
# Roads data
download_osm(x = "roads", mainPath, country, alwaysDownload = TRUE, countryName = TRUE, mostRecent = NULL, defaultClasses = TRUE)
```

# Flood data
```{r}
inputPath_flood <- "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/Flood_extents_Kampala/Flood_extents_full"
copy_input(mainPath, country, inputPath_flood)

```

# Set the projected coordinate reference system
```{r}
set_projection(mainPath, country, mostRecent = TRUE, alwaysSet = TRUE, bestCRS = TRUE)
```

# Process all downloaded layers to prepare them for AccessMod analysis
```{r}
process_inputs(mainPath, country, selectedInputs = NULL, mostRecent = TRUE, alwaysProcess = TRUE, defaultMethods = TRUE, changeRes = TRUE, newRes = 10, popCorrection = TRUE, gridRes = 10)

# In QGIS clip the road data to the study area as this is not done by the process_inputs function
```

# Label land cover classes for AccessMod
```{r}
label_landcover(mainPath, country, mostRecent = TRUE, overwrite = F, defaultLabels = FALSE)

# See WorldCover website for the land cover classes
```

# Compile all processed data for easier import into AccessMod
```{r}
compile_processed_data(mainPath, country, mostRecent = TRUE)
```

# Create merged landcover layer for AccessMod
## First seperate the flood extents based on the unique values of threshold, rainfall, and duration
```{r}
# Flood(x1)x(x2)x(x3), MLC = merged land cover, x1 = rainfall (mm), x2 = rain duration (h), x3 = flood depth (m)
# Load full data
flood_full <- st_read("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Uganda/data/vNaturalPolygons/20240311111000/processed/20240311111005")

# Define unique values for threshold, rainfall, and duration
threshold_values <- c(0.1, 0.2, 0.3)
rainfall_values <- c(20, 40, 60, 80, 100)
duration_values <- c(3600, 10800, 21600)

# Seperate each flood extent based on the unique values of threshold, rainfall, and duration
for (threshold in threshold_values) {
    for (rainfall in rainfall_values) {
        for (duration in duration_values) {
            
            # Filter data based on current combination
            filtered_data <- flood_full[flood_full$threshold == threshold & flood_full$rainfall == rainfall & flood_full$duration == duration, ]
            
            # Check if the filtered data is empty
            if (nrow(filtered_data) > 0) {
                # Define sub-folder name
                subfolder_name <- paste0("flood_", threshold, "_", rainfall, "_", duration)
                
                # Create sub-folder if it doesn't exist
                dir_path <- paste0("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/Flood_extents_Kampala/", subfolder_name)
                if (!dir.exists(dir_path)) {
                    dir.create(dir_path)
                }
                
                # Define shapefile name
                shapefile_name <- paste0(subfolder_name, ".shp")
                
                # Write filtered data to shapefile in the sub-folder
                st_write(filtered_data, file.path(dir_path, shapefile_name), append = FALSE)
            }
        }
    }
}

# Load all flood_extents from subfolders
subfolder_path <- "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/Flood_extents_Kampala"
flood_extents <- list.files(path = subfolder_path, pattern = "\\.shp$", full.names = TRUE, recursive = TRUE)

# Read shapefiles into a list
flood_list <- lapply(flood_extents, function(file) {
    flood <- st_read(file)
})

# Name the flood extents in flood_list based on their subfolder name
flood_extents_t <- list.files(path = subfolder_path, pattern = "\\.shp$", full.names = F, recursive = TRUE)

flood_extents_t <- gsub("/.*", "", flood_extents_t)

for (i in 1:length(flood_list)) {
    flood_list[[i]]$Scenario <- gsub("flood_", "", flood_extents_t[i])
}

# Remove the other variable names 
flood_list <- lapply(flood_list, function(x) {
    x$threshold <- NULL
    x$rainfall <- NULL
    x$duration <- NULL
    return(x)
})

# Create a list of plots for each item in flood_list
plot_list <- lapply(flood_list[1:6], function(data) {
    ggplot() + 
    geom_sf(data = data, aes(fill = Scenario), color = "blue") + 
    theme_minimal() +
    guides(fill = guide_legend(title = "Scenario", position = "bottom"))
})

# Combine the plots into a single plot
grid.arrange(grobs = plot_list, ncol = 2, nrow = 3)

```

## Merge landcover layers for each flood extent
```{r}
# Create merged landcover layers for each flood extent
merged_landcovers <- list()

# Base MLC file path
MLC0 <- raster("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Merged landcovers/MLC0_major_roads_10m.tif")

MLC0 <- projectRaster(MLC0, crs = "+init=EPSG:21096")

for (i in 1:length(flood_list)){
    landcov <- rast(MLC0) # load landcover raster
    grid_landcov <- rast(ext(landcov)) # create a grid with the same extent as the landcover raster 
    res(grid_landcov) <- 10 # set the resolution of the grid to 10m
    crs(grid_landcov) <- crs(landcov) # set the crs of the grid to the same as the landcover raster
    landcover_resamp  <- resample(landcov, grid_landcov, "near") # resample landcover to 10m
    flood_extent <- flood_list[[i]] %>% # load water polygons
        mutate(class = 1) %>% # give all rivers and lakes integer class 1 so that later they can be classified as barriers
        st_transform(st_crs(landcov)) %>% # transform to the same crs as the landcover raster
        as(., "Spatial") %>% # convert to sp object
        vect() # convert to terra object
    flood_extent_rast <- terra::rasterize(flood_extent, grid_landcov, "class", touches = TRUE) # rasterize water polygons
    values(flood_extent_rast)[values(flood_extent_rast) == 0] <- NA # make background value NA
    landcover_merge <- terra::merge(flood_extent_rast, landcov) # merge landcover and water raster
    writeRaster(landcover_merge, paste0("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Merged landcovers/MLC_", flood_extents_t[i], ".tif"), overwrite=T) # write merged landcover to file
}
```
