# InAccessmod.qmd Documentation

## Overview
The `InAccessmod.qmd` file contains the data preparation code for the Accessmod tool. This tool is used for analyzing physical accessibility in the context of undernutrition in Uganda.

## Purpose
The purpose of this file is to preprocess the input data required by the Accessmod tool. It performs various data cleaning, transformation, and aggregation tasks to ensure that the data is in the appropriate format for analysis.

## Usage
To use this file, follow these steps:
1. Ensure that the required input data files are available.
2. Open the `InAccessmod.qmd` file in a compatible code editor or environment.
3. Review the code and make any necessary modifications based on your specific data requirements.
4. Execute the code to preprocess the data.

## Dependencies
This file may have dependencies on other code files or libraries. Ensure that all required dependencies are installed and accessible before running the code.

## Output
The output of this file is the preprocessed data that can be used as input for the Accessmod tool. The specific format and structure of the output data will depend on the requirements of the tool.

## Notes
- It is recommended to review the code and documentation of the Accessmod tool to understand the expected input data format and any additional preprocessing steps that may be required.

## References
- [Accessmod Tool Documentation](https://www.accessmod.org/)

# Install the inAccessMod package from GitHub
```{r}
if (!require("devtools")) install.packages("devtools")

devtools::install_github("unige-geohealth/inAccessMod", build_vignettes = TRUE, dependencies = TRUE)

library(inAccessMod)

```

# Initialize the project for Uganda
```{r}
mainPath <- "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod"

# initiate_project(mainPath) # Only run this line once to create the project folder

country <- "Uganda"
```

# Download and process the required input data for AccessMod
## Administrative boundaries
```{r}
# Parish boundaries

inputPath_adm = "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/Processed/Adm_boundaries_parishes"

copy_input(mainPath, country, inputPath_adm)
```

## Population data
```{r}
#Total population/women of reproductive age (WRA) population/under 5 population
inputPath_pop <- "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/WorldPop_density/Inaccessmod"
copy_input(mainPath, country, inputPath_pop)

```

## Digital Elevation Model (DEM)
```{r}
inputPath_DEM <- "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/DEM/Uganda_SRTM30m"
copy_input(mainPath, country, inputPath_DEM)
```

## Land cover data
```{r}
inputPath_LC <- "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/Landcover/terrascope_download_20240305_144418/ESA_WorldCover_10m_2021_v200_N00E030_Map"
copy_input(mainPath, country, inputPath_LC)
```

## Health facilities data
```{r}
# Health facilities all/ hospitals only
inputPath_HF <- "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/Facilities/suhsharan_health_facilities_UG"

copy_input(mainPath, country, inputPath_HF)
```

## OpenStreetMap (OSM) data
```{r}
# Roads and water lines
download_osm(x = "roads", mainPath, country, alwaysDownload = TRUE, countryName = TRUE, mostRecent = NULL, defaultClasses = TRUE)
download_osm("waterLines", mainPath, country, alwaysDownload = TRUE, countryName = TRUE, mostRecent = NULL, defaultClasses = TRUE)
```

# Set the projected coordinate reference system
```{r}
set_projection(mainPath, country, mostRecent = TRUE, alwaysSet = TRUE, bestCRS = TRUE)
```

# Process all downloaded layers to prepare them for AccessMod analysis
```{r}
process_inputs(mainPath, country, selectedInputs = NULL, mostRecent = TRUE, alwaysProcess = TRUE, defaultMethods = TRUE, changeRes = TRUE, newRes = 100, popCorrection = TRUE, gridRes = 30)

?process_inputs
```

# Label land cover classes for AccessMod
```{r}
label_landcover(mainPath, country, mostRecent = TRUE, overwrite = TRUE, defaultLabels = TRUE)
```

# Compile all processed data for easier import into AccessMod
```{r}
compile_processed_data(mainPath, country, mostRecent = TRUE)
```

# Create merged landcover layer for AccessMod
```{r}
# Flood(x1)x(x2)x(x3), MLC = merged land cover, x1 = rainfall (mm), x2 = rain duration (h), x3 = flood depth (m)
library(sf)

# Load full data
flood_full <- st_read("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/Flood_extents_Kampala/flood_extents.gpkg")

# Define unique values for threshold, rainfall, and duration
threshold_values <- c(0.1, 0.2, 0.3)
rainfall_values <- c(20, 40, 60, 80, 100)
duration_values <- c(3600, 10800, 21600)

# Seperate each flood extent based on the unique values of threshold, rainfall, and duration
for (threshold in threshold_values) {
    for (rainfall in rainfall_values) {
        for (duration in duration_values) {
            
            # Filter data based on current combination
            filtered_data <- flood_full[flood_full$threshold == threshold & flood_full$rainfall == rainfall & flood_full$duration == duration, ]
            
            # Check if the filtered data is empty
            if (nrow(filtered_data) > 0) {
                # Define sub-folder name
                subfolder_name <- paste0("flood_", threshold, "_", rainfall, "_", duration)
                
                # Create sub-folder if it doesn't exist
                dir_path <- paste0("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/Flood_extents_Kampala/", subfolder_name)
                if (!dir.exists(dir_path)) {
                    dir.create(dir_path)
                }
                
                # Define shapefile name
                shapefile_name <- paste0(subfolder_name, ".shp")
                
                # Write filtered data to shapefile in the sub-folder
                st_write(filtered_data, file.path(dir_path, shapefile_name), append = FALSE)
            }
        }
    }
}
```

```{r}
# Libraries
library(dplyr)
library(sf)
library(doParallel)
library(foreach)
library(snow)
library(terra)

# Create merged landcover layers for each flood extent
landcover_merge_function <- function(i) {
    landcov <- rast(landcover[i])
    grid_landcov <- rast(ext(landcov))
    res(grid_landcov) <- 100
    crs(grid_landcov) <- crs(landcov)
    landcover_resamp <- resample(landcov, grid_landcov, "ngb")
    
    roads_new <- st_read(roads[i]) %>%
        group_by(highway) %>%
        summarize() %>%
        filter(highway %in% as.factor(road_classes)) %>%
        mutate(class = recode(highway, !!!reclass_roads),
                class = as.integer(class)) %>%
        st_transform(st_crs(landcov)) %>%
        as(., "Spatial") %>%
        vect()
    roads_rast <- terra::rasterize(roads_new, grid_landcov, "class", touches = TRUE)
    
    water_lines_new <- st_read(water_lines[i]) %>%
        group_by(waterway) %>%
        summarize() %>%
        filter(waterway == "river") %>%
        mutate(class = 1) %>%
        st_transform(st_crs(landcov)) %>%
        as(., "Spatial") %>%
        vect()
    water_lines_rast <- terra::rasterize(water_lines_new, grid_landcov, field = "class", touches = TRUE)
    values(water_lines_rast)[values(water_lines_rast) == 0] <- NA
    
    water_poly_new <- st_read(water_poly[i]) %>%
        mutate(class = 1) %>%
        st_transform(st_crs(landcov)) %>%
        as(., "Spatial") %>%
        vect()
    water_poly_rast <- terra::rasterize(water_poly_new, grid_landcov, "class", touches = TRUE)
    values(water_poly_rast)[values(water_poly_rast) == 0] <- NA
    
    landcover_merge <- terra::merge(water_lines_rast, landcover_resamp)
    landcover_merge <- terra::merge(water_poly_rast, landcover_merge)
    landcover_merge <- terra::merge(roads_rast, landcover_merge)
    values(landcover_merge)[values(landcover_merge) == 1] <- NA
    
    writeRaster(landcover_merge, paste0("data/processed/rLandcover_merge/landcover_merge_", file_name[i], ".tif"), overwrite = TRUE)
}

# parallel processing of landcover merge
cores <- 3 # not to overload your computer
cl <- makeCluster(cores) 
registerDoParallel(cl)

terraOptions(tempdir="/temp/") # set temporary folder for data processing, so R environment does not have memory issues
system.time(foreach(i = 1:length(landcover), .packages= c("sf", "dplyr", "terra")) %dopar% {landcover_merge_function(i)})


```