# Merged Landcovers

This code is used to create merged landcover layers for each flood extent in the AccessMod analysis. It takes flood extent shapefiles and merges them with a base landcover raster to create new landcover layers.

## Step 1: Seperate the flood extents

The code first separates the flood extents based on the unique values of threshold, rainfall, and duration. It loads the full flood extent data and defines the unique values for threshold, rainfall, and duration. Then, it iterates through each combination of threshold, rainfall, and duration and filters the flood extent data based on the current combination. If the filtered data is not empty, it creates a sub-folder for the flood extent and writes the filtered data to a shapefile in the sub-folder.

## Step 2: Load and name the flood extents

The code then loads all the flood extents from the sub-folders and reads them into a list. It names the flood extents in the list based on their sub-folder name. It also removes the other variable names from the flood extents.

## Step 3: Merge landcover layers for each flood extent

The code creates merged landcover layers for each flood extent. It initializes an empty list to store the merged landcover layers. It loads a base landcover raster and sets its projection. Then, it iterates through each flood extent in the flood list and performs the following steps:

1. Loads the landcover raster and creates a grid with the same extent and resolution.
2. Resamples the landcover raster to the grid resolution.
3. Transforms the flood extent polygons to the same projection as the landcover raster.
4. Rasterizes the flood extent polygons using the grid as a template.
5. Sets the background value of the rasterized flood extent to NA.
6. Merges the rasterized flood extent with the landcover raster.
7. Writes the merged landcover layer to a file.

The resulting merged landcover layers are saved in the "Merged landcovers" folder.

# R code
## Load libraries
```{r}
library(sf)
library(dplyr)
library(foreach)
library(snow)
library(terra)
library(raster)
library(ggplot2)
library(gridExtra)
```

# Create merged landcover layer for AccessMod
## First seperate the flood extents based on the unique values of threshold, rainfall, and duration
```{r}
# Flood(x1)x(x2)x(x3), MLC = merged land cover, x1 = rainfall (mm), x2 = rain duration (h), x3 = flood depth (m)
# Load full data
flood_full <- st_read("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Uganda/data/vNaturalPolygons/20240311111000/processed/20240311111005")

# Define unique values for threshold, rainfall, and duration
threshold_values <- c(0.1, 0.2, 0.3)
rainfall_values <- c(20, 40, 60, 80, 100)
duration_values <- c(3600, 10800, 21600)

# Seperate each flood extent based on the unique values of threshold, rainfall, and duration
for (threshold in threshold_values) {
    for (rainfall in rainfall_values) {
        for (duration in duration_values) {
            
            # Filter data based on current combination
            filtered_data <- flood_full[flood_full$threshold == threshold & flood_full$rainfall == rainfall & flood_full$duration == duration, ]
            
            # Check if the filtered data is empty
            if (nrow(filtered_data) > 0) {
                # Define sub-folder name
                subfolder_name <- paste0("flood_", threshold, "_", rainfall, "_", duration)
                
                # Create sub-folder if it doesn't exist
                dir_path <- paste0("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/Flood_extents_Kampala/", subfolder_name)
                if (!dir.exists(dir_path)) {
                    dir.create(dir_path)
                }
                
                # Define shapefile name
                shapefile_name <- paste0(subfolder_name, ".shp")
                
                # Write filtered data to shapefile in the sub-folder
                st_write(filtered_data, file.path(dir_path, shapefile_name), append = FALSE)
            }
        }
    }
}

# Load all flood_extents from subfolders
subfolder_path <- "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/RAW/Flood_extents_Kampala"
flood_extents <- list.files(path = subfolder_path, pattern = "\\.shp$", full.names = TRUE, recursive = TRUE)

# Read shapefiles into a list
flood_list <- lapply(flood_extents, function(file) {
    flood <- st_read(file)
})

# Name the flood extents in flood_list based on their subfolder name
flood_extents_t <- list.files(path = subfolder_path, pattern = "\\.shp$", full.names = F, recursive = TRUE)

flood_extents_t <- gsub("/.*", "", flood_extents_t)

for (i in 1:length(flood_list)) {
    flood_list[[i]]$Scenario <- gsub("flood_", "", flood_extents_t[i])
}

# Remove the other variable names 
flood_list <- lapply(flood_list, function(x) {
    x$threshold <- NULL
    x$rainfall <- NULL
    x$duration <- NULL
    return(x)
})
```

## Merge landcover layers for each flood extent
```{r}
# Create merged landcover list
merged_landcovers <- list()

# Base MLC file path
MLC0 <- raster("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Merged landcovers/MLC0_major_roads_10m.tif")

## Set projection
MLC0 <- projectRaster(MLC0, crs = "+init=EPSG:21096")

## Iterate through each flood extent in the flood list and create merged landcover layers for each flood scenario 
for (i in 1:length(flood_list)){
    landcov <- rast(MLC0) # load landcover raster
    grid_landcov <- rast(ext(landcov)) # create a grid with the same extent as the landcover raster 
    res(grid_landcov) <- 10 # set the resolution of the grid to 10m
    crs(grid_landcov) <- crs(landcov) # set the crs of the grid to the same as the landcover raster
    landcover_resamp  <- resample(landcov, grid_landcov, "near") # resample landcover to 10m
    flood_extent <- flood_list[[i]] %>% # load water polygons
        mutate(class = 1) %>% # give all rivers and lakes integer class 1 so that later they can be classified as barriers
        st_transform(st_crs(landcov)) %>% # transform to the same crs as the landcover raster
        as(., "Spatial") %>% # convert to sp object
        vect() # convert to terra object
    flood_extent_rast <- terra::rasterize(flood_extent, grid_landcov, "class", touches = TRUE) # rasterize water polygons
    values(flood_extent_rast)[values(flood_extent_rast) == 0] <- NA # make background value NA
    landcover_merge <- terra::merge(flood_extent_rast, landcov) # merge landcover and water raster
    writeRaster(landcover_merge, paste0("D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/InAccessmod/Merged landcovers/MLC_", flood_extents_t[i], ".tif"), overwrite=T) # write merged landcover to file
}
```
