---
title: "Data Analysis Plan: Assessing the Impact of Flooding on Healthcare Accessibility in Kampala"
format: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.show='hide', eval=FALSE)

#fig.show='hide', eval=FALSE
```

# Loading packages

```{r package loading}
# List of required packages
required_packages <- c(
  "NCmisc", "tidyverse", "haven", "lattice", "data.table", "knitr", "survey",
  "expss", "labelled", "table1", "readxl", "GGally", "kableExtra", "xtable",
  "modelsummary", "performance", "finalfit", "naniar", "rms", "openxlsx",
  "rcompanion", "questionr", "flextable", "MASS", "car", "cowplot", "heatmaply",
  "esquisse", "visreg", "forcats", "mosaic", "DT", "mice", "gtsummary", "styler",
  "languageserver", "httpgd", "DBI", "RSQLite")
)

## Function to check and install missing packages
install_missing_packages <- function(packages) {
  missing_packages <- packages[!(packages %in% installed.packages()[,"Package"])]
  if (length(missing_packages) > 0) {
    install.packages(missing_packages, dependencies = TRUE)
  }
}

##Check and install missing packages
install_missing_packages(required_packages)


# Load the installed packages
library(tidyverse)
library(haven)
library(lattice)
library(data.table)
library(knitr)
library(survey)
  options(survey.lonely.psu = "adjust")
library(expss)
library(labelled)
library(table1)
library(readxl)
library(GGally)
library(kableExtra)
library(xtable)
library(modelsummary)
library(performance)
library(finalfit)
library(naniar)
library(rms)
library(openxlsx)
library(rcompanion)
library(questionr)
library(flextable)
library(MASS)
library(car)
library(cowplot)
library(heatmaply)
library(esquisse)
library(visreg)
library(forcats)
library(mosaic)
library(DT)
library(mice)
library(gtsummary)
library(languageserver)
library(httpgd)
library(DBI)
library(RSQLite)
```


# Data handling
## Loading data and creating subset

```{r Data load and subsetting}
# Load the children dataset from a .SAV file 
## depending on computer  "C:/Users/Rik/OneDrive/PhD_Undernutrition_Uganda_Physical_Accessbility/1_Study_1_Physical_access/Data/DATA/Uganda Full/DHS Clusters/UGKR7BSV_Children/UGKR7BFL.SAV"
##                        "D:/Onedrive/PhD_Undernutrition_Uganda_Physical_Accessbility/1_Study_1_Physical_access/Data/DATA/Uganda Full/DHS Clusters/UGKR7BSV_Children/UGKR7BFL.SAV"

DHS_children_kamp <- read_sav("C:/Users/Rik/OneDrive/PhD_Undernutrition_Uganda_Physical_Accessbility/1_Study_1_Physical_access/Data/DATA/Uganda Full/DHS Clusters/UGKR7BSV_Children/UGKR7BFL.SAV"
)  %>%
  dplyr::filter(V024 == 0)  %>% #only households in Kampala 
  dplyr::select( 
    CASEID,
    V001, V005, V002,
    V023, V024,
    V149, V729,
    V191, V212,
    V445, V714,
    BIDX, B0, B4,
    B5,   BORD,
    HW5,  HW8,
    HW11, M19,
    B11,  H11,
    V113, M5,
    V136,
    V481, H22,
    V505, H7
  ) # only select relevant variables


con <- dbConnect(RSQLite::SQLite(), dbname = "C:/Users/Rik/OneDrive/PhD_Undernutrition_Uganda_Physical_Accessbility/2_Study_2_Flooding_Analysis/Data/Processed/Simulation/DHS_Clusters_Travel_time_simulation.sqlite")
df1 <- dbReadTable(con, "dhs_clusters_travel_time_simulation")

df1_mean <- df1 %>%
  group_by(dhsclust) %>%
  dplyr::summarize(mean_travel_time = mean(sample_1, na.rm = TRUE))

DHS_children_kamp <- merge(DHS_children_kamp, df1_mean, 
  by.x = "V001", 
  by.y = "dhsclust"
)

```

## Removing unnecessary SPSS metadata

```{r removing metadata}
# Remove unnecessary SPSS metadata from the children dataset
## Remove variable formats
DHS_children_kamp <- zap_formats(DHS_children_kamp)

## Remove variable widths
DHS_children_kamp <- zap_widths(DHS_children_kamp)

## Remove variable labels
DHS_children_kamp <- zap_labels(DHS_children_kamp)

## Remove variable label
DHS_children_kamp <- zap_label(DHS_children_kamp)
```

## Defining missing data
```{r missing data}
```{r missing data}
# Replace specific values with NA (missing data) in various columns
## For column HW5, replace 9998 and 9999 with NA
DHS_children_kamp$HW5[DHS_children_kamp$HW5 == 9998] = NA
DHS_children_kamp$HW5[DHS_children_kamp$HW5 == 9999] = NA

## For column HW8, replace 9998 and 9999 with NA
DHS_children_kamp$HW8[DHS_children_kamp$HW8 == 9998] = NA
DHS_children_kamp$HW8[DHS_children_kamp$HW8 == 9999] = NA

## For column HW11, replace 9998 and 9999 with NA
DHS_children_kamp$HW11[DHS_children_kamp$HW11 == 9998] = NA
DHS_children_kamp$HW11[DHS_children_kamp$HW11 == 9999] = NA

## For column V149, replace 9 with NA
DHS_children_kamp$V149[DHS_children_kamp$V149 == 9] = NA

## For column V729, replace 9 with NA
DHS_children_kamp$V729[DHS_children_kamp$V729 == 9] = NA

## For column V445, replace 9998 and 9999 with NA
DHS_children_kamp$V445[DHS_children_kamp$V445 == 9998] = NA
DHS_children_kamp$V445[DHS_children_kamp$V445 == 9999] = NA

## For column V714, replace 9 with NA
DHS_children_kamp$V714[DHS_children_kamp$V714 == 9] = NA

## For column M19, replace 9996, 9998, and 9999 with NA
DHS_children_kamp$M19[DHS_children_kamp$M19 == 9996] = NA
DHS_children_kamp$M19[DHS_children_kamp$M19 == 9998] = NA
DHS_children_kamp$M19[DHS_children_kamp$M19 == 9999] = NA

## For column H11, replace 8 and 9 with NA
DHS_children_kamp$H11[DHS_children_kamp$H11 == 8] = NA
DHS_children_kamp$H11[DHS_children_kamp$H11 == 9] = NA

## For column M5, replace 93, 94, 97, 98, and 99 with NA
DHS_children_kamp$M5[DHS_children_kamp$M5 == 93] = NA
DHS_children_kamp$M5[DHS_children_kamp$M5 == 94] = NA
DHS_children_kamp$M5[DHS_children_kamp$M5 == 97] = NA
DHS_children_kamp$M5[DHS_children_kamp$M5 == 98] = NA
DHS_children_kamp$M5[DHS_children_kamp$M5 == 99] = NA

## For column V113, replace 99, 97, and 96 with NA
DHS_children_kamp$V113[DHS_children_kamp$V113 == 99] = NA
DHS_children_kamp$V113[DHS_children_kamp$V113 == 97] = NA
DHS_children_kamp$V113[DHS_children_kamp$V113 == 96] = NA

## For column H7, replace 8 and 9 with NA
DHS_children_kamp$H7[DHS_children_kamp$H7 == 8] = NA
DHS_children_kamp$H7[DHS_children_kamp$H7 == 9] = NA

## For column V481, replace 9 with NA
DHS_children_kamp$V481[DHS_children_kamp$V481 == 9] = NA

## For column H22, replace 8 and 9 with NA
DHS_children_kamp$H22[DHS_children_kamp$H22 == 8] = NA
DHS_children_kamp$H22[DHS_children_kamp$H22 == 9] = NA

## For column V505, replace 98 and 99 with NA
DHS_children_kamp$V505[DHS_children_kamp$V505 == 98] = NA
DHS_children_kamp$V505[DHS_children_kamp$V505 == 99] = NA
```

## Recoding variables

```{r recoding}
# Transforming variables into more interpretable formats
## Birth interval: Dichotomize into more or less than 24 months
DHS_children_kamp$B11 = ifelse(DHS_children_kamp$B11 >= 24, 0, 1)

## Combined diarrhea in the last 24 hours and 2 weeks into one category
DHS_children_kamp$H11 = ifelse(DHS_children_kamp$H11 == 0, 0, 1)

## Recode source of drinking water into simplified categories
DHS_children_kamp$V113 = expss::recode(DHS_children_kamp$V113, 
  11 %thru% 14 ~ 1, 21 ~ 2, 31 %thru% 32 ~ 3, 41 %thru% 43 ~ 4, 51 ~ 5, 63 ~ 6, 71 ~ 6)

## Convert the variables into factors
DHS_children_kamp$B11 = factor(DHS_children_kamp$B11,
  levels = c(0, 1),
  labels = c("Birth interval more than 24 months", "Birth interval less than 24 months")
)

DHS_children_kamp$H11 = factor(DHS_children_kamp$H11,
  levels = c(0, 1),
  labels = c("No diarrhea in the last 2 weeks", "Had diarrhea in the last 2 weeks")
)

## Recode the sex of the child
DHS_children_kamp$B4 = ifelse(DHS_children_kamp$B4 == 1, 0, 1) %>% factor(levels = c(0, 1), labels = c("Male","Female"))

## Create a factor for twin births
DHS_children_kamp$B0 = factor(DHS_children_kamp$B0, levels = c(0:3),
  labels = c("Single Birth", "1st of multiple", "2nd of multiple", "3rd of multiple"))

## Recode DPT 3 into yes and no answers and make it a factor
DHS_children_kamp$H7 = ifelse(DHS_children_kamp$H7 == 0, 0, 1)
DHS_children_kamp$H7 = factor(DHS_children_kamp$H7, 
  levels = c(0, 1), 
  labels = c("No have not received", "Yes, received DPT 3")
)

## Set other variables as factors
DHS_children_kamp$V149 = factor(DHS_children_kamp$V149, 
  levels = c(0, 1, 2, 3, 4, 5), 
  labels = c("No education", "Incomplete Primary", "Complete primary", "Incomplete secondary", "Complete secondary", "Higher")
)

DHS_children_kamp$V729 = factor(DHS_children_kamp$V729, 
  levels = c(0, 1, 2, 3, 4, 5), 
  labels = c("No education", "Incomplete Primary", "Complete primary", "Incomplete secondary", "Complete secondary", "Higher")
)

DHS_children_kamp$V714 = factor(DHS_children_kamp$V714, 
  levels = c(0, 1), 
  labels = c("Not working", "Currently working")
)

DHS_children_kamp$V113 = factor(DHS_children_kamp$V113, 
  levels = c(1, 2, 3, 4, 5, 6), 
  labels = c("Piped Water", "Tube Well Water", "Dug Well (Open/Protected", "Surface from Spring", "Rainwater", "Other")
)

DHS_children_kamp$V481 = factor(DHS_children_kamp$V481, 
  levels = c(0, 1), 
  labels = c("No", "Yes")
)

DHS_children_kamp$H22 = factor(DHS_children_kamp$H22, 
  levels = c(0, 1), 
  labels = c("No", "Yes")
)

## Scaling certain variables for better interpretation
DHS_children_kamp$HW5 = DHS_children_kamp$HW5 / 100
DHS_children_kamp$HW8 = DHS_children_kamp$HW8 / 100
DHS_children_kamp$HW11 = DHS_children_kamp$HW11 / 100
DHS_children_kamp$V445 = DHS_children_kamp$V445 / 100
DHS_children_kamp$V191 = DHS_children_kamp$V191 / 1000
```

## Providing labels

```{r providing labels}
# Assigning variable labels to the dataset
var_lab(DHS_children_kamp$V001) = "Cluster number"
var_lab(DHS_children_kamp$V005) = "Women's individual sample weight"
var_lab(DHS_children_kamp$V023) = "Stratification used in sample design"
var_lab(DHS_children_kamp$BORD) = "Birth order"
var_lab(DHS_children_kamp$V149) = "Educational attainment of the mother"
var_lab(DHS_children_kamp$V729) = "Educational attainment of the father"  
var_lab(DHS_children_kamp$V191) = "Wealth index"
var_lab(DHS_children_kamp$V212) = "Age of respondent at first birth"
var_lab(DHS_children_kamp$V445) = "BMI of mother (kg/m^2)"
var_lab(DHS_children_kamp$V714) = "Work status of mother"
var_lab(DHS_children_kamp$B4)   = "Sex of child"
var_lab(DHS_children_kamp$HW5)  = "Stunting (continious)"
var_lab(DHS_children_kamp$HW8)  = "Underweight (continious)"  
var_lab(DHS_children_kamp$HW11) = "Wasting (continious)"
var_lab(DHS_children_kamp$V024) = "Region"
var_lab(DHS_children_kamp$M19)  = "Birth weight in grams"
var_lab(DHS_children_kamp$M5)   = "Duration of breastfeeding"
var_lab(DHS_children_kamp$V113) = "Source of drinking water"
var_lab(DHS_children_kamp$V136) = "Number of household members"
var_lab(DHS_children_kamp$V481) = "Covered by health insurance"
var_lab(DHS_children_kamp$H22)  = "Fever last two weeks"
var_lab(DHS_children_kamp$V505) = "Number of other wives"
var_lab(DHS_children_kamp$V714) = "Mother is currently working"
var_lab(DHS_children_kamp$V729) = "Education of father/partner"
var_lab(DHS_children_kamp$V024) = "Region of residence"
var_lab(DHS_children_kamp$V149) = "Education of mother"
var_lab(DHS_children_kamp$B4)   = "Sex"
var_lab(DHS_children_kamp$H11)  = "Had diarrhea in the last two weeks"
var_lab(DHS_children_kamp$B11)  = "Preceding birth interval <=24"
var_lab(DHS_children_kamp$B0)   = "Child is twin"
var_lab(DHS_children_kamp$H7)  = "Child has received DPT3 vaccination"
```

## Removing invalid data

```{r removing invalid data}
# Removing invalid data
## Filter out rows with children who are no longer alive (B5 == 1)
DHS_children_kamp = subset(DHS_children_kamp, B5 == 1)

## Keep only children with anthropometric measurements (non-missing HW5 values)
DHS_children_kamp = DHS_children_kamp %>% drop_na(HW5)

```

```{r}
summary(DHS_children_kamp$mean_travel_time)

table(DHS_children_kamp$mean_travel_time, DHS_children_kamp$V001)

ggplot(DHS_children_kamp, aes(x = mean_travel_time)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  labs(title = "Mean Travel Time Distribution",
    x = "Mean Travel Time",
    y = "Frequency") +
  theme_minimal()

DHS_children_kamp$travel_time_pf <- DHS_children_kamp$mean_travel_time + runif(nrow(DHS_children_kamp), min = 0, max = 60)

DHS_children_kamp$travel_time_pf

esquisser(DHS_children_kamp)
```

## Exploring Missing Data

```{r exploring missing data}
# Investigating missing data
## Calculate the proportion of missing values for each variable
prop_miss(DHS_children_kamp)

## Count the number of missing values for each variable
n_miss(DHS_children_kamp)

## Summary of missing values by variable
miss_var_summary(DHS_children_kamp)

## Summary of missing values by case
miss_case_summary(DHS_children_kamp)

## Create a table showing patterns of missing data
miss_case_table(DHS_children_kamp)

## Visualize missing data
vis_miss(DHS_children_kamp) + theme(axis.text.x = element_text(angle=80))

gg_miss_upset(DHS_children_kamp)

## Define dependent and explanatory variables for missing data analysis
dependent1 = "HW5"
dependent2 = "HW8"
dependent3 = "HW11"
explanatory = c("V149", "V729", "V191", "V212", "V445", "V714", "B4", "BORD", "M19", "B11", "H11", "V113", "V136", "H7", "V481", "H22", "V505")

## Create missing data plots for dependent1, dependent2, and dependent3
missing_plot(DHS_children_kamp, dependent1, explanatory)
missing_plot(DHS_children_kamp, dependent2, explanatory)
missing_plot(DHS_children_kamp, dependent3, explanatory)

## Generate summary tables for missing data
missing1 = ff_glimpse(DHS_children_kamp, dependent1, explanatory)
flextable(missing1$Continuous) #%>%
  save_as_image("missing_cont.png")

ff_glimpse(DHS_children_kamp, dependent2, explanatory)
ff_glimpse(DHS_children_kamp, dependent3, explanatory)

## Create missing data pair plots for dependent1, dependent2, and dependent3
DHS_children_kamp %>%
  missing_pairs(dependent1, explanatory, position = "fill")

DHS_children_kamp %>%
  missing_pairs(dependent2, explanatory, position = "fill")

DHS_children_kamp %>%
  missing_pairs(dependent3, explanatory, position = "fill")
```

# Descriptive Analysis

```{r}
# Summarize key variables
summary(DHS_data)
summary(GIS_data)

# Initial data exploration and visualization
ggplot(DHS_data, aes(x = variable, fill = factor)) +
  geom_bar() +
  theme_minimal()
# Replace 'variable' and 'factor' with relevant column names
```

# Bivariate Analysis

```{r}
# Assess relationships between variables
cor.test(DHS_data$variable1, DHS_data$variable2)
# Replace 'variable1' and 'variable2' with relevant column names

# Compare mobility levels and undernutrition
# Add your code here
```

# Multinomial Regression

```{r}
# Fit a multinomial regression model
multinom_model <- multinom(response ~ predictor1 + predictor2, data = DHS_data)
summary(multinom_model)
# Replace 'response', 'predictor1', 'predictor2' with relevant column names
```

# Spatial Analysis

```{r}
# Analyze spatial data
# Plot flooding extents and healthcare facilities
ggplot() +
  geom_sf(data = GIS_data) +
  theme_minimal()
# Add other spatial analysis code
```

# Equity Assessment

```{r}
# Analyze equity implications of findings
# Add code for equity assessment, possibly involving comparisons of different groups
```

