# InAccessmod.qmd Documentation

## Overview
The `InAccessmod.qmd` file contains the data preparation code for the Accessmod tool.

## Purpose
The purpose of this file is to preprocess the input data required by the Accessmod tool. It performs various data cleaning, transformation, and aggregation tasks to ensure that the data is in the appropriate format for analysis.

## Usage
To use this file, follow these steps:
1. Ensure that the required input data files are available.
2. Open the `InAccessmod.qmd` file in a compatible code editor or environment.
3. Review the code and make any necessary modifications based on your specific data requirements.
4. Execute the code to preprocess the data.

## Dependencies
This file has dependencies on the `inAccessMod` package. Make sure to install the package using the `devtools::install_github` function before running the code.

## Output
The output of this file is the preprocessed data that can be used as input for the Accessmod tool.

## Notes
- It is recommended to review the code and documentation of the Accessmod tool to understand the expected input data format and any additional preprocessing steps that may be required.

## References
- [Accessmod Tool Documentation](https://www.accessmod.org/)

# Install the inAccessMod package from GitHub & load the library
```{r}
# Set up renv for reproducible package management
if (!requireNamespace("renv", quietly = TRUE)) {
    install.packages("renv")
} else {
    message("renv is already installed.")
}

# Load the renv package for environment management
library(renv)

# Activate renv if the project environment is not synchronized
if (!renv::status()$synchronized) {
    renv::activate()
}

# Restore the package environment (useful for collaborators)
renv::restore()

# Load required libraries for data processing and AccessMod preparation
library(inAccessMod)
library(terra)
library(here)
library(Spatial.align)
```

# Initialize the project for Uganda
```{r}
# Define the main project directory
# All input and output files for AccessMod will be organized under this root folder
mainPath <- here("AccessMod/InAccessMod")

# Specify the country for analysis
# Used throughout the script to identify country-specific data and processing steps
country <- "Uganda"
```

# Download or process the required input data for AccessMod
## Administrative boundaries
```{r}
# Load parish administrative boundaries
# This step copies the processed parish boundaries to the AccessMod input folder for Uganda.
inputPath_adm <- here("Data/processed/Adm_boundaries")
copy_input(mainPath, country, inputPath_adm)
```

## Population data
```{r}
# Total population
inputPath_pop <- here("Data/processed/Worldpop_density/worldpop_density_kampala.tif")
copy_input(mainPath, country, inputPath_pop)
```

## Digital Elevation Model (DEM)
```{r}
inputPath_DEM <- here("Data/processed/DEM/dem30m_kampala.tif")
copy_input(mainPath, country, inputPath_DEM)
```

## Land cover data
```{r}
# Load land cover raster data
# This step copies the processed land cover raster to the AccessMod input folder for Uganda.
inputPath_LC <- here("Data/processed/Landcover/landcover10m_kampala.tif")
copy_input(mainPath, country, inputPath_LC)
```

## Health facilities data
```{r}
# Load health facilities data (all facilities and hospitals only)
# This step copies the processed health facilities shapefile to the AccessMod input folder for Uganda.
inputPath_HF <- here("Data/processed/Facilities")
copy_input(mainPath, country, inputPath_HF)
```

## OpenStreetMap (OSM) data
```{r}
# Load OpenStreetMap (OSM) roads data
# This step copies the processed OSM roads shapefile to the AccessMod input folder for Uganda.
inputPath_roads <- here("Data/processed/OSM_roads")
copy_input(mainPath, country, inputPath_roads)
```

# Flood data
```{r}
# Load flood extent data
# This step copies the processed flood extent shapefile to the AccessMod input folder for Uganda.
inputPath_flood <- here("Data/processed/Flood_extents/Flood_extents_full")
copy_input(mainPath, country, inputPath_flood)
```

# Water bodies data
```{r}
# Load lakes and rivers vector data
# This step copies the processed lakes and rivers shapefile to the AccessMod input folder for Uganda.
inputPath_water <- here("Data/processed/Water_lines")
copy_input(mainPath, country, inputPath_water)
```

# Set the projected coordinate reference system
```{r}
# Set the projected coordinate reference system (CRS) for all vector data inputs
# This ensures all vector layers are aligned to the optimal CRS for analysis in AccessMod.
set_projection(mainPath, country, mostRecent = TRUE, alwaysSet = TRUE, bestCRS = TRUE)
```

# Process the input data for AccessMod This cannot be done for rasters using the `process_inputs` function as we need to do a manual disaggregation correction for the population data (see 02.pop_raster_correction.qmd.). 
```{r}
# Process vector data inputs (administrative boundaries, health facilities, OSM roads, water bodies)
# Rasters are handled separately and should not be processed with this function.
process_inputs(mainPath, country, mostRecent = TRUE, alwaysProcess = FALSE)

# Manually process raster inputs (population, DEM, landcover)
# Population raster processing is detailed in 02.pop_raster_correction.qmd

dem <- rast(here("Data/processed/DEM/dem30m_kampala.tif"))
landcover <- rast(here("Data/processed/Landcover/landcover10m_kampala.tif"))

# Align DEM and landcover rasters to the population raster:
# - Set CRS to match population raster
# - Crop to population raster extent
# - Resample to population raster resolution
crs(dem) <- crs(pop_total_10m)
dem <- crop(dem, pop_total_10m)
dem <- resample(dem, pop_total_10m, method = "bilinear")

crs(landcover) <- crs(pop_total_10m)
landcover <- crop(landcover, pop_total_10m)
landcover <- resample(landcover, pop_total_10m, method = "near")
```

# Compile all processed data for easier import into AccessMod
```{r}
# Compile processed vector data using the compile_processed_data function
# This step organizes all processed vector inputs into the AccessMod folder structure for Uganda.
compile_processed_data(mainPath, country, mostRecent = FALSE)

# Save processed raster data manually to the zToAccessMod folder
# DEM and landcover rasters are written to the AccessMod input directory for Uganda.
writeRaster(dem, here("AccessMod/InAccessMod/Uganda/data/zToAccessMod/20250108115609/UG_DEM_kamp.tif"), overwrite = TRUE)
writeRaster(landcover, here("AccessMod/InAccessMod/Uganda/data/zToAccessMod/20250108115609/UG_Landcover_kamp.tif"), overwrite = TRUE)
```
