# Block Averaging of Population Raster Data

## Overview

This QMD file provides a script for performing block averaging of population raster data using R. Block averaging involves aggregating high-resolution population data into larger blocks to create a lower-resolution representation. This process can be useful for downsampling data or preparing it for analysis at a coarser scale.


# Load libraries
```{r}
# Set up renv for reproducible package management
if (!requireNamespace("renv", quietly = TRUE)) {
    install.packages("renv")
} else {
    message("renv is already installed.")
}

# Load renv for environment management
library(renv)

# Activate renv if the environment is not synchronized
if (!renv::status()$synchronized) {
    renv::activate()
}

# Restore the package environment (recommended for collaborators)
renv::restore()

# Load required libraries for raster processing
library(terra)
```

# Load data
```{r}
# Load the population raster file (WorldPop density for Kampala)
pop_raster_total <- rast(here("Data/processed/Worldpop_density/worldpop_density_kampala.tif"))

# Load the study area shapefile (parish boundaries for Kampala)
study_area <- vect(here("Data/processed/Adm_boundaries/parishes_kampala.shp"))
```

# Transform the projection of the population raster
```{r}
# Reproject the study area shapefile to match the CRS of the population raster for consistent spatial analysis
study_area <- project(study_area, crs(pop_raster_total))
```

# disaggregate the data to 10x10m resolution
```{r}
# Disaggregate population raster to finer resolution (10x10 times more cells)
pop_total_10m <- disagg(pop_raster_total, c(10, 10))

# Adjust population values so total population is preserved after disaggregation
# Divide each original cell's value by the number of new cells created (10 x 10 = 100)
pop_total_10m <- pop_total_10m / 100
```

# Save the data to new TIFF files
```{r}
# Save the disaggregated population raster to the AccessMod input folder for further analysis
writeRaster(pop_total_10m, here("AccessMod/InAccessMod/Uganda/data/zToAccessMod/20250108115609/pop_total_10m.tif"), overwrite = TRUE)

# Save the disaggregated population raster to the processed WorldPop density folder for record-keeping and reproducibility
writeRaster(pop_total_10m, here("Data/processed/Worldpop_density/pop_total_10m.tif"), overwrite = TRUE)
```
