# Merged Landcovers

This code is used to create merged landcover layers for each flood extent in the AccessMod analysis. It takes flood extent shapefiles and merges them with a base landcover raster to create new landcover layers.

## Step 1: Seperate the flood extents

The code first separates the flood extents based on the unique values of rainfall and duration. It loads the full flood extent data and defines the unique values for threshold, rainfall, and duration. Then, it iterates through each combination of threshold, rainfall, and duration and filters the flood extent data based on the current combination. If the filtered data is not empty, it creates a sub-folder for the flood extent and writes the filtered data to a shapefile in the sub-folder.

## Step 2: Load and name the flood extents

The code then loads all the flood extents from the sub-folders and reads them into a list. It names the flood extents in the list based on their sub-folder name. It also removes the other variable names from the flood extents.

## Step 3: Merge landcover layers for each flood extent

The code creates merged landcover layers for each flood extent. It initializes an empty list to store the merged landcover layers. It loads a base landcover raster and sets its projection. Then, it iterates through each flood extent in the flood list and performs the following steps:

1. Loads the landcover raster and creates a grid with the same extent and resolution.
2. Resamples the landcover raster to the grid resolution.
3. Transforms the flood extent polygons to the same projection as the landcover raster.
4. Rasterizes the flood extent polygons using the grid as a template.
5. Sets the background value of the rasterized flood extent to NA.
6. Merges the rasterized flood extent with the landcover raster.
7. Writes the merged landcover layer to a file.

The resulting merged landcover layers are saved in the "Merged landcovers" folder.

# R code
## Load libraries
```{r}
# Set up renv for reproducible package management
if (!requireNamespace("renv", quietly = TRUE)) {
    install.packages("renv")
} else {
    message("renv is already installed.")
}

# Load renv for environment management
library(renv)

# Activate renv if not already synchronized
if (!renv::status()$synchronized) {
    renv::activate()
}

# Restore package environment (for collaborators)
renv::restore()

# Load required libraries for spatial data processing, plotting, and utilities
library(sf)            # Simple features for vector data
library(dplyr)         # Data manipulation
library(foreach)       # Parallel processing
library(snow)          # Parallel computing support
library(terra)         # Raster and vector spatial data
library(raster)        # Legacy raster support
library(ggplot2)       # Plotting
library(gridExtra)     # Grid graphics
library(tidyterra)     # Tidyverse support for terra objects
library(cowplot)       # Plot composition
library(patchwork)     # Plot composition
library(Spatial.align) # Spatial alignment tools
library(ggspatial)     # Spatial plotting enhancements
library(here)          # Project directory management
library(stringr)       # String manipulation
library(tidyr)         # Data tidying
library(scales)        # Scale functions for visualization
```

# Create merged landcover layer for AccessMod
## First seperate the flood extents based on the unique values of threshold, rainfall, and duration
```{r}
# Separate flood extents by rainfall (mm) and duration (hours), and save each as a shapefile in a dedicated subfolder.
# Load the full flood extent dataset
flood_full <- st_read(here("AccessMod/InAccessMod/Uganda/data/vNaturalPolygons/20250107104606/processed/20250107121314"))

# Define rainfall and duration values to iterate over (duration in hours)
rainfall_values <- c(20, 40, 60, 80, 100)
duration_values <- c(1, 3, 6)

# Loop through each rainfall-duration combination and export corresponding flood polygons
for (rainfall in rainfall_values) {
    for (duration in duration_values) {
        # Filter polygons for the current rainfall-duration scenario
        filtered_data <- flood_full[flood_full$rainfall == rainfall & flood_full$duration == duration, ]
        
        # If any polygons exist for this scenario, save them as a shapefile
        if (nrow(filtered_data) > 0) {
            # Construct subfolder name for this scenario
            subfolder_name <- paste0("flood_", rainfall, "_", duration)
            
            # Path to the scenario subfolder within Flood_extents
            dir_path <- here("Data", "raw", "Flood_extents", subfolder_name)
            
            # Create the subfolder if it doesn't already exist
            if (!dir.exists(dir_path)) {
                dir.create(dir_path, recursive = TRUE)
            }
            
            # Name for the output shapefile
            shapefile_name <- paste0(subfolder_name, ".shp")
            
            # Write the filtered polygons to the shapefile in the scenario subfolder
            st_write(filtered_data, file.path(dir_path, shapefile_name), append = FALSE)
        }
    }
}

# Load all scenario flood extent shapefiles from subfolders into a list
flood_list <- list.files(path = here("Data", "raw", "Flood_extents"), pattern = "\\.shp$", full.names = TRUE, recursive = TRUE) %>%
    lapply(st_read)

# Extract scenario names from subfolder names for each flood extent
flood_extents_names <- basename(dirname(list.files(path = here("Data", "raw", "Flood_extents"), pattern = "\\.shp$", full.names = TRUE, recursive = TRUE)))

# Add a Scenario column to each flood extent object, based on its subfolder name
for (i in 1:length(flood_list)) {
    flood_list[[i]]$Scenario <- gsub("flood_", "", flood_extents_names[i])
}

# Assign scenario names as list element names for easy reference
names(flood_list) <- flood_extents_names

# Remove rainfall and duration columns from each flood extent object, retaining only geometry and scenario info
flood_list <- lapply(flood_list, function(x) {
    x$rainfall <- NULL
    x$duration <- NULL
    return(x)
})

```

## Merge landcover layers for each flood extent
```{r}
# Create a list to store merged landcover rasters for each flood scenario
merged_landcovers <- list()

# Load the base merged landcover raster (MLC0)
MLC0 <- rast(here("Data/processed/Merged_landcovers/raster_land_cover_merged_MLC0.img"))

# Load landcover raster to set projection (ensures CRS consistency)
landcover <- rast(here("AccessMod/InAccessMod/Uganda/data/zToAccessMod/20250108115609/UG_Landcover_kamp.tif"))

# Set the CRS of the base raster to match the landcover raster
crs(MLC0) <- crs(landcover)

## For each flood scenario, merge the flood extent polygons with the base landcover raster
for (i in 1:length(flood_list)) {
    landcov <- MLC0 # Use base landcover raster for merging
    flood_extent <- flood_list[[i]] %>% # Get flood polygons for current scenario
        mutate(class = case_when(
            threshold == 0.1 ~ 1, # Assign flood depth class 1 for 10cm
            threshold == 0.2 ~ 2, # Assign flood depth class 2 for 20cm
            threshold == 0.3 ~ 3, # Assign flood depth class 3 for 30cm
            TRUE ~ NA_real_
        )) %>% 
        as(., "Spatial") %>% # Convert sf object to sp object
        vect() # Convert sp object to terra vector

    # Rasterize flood extent polygons to match the base raster grid, using the assigned class
    flood_extent_rast <- terra::rasterize(flood_extent, landcov, "class", touches = TRUE)

    # Mask rasterized flood extent to preserve NA values from the base raster (avoids filling gaps)
    flood_extent_rast <- mask(flood_extent_rast, landcov, inverse = FALSE) 

    # Merge the flood raster with the base landcover raster, prioritizing flood classes where present
    landcover_merge <- terra::cover(flood_extent_rast, landcov) 

    # Save the merged raster for the current scenario to disk
    writeRaster(landcover_merge, here("Data", "processed", "Merged_landcovers", paste0("MLC_", flood_extents_names[i], ".tif")), overwrite = TRUE)
}
```

# Prepare merged landcover layers for plotting
```{r}	
# Load all merged landcover rasters from the "Merged landcovers" folder
merged_landcovers <- list.files(path = here("Data", "processed", "Merged_landcovers"), pattern = "\\.tif$", full.names = TRUE, recursive = TRUE) %>%
    lapply(terra::rast)

# Extract file names for merged landcover rasters (used for naming and ordering)
merged_landcovers_names <- list.files(path = here("Data", "processed", "Merged_landcovers"), pattern = "\\.tif$", full.names = FALSE, recursive = TRUE)

# Order merged landcover rasters by flood depth (mm) and duration (hr) extracted from file names
# Extract flood depths and durations from file names using regular expressions
flood_depths <- as.numeric(sub(".*_flood_([0-9]+)_.*", "\\1", merged_landcovers_names))
durations <- as.numeric(sub(".*_([0-9]+)\\.tif$", "\\1", merged_landcovers_names))

# Assign Inf to NA values for files that do not match the expected pattern (e.g., summary rasters)
flood_depths[is.na(flood_depths)] <- Inf
durations[is.na(durations)] <- Inf

# Reorder file names by flood depth and duration
merged_landcovers_names <- merged_landcovers_names[order(flood_depths, durations, merged_landcovers_names)]

# Remove summary rasters and NA entries from the list (e.g., "MLC_Flood_extents_full.tif")
merged_landcovers_names <- merged_landcovers_names[!grepl("full", merged_landcovers_names) & !is.na(merged_landcovers_names)]

# Reorder the merged_landcovers list to match the ordered file names
merged_landcovers <- merged_landcovers[order(flood_depths, durations)]

# Assign each merged landcover raster to a named object in the environment for convenience
for (i in 1:length(merged_landcovers)) {
    assign(gsub(".tif", "", merged_landcovers_names[i]), merged_landcovers[[i]])
}

# Define legend colors and names for landcover and flood classes
legend_colour <- data.frame(
    value = c(1, 2, 3, 10, 20, 30, 40, 50, 60, 80, 90, 1001, 1002, 1003, 1004, 1005),
    color = c("#169FFB", "#3C3CCF", "#170AA6", "#A8E16E", "#FFB400", "#FFFF1B", "#D7B5D8", "#FF0000", "#30123B", "#0312DC", "#15C44C", "#DE0004", "#CE0FA4", "#336CE6", "#34EF05", "#333333")
)

legend_names <- c(
    "Flood 10cm",
    "Flood 20cm",
    "Flood 30cm",
    "Tree Cover",
    "Shrubs",
    "Grass Land",
    "Agriculture",
    "Urban",
    "Bare",
    "Water Bodies",
    "Herbaceous wetland",
    "Trunk",
    "Primary",
    "Motorway",
    "Secondary",
    "Tertiary"
)

# Create a data frame mapping raster values to legend categories
levels_df <- data.frame(
  ID = legend_colour$value,   # Raster values
  Category = legend_names     # Corresponding category names
)

# Assign legend levels to each raster for proper labeling in plots
merged_landcovers <- lapply(merged_landcovers, function(r) {
    levels(r) <- levels_df
    r
})

# Convert each raster to a factor for categorical plotting
merged_landcovers <- lapply(merged_landcovers, function(r) {
    as.factor(r)
})

# Assign descriptive names to each merged landcover raster based on flood scenario
names(merged_landcovers) <- c("20 mm / 1hr", "20 mm / 3hr", "20 mm / 6hr", "40 mm / 1hr", "40 mm / 3hr", "40 mm / 6hr", "60 mm / 1hr", "60 mm / 3hr", "60 mm / 6hr", "80 mm / 1hr", "80 mm / 3hr", "80 mm / 6hr", "100 mm / 1hr", "100 mm / 3hr", "100 mm / 6hr")

# Create a named vector for color mapping in plots
color_mapping <- setNames(legend_colour$color, legend_colour$value)
```

# Plot
```{r}
# -----------------------------
# 1) Parse scenario names, order layers, and build a 15-layer SpatRaster stack
# -----------------------------
# Assumes: merged_landcovers is a named list of SpatRaster layers, each representing a flood scenario
meta <- tibble(name = names(merged_landcovers)) |>
    mutate(
        mm = as.integer(str_extract(name, "(?i)(?<=^|[^0-9])\\d+(?=\\s*mm)")),
        hr = as.integer(str_extract(name, "(?i)(?<=^|[^0-9])\\d+(?=\\s*h)")),
        lab = sprintf("%d mm / %d hr", mm, hr)
    ) |>
    filter(!is.na(mm), !is.na(hr)) |>
    arrange(mm, factor(hr, levels = c(1, 3, 6)))

ord <- meta$name

# Stack layers in the desired order for plotting
mlc_stack <- rast(merged_landcovers[ord])

# Assign tidy facet labels to each layer
names(mlc_stack) <- meta$lab

# -----------------------------
# 2) Define legend, color palette, and alpha mapping for categorical raster values
# -----------------------------
legend_colour <- data.frame(
    value = c(1, 2, 3, 10, 20, 30, 40, 50, 60, 80, 90, 1001, 1002, 1003, 1004, 1005),
    color = c("#169FFB", "#3C3CCF", "#170AA6", "#A8E16E", "#FFB400", "#FFFF1B",
                        "#D7B5D8", "#FF0000", "#30123B", "#0312DC", "#15C44C",
                        "#DE0004", "#CE0FA4", "#336CE6", "#34EF05", "#333333")
)

legend_names <- c(
    "Flood 10cm", "Flood 20cm", "Flood 30cm",
    "Tree Cover", "Shrubs", "Grass Land", "Agriculture", "Urban", "Bare",
    "Water Bodies", "Herbaceous wetland",
    "Trunk", "Primary", "Motorway", "Secondary", "Tertiary"
)

# Create named vectors for ggplot2 manual scales
pal   <- setNames(legend_colour$color, as.character(legend_colour$value))
labs  <- setNames(legend_names,            as.character(legend_colour$value))

# Set alpha: fully opaque for flood classes, semi-transparent for other landcover types
alpha_map <- setNames(
    ifelse(as.integer(names(pal)) %in% c(1, 2, 3), 1, 0.6),
    names(pal)
)

# To restrict legend to only present classes, uncomment below:
# present_vals <- as.character(sort(unique(stats::na.omit(values(mlc_stack[])))))
# pal   <- pal[names(pal) %in% present_vals]
# labs  <- labs[names(labs) %in% present_vals]
# alpha_map <- alpha_map[names(alpha_map) %in% present_vals]

# -----------------------------
# 3) Convert overlay data to sf format if needed
# -----------------------------
# OSM_roads and water_lines may be SpatVector or sf; convert to sf for plotting
if (inherits(OSM_roads, "SpatVector"))  OSM_roads  <- sf::st_as_sf(OSM_roads)
if (inherits(water_lines, "SpatVector")) water_lines <- sf::st_as_sf(water_lines)

# -----------------------------
# 4) Build faceted panel plot of merged landcover and flood scenarios
# -----------------------------
# Use geom_spatraster to plot the 15-layer SpatRaster, faceted by scenario
# Fill and alpha are mapped to categorical raster values
gridplot_mlc <-
        ggplot() +
        geom_spatraster(
                data = mlc_stack,
                aes(fill = as.factor(after_stat(value)), alpha = as.factor(after_stat(value)))
        ) +
        facet_wrap(~lyr, nrow = 5, ncol = 3) +
        # Overlay OSM roads and water lines with thin lines
        geom_sf(data = OSM_roads,  color = "black", linewidth = 0.15, inherit.aes = FALSE) +
        geom_sf(data = water_lines, color = "blue",  linewidth = 0.12, linetype = 2, inherit.aes = FALSE) +
        # Manual fill and alpha scales for categorical legend
        scale_fill_manual(
                values = pal,
                breaks = names(pal),
                labels = labs[names(pal)],
                name   = "Merged land-cover, barriers and floods",
                na.translate = FALSE,
                na.value = "transparent",
                guide = guide_legend(
                        ncol = 2,
                        title.position = "top",
                        # Set legend key alpha to match facet alpha by class
                        override.aes = list(alpha = unname(alpha_map[names(pal)]))
                )
        ) +
        scale_alpha_manual(values = alpha_map, guide = "none") +
        coord_sf(expand = FALSE) +
        theme_minimal(base_size = 8) +  # Set base font size to 8
        theme(
                plot.background = element_rect(fill = "white", color = NA),
                panel.background = element_rect(fill = "white", color = NA),
                axis.text = element_blank(),
                axis.title = element_blank(),
                panel.grid = element_blank(),
                strip.text = element_text(size = 8, face = "plain"),
                strip.background = element_rect(fill = "white", colour = NA),
                legend.position = "right",
                legend.title = element_text(size = 8, face = "bold"),
                legend.text  = element_text(size = 8),
                panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6)
        )

# Save the faceted plot as a PDF for publication
ggsave(
        filename = here("Figures", "Pdf_final", "app_fig_1_merged_landcovers_w_floods.pdf"),
        plot = gridplot_mlc,
        width = 15.92, height = 24.62, units = "cm"
)
```
