# Population impacted weighted delta travel time calculation and visualization
```{r}
# Set up renv for reproducible package management
if (!requireNamespace("renv", quietly = TRUE)) {
    install.packages("renv")
} else {
    message("renv is already installed.")
}

# Load the renv package
library(renv)

# Activate renv if not already synchronized
if (!renv::status()$synchronized) {
    renv::activate()
}

# Restore the environment (useful for collaborators to install required packages)
renv::restore()

# Load all required libraries for analysis and visualization
library(terra)           # Raster and vector spatial data
library(Spatial.align)   # Raster alignment
library(patchwork)       # Plot composition
library(sf)              # Simple features for vector data
library(biscale)         # Bivariate mapping
library(classInt)        # Classification intervals
library(here)            # Relative file paths
library(ggplot2)         # Plotting
library(tidyterra)       # Tidyverse support for terra objects
library(tidyverse)       # Data manipulation and visualization
library(flextable)       # Table formatting for Word export
library(officer)         # Word document creation
library(kableExtra)      # Enhanced HTML table formatting
```

# Weighted Delta Travel Time Calculation for scenarios with All Facilities
```{r}
# Load delta travel time rasters for all facilities
## For delta travel time go to script 02.1_maps_tt_and_delta_TT
### After running that script, create a list using the following code:

delta_tt_list_all <- mget(ls(pattern = "delta_tt_all"))

# Remove any non-raster objects from delta_tt_list_all (e.g., "delta_tt_all_weighted")
delta_tt_list_all <- delta_tt_list_all[grepl("^delta_tt_all_\\d+_\\d+$", names(delta_tt_list_all))]

# Load population density raster
pop_density_raster <- rast(here("AccessMod", "InAccessMod", "Uganda", "data", "zToAccessMod", "20250108115609", "pop_total_10m.tif"))

# Load parish boundaries as SpatVector
parish_boundaries_all <- vect(here("Data", "processed", "Adm_boundaries", "parishes_kampala.shp"))

# Step 1: Multiply each delta travel time raster by the population density raster to get population-weighted delta TT
weighted_delta_tt_list_all <- lapply(delta_tt_list_all, function(delta_tt_raster) {
    delta_tt_raster * pop_density_raster
})

# Rename weighted delta TT rasters for clarity (keep only scenario info)
names(weighted_delta_tt_list_all) <- sapply(names(weighted_delta_tt_list_all), function(name) {
    parts <- strsplit(name, "_")[[1]]
    paste(tail(parts, 2), collapse = "_")
})

# Step 2: Save each weighted delta TT raster to disk in a dedicated folder
output_folder <- here("Data", "processed", "delta_travel", "weighted_delta_tt")
if (!dir.exists(output_folder)) {
    dir.create(output_folder, recursive = TRUE)
}

for (i in seq_along(weighted_delta_tt_list_all)) {
    raster_obj <- weighted_delta_tt_list_all[[i]]
    scenario_name <- names(weighted_delta_tt_list_all)[i]
    file_path <- file.path(output_folder, paste0("weighted_delta_tt_", scenario_name, ".tif"))
    writeRaster(raster_obj, file_path, overwrite = TRUE)
}

# Step 3: Calculate zonal sum of weighted delta TT for each parish (all facilities)
for (i in seq_along(weighted_delta_tt_list_all)) {
    sum_weighted_delta_tt <- zonal(weighted_delta_tt_list_all[[i]], parish_boundaries_all, fun = "sum", na.rm = TRUE)
    original_name <- names(weighted_delta_tt_list_all)[i]
    parish_boundaries_all[[paste0("Weighted_delta_tt_", original_name)]] <- sum_weighted_delta_tt[, 1]
}

# Step 4: Calculate total population per parish
parish_boundaries_all$Population_Sum <- zonal(pop_density_raster, parish_boundaries_all, fun = "sum", na.rm = TRUE)

# Step 5: Calculate average population-weighted delta TT per parish for each scenario
for (i in seq_along(weighted_delta_tt_list_all)) {
    original_name <- names(weighted_delta_tt_list_all)[i]
    parish_boundaries_all[[paste0("Average_Weighted_Delta_TT_", original_name)]] <-
        parish_boundaries_all[[paste0("Weighted_delta_tt_", original_name)]] / parish_boundaries_all$Population_Sum
}
```

# Create a bivariate map of accessibility reduction and population sum for all facilities
```{r}	
# Bivariate map of population-weighted delta travel time and population sum for ALL FACILITIES
# STEP 1: Convert parish boundaries to sf object for plotting
parish_boundaries_all_sf <- st_as_sf(parish_boundaries_all)

# STEP 2: Ensure all delta TT values are positive for visualization
for (col in names(parish_boundaries_all_sf)[grepl("Average_Weighted_Delta_TT_", names(parish_boundaries_all_sf))]) {
    parish_boundaries_all_sf[[col]] <- abs(parish_boundaries_all_sf[[col]])
}

# STEP 3: Extract and order scenario columns by rainfall intensity and duration
scenario_columns_all <- grep("Average_Weighted_Delta_TT_", names(parish_boundaries_all_sf), value = TRUE)
scenario_columns_ordered_all <- scenario_columns_all[order(
    as.numeric(sub(".*_(\\d+)_.*", "\\1", scenario_columns_all)),
    as.numeric(sub(".*_(\\d+)$", "\\1", sub(".*_\\d+_", "", scenario_columns_all)))
)]

# STEP 4: Bin population sum into categories for bivariate mapping
parish_boundaries_all_sf$Population_bin <- factor(cut(
    parish_boundaries_all_sf$Population_Sum,
    breaks = c(0, 10000, 30000, Inf),
    include.lowest = TRUE,
    labels = c("Low", "Medium", "High")
), levels = c("High", "Medium", "Low"))

# STEP 5: Define legend labels for bivariate map
labels_all <- list(
    bi_x = c("0-30", "30-60", "60+"),
    bi_y = c("30.000+", "10.000 - 30.000", "0 - 10.000")
)

# STEP 6: Create bivariate map plots for each scenario
plot_list_all <- list()

for (i in seq_along(scenario_columns_ordered_all)) {
    scenario_col <- scenario_columns_ordered_all[i]

    # Extract rainfall intensity (mm) and duration (hours) for plot title
    numbers <- sub(".*_(\\d+)_(\\d+)$", "\\1 \\2", scenario_col)
    mm <- unlist(strsplit(numbers, " "))[1]
    hours <- as.numeric(unlist(strsplit(numbers, " "))[2]) / 3600
    title_text <- paste(mm, "mm x", hours, "hour")

    # Bin delta TT values for bivariate mapping
    parish_boundaries_all_sf$Weighted_bin <- factor(cut(
        parish_boundaries_all_sf[[scenario_col]],
        breaks = c(0, 30, 60, Inf),
        include.lowest = TRUE,
        labels = c("0-30", "30-60", "60+")
    ), levels = c("0-30", "30-60", "60+"))

    # Bivariate classification using biscale
    parish_boundaries_bi_all <- bi_class(
        parish_boundaries_all_sf,
        x = "Weighted_bin",
        y = "Population_bin",
        dim = 3
    )

    # Plot bivariate map for this scenario
    p <- ggplot() +
        geom_sf(data = parish_boundaries_bi_all, aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
        bi_scale_fill(pal = "DkViolet", dim = 3, flip_axes = FALSE) +
        labs(title = title_text) +
        bi_theme() +
        theme(
            plot.title = element_text(size = 8, face = "bold", hjust = 0.5),
            legend.position = "none"
        )

    plot_list_all[[i]] <- p
}

# STEP 7: Create shared legend for bivariate map
legend_all <- bi_legend(
    pal = "DkViolet",
    dim = 3,
    xlab = "Δ Travel Time (min)",
    ylab = "Population Sum",
    size = 8,
    breaks = labels_all,
    flip_axes = FALSE,
    arrows = FALSE
)

# STEP 8: Combine all scenario plots into a grid (without legend)
weighted_delta_TT_all <- wrap_plots(plot_list_all, ncol = 3) +
    plot_layout(guides = "collect") &
    theme(plot.title = element_text(hjust = 0.5, size = 8, face = "bold"))

# STEP 9: Save the main bivariate map grid and legend as PDF files
ggsave(
    filename = here("Figures", "Pdf_Final", "Figure_3_weighted_delta_TT_all.pdf"),
    plot = weighted_delta_TT_all,
    width = 15.92,
    height = 24.62,
    units = "cm"
)

ggsave(
    filename = here("Figures", "Pdf_Final", "Figure_3_weighted_delta_TT_all_legend.pdf"),
    plot = legend_all,
    width = 6,
    height = 6,
    units = "cm"
)
```

# Highlight parishes with the highest decrease in travel time for each scenario for all facilities
```{r}
# Format scenario titles for display (e.g., "100_21600" → "100 mm x 6 hours")
format_scenario_title <- function(scenario) {
    numbers <- sub(".*_(\\d+)_(\\d+)$", "\\1 \\2", scenario)
    numbers_split <- unlist(strsplit(numbers, " "))
    hours <- as.numeric(numbers_split[2]) / 3600
    paste(numbers_split[1], "mm x", hours, "hours")
}

# Define scenario column names for ALL facilities
scenario_columns_ordered_all <- c(
    "Average_Weighted_Delta_TT_20_3600", "Average_Weighted_Delta_TT_20_10800", "Average_Weighted_Delta_TT_20_21600",
    "Average_Weighted_Delta_TT_40_3600", "Average_Weighted_Delta_TT_40_10800", "Average_Weighted_Delta_TT_40_21600",
    "Average_Weighted_Delta_TT_60_3600", "Average_Weighted_Delta_TT_60_10800", "Average_Weighted_Delta_TT_60_21600",
    "Average_Weighted_Delta_TT_80_3600", "Average_Weighted_Delta_TT_80_10800", "Average_Weighted_Delta_TT_80_21600",
    "Average_Weighted_Delta_TT_100_3600", "Average_Weighted_Delta_TT_100_10800", "Average_Weighted_Delta_TT_100_21600"
)

# Convert parish boundaries to data frame for analysis
parish_boundaries_all_df <- as.data.frame(parish_boundaries_all)

# Identify top 10 parishes with highest travel time increase for each scenario
top_parishes_list_all <- list()

for (scenario_col in scenario_columns_ordered_all) {
    if (scenario_col %in% names(parish_boundaries_all_df)) {
        travel_times <- parish_boundaries_all_df[[scenario_col]]
        if (!any(is.na(travel_times))) {
            top_parishes <- parish_boundaries_all_df[order(-travel_times), ][1:10, ]
            top_parishes_list_all[[scenario_col]] <- data.frame(
                Parish_Name = top_parishes$PName2016,
                Travel_Time_Increase = top_parishes[[scenario_col]],
                Scenario = scenario_col
            )
        } else {
            warning(paste("Scenario column", scenario_col, "contains NAs. Skipping."))
        }
    } else {
        warning(paste("Scenario column", scenario_col, "not found in all dataset."))
    }
}



# Calculate frequency of parishes appearing in top 10 lists across all scenarios
parish_frequency_all <- table(unlist(lapply(top_parishes_list_all, function(df) df$Parish_Name)))
parish_frequency_all_df <- as.data.frame(parish_frequency_all)
colnames(parish_frequency_all_df) <- c("Parish_Name", "Frequency")
parish_frequency_all_df$Frequency <- as.numeric(parish_frequency_all_df$Frequency)
parish_frequency_all_df <- parish_frequency_all_df[order(-parish_frequency_all_df$Frequency), ]
top_10_parish_frequency_all <- head(parish_frequency_all_df, 10)



# Add indicator to SpatVector for top 10 parishes by frequency
parish_boundaries_all$Top_10 <- ifelse(parish_boundaries_all$PName2016 %in% top_10_parish_frequency_all$Parish_Name, "Top 10", "Not in Top 10")

# Calculate centroids for parish polygons
parish_centroids_all <- centroids(parish_boundaries_all)
centroids_coords_all <- geom(parish_centroids_all)
parish_centroids_all_df <- as.data.frame(parish_centroids_all) %>%
    mutate(PName2016 = parish_boundaries_all$PName2016,
           Top_10 = parish_boundaries_all$Top_10) %>%
    cbind(centroids_coords_all[, c("x", "y")])

# Plot map highlighting top 10 parishes by frequency
highlighted_plot_all <- ggplot() +
    geom_spatvector(data = parish_boundaries_all, aes(fill = Top_10)) +
    geom_label(
        data = parish_centroids_all_df %>% filter(Top_10 == "Top 10"),
        aes(label = PName2016, x = x, y = y),
        color = "black", fill = "white", size = 3,
        label.padding = unit(0.2, "lines"), label.size = 0.25,
        hjust = 0.5, vjust = -0.5
    ) +
    scale_fill_manual(values = c("Top 10" = "red", "Not in Top 10" = "gray"), name = "Top 10 Parishes") +
    theme_minimal() +
    ggtitle("Top 10 Parishes with Highest Travel Time Increase (All Facilities)") +
    theme(
        legend.position = "bottom",
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid.major = element_line(color = "gray80", size = 0.5),
        panel.grid.minor = element_line(color = "gray90", size = 0.25),
        strip.background = element_rect(fill = "white")
    )

# Display highlighted map
print(highlighted_plot_all)
```

# Densitiy plots of the average weighted delta travel time for each scenario across all parishes for all facilities 
```{r}
# Define scenario columns for ALL FACILITIES (ordered by rainfall intensity and duration)
scenario_columns_ordered_all <- c(
    "Average_Weighted_Delta_TT_20_3600", "Average_Weighted_Delta_TT_20_10800", "Average_Weighted_Delta_TT_20_21600",
    "Average_Weighted_Delta_TT_40_3600", "Average_Weighted_Delta_TT_40_10800", "Average_Weighted_Delta_TT_40_21600",
    "Average_Weighted_Delta_TT_60_3600", "Average_Weighted_Delta_TT_60_10800", "Average_Weighted_Delta_TT_60_21600",
    "Average_Weighted_Delta_TT_80_3600", "Average_Weighted_Delta_TT_80_10800", "Average_Weighted_Delta_TT_80_21600",
    "Average_Weighted_Delta_TT_100_3600", "Average_Weighted_Delta_TT_100_10800", "Average_Weighted_Delta_TT_100_21600"
)

# Compute global axis limits for density plots across all scenarios
x_limits_all <- range(unlist(parish_boundaries_all_df[scenario_columns_ordered_all]), na.rm = TRUE)
y_limits_all <- c(0, max(sapply(scenario_columns_ordered_all, function(col) {
    max(density(parish_boundaries_all_df[[col]], na.rm = TRUE)$y)
})))

# Create density plots for each scenario (all facilities)
density_plot_list_all <- list()

for (i in seq_along(scenario_columns_ordered_all)) {
    scenario_col <- scenario_columns_ordered_all[i]
    if (scenario_col %in% names(parish_boundaries_all_df)) {
        p <- ggplot(parish_boundaries_all_df, aes(x = .data[[scenario_col]])) + 
            geom_density(fill = "blue", color = "black", alpha = 0.7) +
            xlim(x_limits_all) +
            ylim(y_limits_all) +
            theme_minimal() +
            ggtitle(format_scenario_title(scenario_col)) +
            xlab("Average Weighted Delta Travel Time") +
            ylab("Density") +
            theme(
                plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
                axis.title = element_text(size = 12),
                axis.text = element_text(size = 10)
            )
        density_plot_list_all[[i]] <- p
    }
}

# Combine all scenario density plots into a grid (all facilities)
final_density_plot_all <- wrap_plots(density_plot_list_all, ncol = 3) + 
    plot_layout(guides = "collect") &
    plot_annotation(
        title = "Density Plots of Average Weighted Delta Travel Time Scenarios (All Facilities)",
        theme = theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
        )
    )

print(final_density_plot_all)

# Create overlayed density plots by rainfall duration (1h, 3h, 6h) for all facilities
overlayed_plot_list_all <- list()
hours_list <- c(3600, 10800, 21600)
color_palette <- c("blue", "green", "orange", "purple", "red")

for (hour in hours_list) {
    scenarios_for_hour <- grep(paste0("_", hour), scenario_columns_ordered_all, value = TRUE)
    combined_df <- data.frame()
    
    for (scenario_col in scenarios_for_hour) {
        mm_value <- sub(".*_(\\d+)_\\d+$", "\\1", scenario_col)
        temp_df <- data.frame(
            mm = as.numeric(mm_value),
            travel_times = parish_boundaries_all_df[[scenario_col]]
        )
        combined_df <- rbind(combined_df, temp_df)
    }
    
    p <- ggplot(combined_df, aes(x = travel_times, color = as.factor(mm))) + 
        geom_density(fill = NA, alpha = 0.7) +
        theme_minimal() +
        ggtitle(paste(hour / 3600, "hour(s)")) +
        xlab("Average Weighted Delta Travel Time") +
        ylab("Density") +
        scale_color_manual(values = color_palette, name = "mm Values", breaks = c(20, 40, 60, 80, 100)) +
        theme(
            plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
            axis.title = element_text(size = 12),
            axis.text = element_text(size = 10)
        )
    overlayed_plot_list_all[[as.character(hour)]] <- p
}

# Combine overlayed density plots for all durations (all facilities)
final_overlayed_plot_all <- wrap_plots(overlayed_plot_list_all, ncol = 1) +
    plot_annotation(
        title = "Overlayed Density Plots for Different mm by Duration (All Facilities)",
        theme = theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
        )
    )

print(final_overlayed_plot_all)
```

########################################################
Hospiital only scenarios
########################################################

# Weighted Delta Travel Time Calculation for scenarios with Hospitals only
# Load delta travel time and population distribution
```{r}
# Load hospital-only delta travel time rasters
## For delta travel time go to script 05. maps_tt_and_delta_TT
## After running that script, create a list using the following code:
delta_tt_list_hosp <- mget(ls(pattern = "delta_tt_hosp"))

# Load population density raster
pop_density_raster <- rast(here("AccessMod", "InAccessMod", "Uganda", "data", "zToAccessMod", "20250108115609", "pop_total_10m.tif"))

# Load parish boundaries as SpatVector for hospital-only analysis
parish_boundaries_hosp <- vect(here("Data", "processed", "Adm_boundaries", "parishes_kampala.shp"))

# Remove any non-raster objects from delta_tt_list_hosp (e.g., plots, lists, etc.)
delta_tt_list_hosp <- delta_tt_list_hosp[grepl("^delta_tt_hosp_\\d+_\\d+$", names(delta_tt_list_hosp))]

# Align all hospital-only delta TT rasters to the population raster grid
delta_tt_list_hosp <- lapply(delta_tt_list_hosp, function(x) Spatial.align::align(x, pop_density_raster))

# Multiply each hospital-only delta TT raster by the population raster to get population-weighted delta TT
weighted_delta_tt_list_hosp <- lapply(delta_tt_list_hosp, function(delta_tt_raster) {
    delta_tt_raster * pop_density_raster
})

# Rename weighted delta TT rasters for clarity (keep only scenario info)
names(weighted_delta_tt_list_hosp) <- sapply(names(weighted_delta_tt_list_hosp), function(name) {
    parts <- strsplit(name, "_")[[1]]
    paste(tail(parts, 2), collapse = "_")
})

# Save each weighted delta TT raster to disk in a dedicated folder for hospitals only
output_folder_hosp <- here("Data", "processed", "delta_travel", "weighted_delta_tt_hosp")
if (!dir.exists(output_folder_hosp)) {
    dir.create(output_folder_hosp, recursive = TRUE)
}

for (i in seq_along(weighted_delta_tt_list_hosp)) {
    raster_obj <- weighted_delta_tt_list_hosp[[i]]
    scenario_name <- names(weighted_delta_tt_list_hosp)[i]
    file_path <- file.path(output_folder_hosp, paste0("weighted_delta_tt_hosp_", scenario_name, ".tif"))
    writeRaster(raster_obj, file_path, overwrite = TRUE)
}

# Calculate zonal sum of weighted delta TT for each parish (hospital-only scenarios)
for (i in seq_along(weighted_delta_tt_list_hosp)) {
    sum_weighted_delta_tt <- zonal(weighted_delta_tt_list_hosp[[i]], parish_boundaries_hosp, fun = "sum", na.rm = TRUE)
    original_name <- names(weighted_delta_tt_list_hosp)[i]
    parish_boundaries_hosp[[paste0("Weighted_delta_tt_", original_name)]] <- sum_weighted_delta_tt[, 1]
}

# Calculate total population per parish for hospital-only scenarios
parish_boundaries_hosp$Population_Sum <- zonal(pop_density_raster, parish_boundaries_hosp, fun = "sum", na.rm = TRUE)

# Calculate average population-weighted delta TT per parish for each hospital-only scenario
for (i in seq_along(weighted_delta_tt_list_hosp)) {
    original_name <- names(weighted_delta_tt_list_hosp)[i]
    parish_boundaries_hosp[[paste0("Average_Weighted_Delta_TT_", original_name)]] <-
        parish_boundaries_hosp[[paste0("Weighted_delta_tt_", original_name)]] / parish_boundaries_hosp$Population_Sum
}
```

```{r}
# Create a bivariate map of population-weighted delta travel time and population sum for hospital-only scenarios

# STEP 1: Convert parish boundaries to sf object for plotting
parish_boundaries_hosp_sf <- st_as_sf(parish_boundaries_hosp)

# STEP 2: Ensure all delta TT values are positive for visualization
for (col in names(parish_boundaries_hosp_sf)[grepl("Average_Weighted_Delta_TT_", names(parish_boundaries_hosp_sf))]) {
    parish_boundaries_hosp_sf[[col]] <- abs(parish_boundaries_hosp_sf[[col]])
}

# STEP 3: Extract and order scenario columns by rainfall intensity and duration
scenario_columns_hosp <- grep("Average_Weighted_Delta_TT_", names(parish_boundaries_hosp_sf), value = TRUE)
scenario_columns_ordered_hosp <- scenario_columns_hosp[order(
    as.numeric(sub(".*_(\\d+)_.*", "\\1", scenario_columns_hosp)),
    as.numeric(sub(".*_(\\d+)$", "\\1", sub(".*_\\d+_", "", scenario_columns_hosp)))
)]

# STEP 4: Bin population sum into categories for bivariate mapping
parish_boundaries_hosp_sf$Population_bin <- factor(cut(
    parish_boundaries_hosp_sf$Population_Sum,
    breaks = c(0, 10000, 30000, Inf),
    include.lowest = TRUE,
    labels = c("Low", "Medium", "High")
), levels = c("High", "Medium", "Low"))

# STEP 5: Define legend labels for bivariate map
labels_hosp <- list(
    bi_x = c("0-30", "30-60", "60+"),
    bi_y = c("30.000+", "10.000 - 30.000", "0 - 10.000")
)

# STEP 6: Create bivariate map plots for each hospital-only scenario
plot_list_hosp <- list()

for (i in seq_along(scenario_columns_ordered_hosp)) {
    scenario_col <- scenario_columns_ordered_hosp[i]

    # Extract rainfall intensity (mm) and duration (hours) for plot title
    mm <- sub(".*_(\\d+)_\\d+$", "\\1", scenario_col)
    hr <- as.numeric(sub(".*_(\\d+)$", "\\1", scenario_col)) / 3600
    title_text <- paste(mm, "mm x", hr, "hour")

    # Bin delta TT values for bivariate mapping
    parish_boundaries_hosp_sf$Weighted_bin <- factor(cut(
        parish_boundaries_hosp_sf[[scenario_col]],
        breaks = c(0, 30, 60, Inf),
        include.lowest = TRUE,
        labels = c("0-30", "30-60", "60+")
    ), levels = c("0-30", "30-60", "60+"))

    # Bivariate classification using biscale
    parish_boundaries_bi_hosp <- bi_class(
        parish_boundaries_hosp_sf,
        x = "Weighted_bin",
        y = "Population_bin",
        dim = 3
    )

    # Plot bivariate map for this scenario
    p <- ggplot() +
        geom_sf(data = parish_boundaries_bi_hosp, aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
        bi_scale_fill(pal = "DkViolet", dim = 3, flip_axes = FALSE) +
        labs(title = title_text) +
        bi_theme() +
        theme(
            plot.title = element_text(size = 8, face = "bold", hjust = 0.5),
            legend.position = "none"
        )

    plot_list_hosp[[i]] <- p
}

# STEP 7: Create shared legend for bivariate map
legend_hosp <- bi_legend(
    pal = "DkViolet",
    dim = 3,
    xlab = "Δ Travel Time (min)",
    ylab = "Population Sum",
    size = 8,
    breaks = labels_hosp,
    flip_axes = FALSE,
    arrows = FALSE
)

# STEP 8: Combine all scenario plots into a grid (without legend)
weighted_delta_TT_hosp <- wrap_plots(plot_list_hosp, ncol = 3) +
    plot_layout(guides = "collect") &
    theme(plot.title = element_text(hjust = 0.5, size = 8, face = "bold"))

# STEP 9: Save the main bivariate map grid and legend as PDF files
ggsave(
    filename = here("Figures", "Pdf_Final", "Figure_4_weighted_delta_TT_hosp.pdf"),
    plot = weighted_delta_TT_hosp,
    width = 15.92,
    height = 24.62,
    units = "cm"
)

ggsave(
    filename = here("Figures", "Pdf_Final", "Figure_4_weighted_delta_TT_hosp_legend.pdf"),
    plot = legend_hosp,
    width = 6,
    height = 6,
    units = "cm"
)
```

# Highlight parishes with the highest decrease in travel time for each scenario for all facilities
```{r}


# Convert hospital-only parish boundaries to data frame for analysis
parish_boundaries_hosp_df <- as.data.frame(parish_boundaries_hosp)

# Extract hospital-only scenario columns for average weighted delta travel time
scenario_columns_ordered_hosp <- grep("Average_Weighted_Delta_TT_", names(parish_boundaries_hosp_df), value = TRUE)

# Initialize list to store top 10 parishes for each scenario
top_parishes_list_hosp <- list()

# Identify top 10 parishes with highest travel time increase for each hospital-only scenario
for (i in seq_along(scenario_columns_ordered_hosp)) {
    scenario_col <- scenario_columns_ordered_hosp[i]

    if (scenario_col %in% names(parish_boundaries_hosp_df)) {
        travel_times <- parish_boundaries_hosp_df[[scenario_col]]
        if (!all(is.na(travel_times))) {
            top_parishes <- parish_boundaries_hosp_df[order(-travel_times), ][1:10, ]
            top_parishes_list_hosp[[scenario_col]] <- data.frame(
                Parish_Name = top_parishes$PName2016,
                Travel_Time_Increase = top_parishes[[scenario_col]],
                Scenario = scenario_col
            )
        }
    }
}



# Calculate frequency of parishes appearing in top 10 lists across all hospital-only scenarios
parish_frequency_hosp <- table(unlist(lapply(top_parishes_list_hosp, function(df) df$Parish_Name)))
parish_frequency_df_hosp <- as.data.frame(parish_frequency_hosp)
colnames(parish_frequency_df_hosp) <- c("Parish_Name", "Frequency")
parish_frequency_df_hosp$Frequency <- as.numeric(parish_frequency_df_hosp$Frequency)
parish_frequency_df_hosp <- parish_frequency_df_hosp[order(-parish_frequency_df_hosp$Frequency), ]
top_10_parish_frequency_hosp <- head(parish_frequency_df_hosp, 10)



# Add indicator to SpatVector for top 10 parishes by frequency (hospital-only)
parish_boundaries_hosp$Top_10 <- ifelse(parish_boundaries_hosp$PName2016 %in% top_10_parish_frequency_hosp$Parish_Name, "Top 10", "Not in Top 10")

# Calculate centroids for parish polygons (hospital-only)
parish_centroids_hosp <- centroids(parish_boundaries_hosp)
centroids_coords_hosp <- geom(parish_centroids_hosp)

# Create data frame for centroid coordinates and attributes
parish_centroids_df_hosp <- as.data.frame(parish_centroids_hosp) %>%
    mutate(
        PName2016 = parish_boundaries_hosp$PName2016,
        Top_10 = parish_boundaries_hosp$Top_10,
        x = centroids_coords_hosp[, "x"],
        y = centroids_coords_hosp[, "y"]
    )

# Define color palette for highlighting top 10 parishes
top_10_palette <- c("Top 10" = "red", "Not in Top 10" = "gray")

# Plot map highlighting top 10 parishes by frequency (hospital-only)
highlighted_plot_hosp <- ggplot() +
    geom_spatvector(data = parish_boundaries_hosp, aes(fill = Top_10)) +
    geom_label(
        data = parish_centroids_df_hosp %>% filter(Top_10 == "Top 10"),
        aes(label = PName2016, x = x, y = y),
        color = "black",
        fill = "white",
        size = 3,
        label.padding = unit(0.2, "lines"),
        label.size = 0.25
    ) +
    scale_fill_manual(values = top_10_palette, name = "Top 10 Parishes") +
    theme_minimal() +
    ggtitle("Top 10 Parishes with Highest Travel Time Increase (Hospitals Only)") +
    theme(
        legend.position = "bottom",
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid.major = element_line(color = "gray80", size = 0.5),
        panel.grid.minor = element_line(color = "gray90", size = 0.25),
        strip.background = element_rect(fill = "white")
    )

# Display highlighted map for hospital-only scenarios
print(highlighted_plot_hosp)
```

# Densitiy plots of the average weighted delta travel time for each scenario across all parishes for hospitals 
```{r}
# Density plots of the average weighted delta travel time for each hospital-only scenario across all parishes

# Define the list of columns for hospital-only delta TT scenarios
scenario_columns_ordered_hosp <- c(
    "Average_Weighted_Delta_TT_20_3600", "Average_Weighted_Delta_TT_20_10800", "Average_Weighted_Delta_TT_20_21600",
    "Average_Weighted_Delta_TT_40_3600", "Average_Weighted_Delta_TT_40_10800", "Average_Weighted_Delta_TT_40_21600",
    "Average_Weighted_Delta_TT_60_3600", "Average_Weighted_Delta_TT_60_10800", "Average_Weighted_Delta_TT_60_21600",
    "Average_Weighted_Delta_TT_80_3600", "Average_Weighted_Delta_TT_80_10800", "Average_Weighted_Delta_TT_80_21600",
    "Average_Weighted_Delta_TT_100_3600", "Average_Weighted_Delta_TT_100_10800", "Average_Weighted_Delta_TT_100_21600"
)

# Compute global axis limits for density plots across all hospital-only scenarios
x_limits <- range(unlist(parish_boundaries_hosp_df[scenario_columns_ordered_hosp]), na.rm = TRUE)

y_limits <- c(0, max(sapply(scenario_columns_ordered_hosp, function(col) {
    max(density(parish_boundaries_hosp_df[[col]], na.rm = TRUE)$y)
})))

# Create density plots for each hospital-only scenario
density_plot_list_hosp <- list()

for (i in seq_along(scenario_columns_ordered_hosp)) {
    scenario_col <- scenario_columns_ordered_hosp[i]
    if (scenario_col %in% names(parish_boundaries_hosp_df)) {
        p <- ggplot(parish_boundaries_hosp_df, aes(x = .data[[scenario_col]])) +
            geom_density(fill = "blue", color = "black", alpha = 0.7) +
            xlim(x_limits) +
            ylim(y_limits) +
            theme_minimal() +
            ggtitle(format_scenario_title(scenario_col)) +
            xlab("Average Weighted Delta Travel Time") +
            ylab("Density") +
            theme(
                plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
                axis.title = element_text(size = 12),
                axis.text = element_text(size = 10)
            )
        density_plot_list_hosp[[i]] <- p
    }
}

# Combine all hospital-only scenario density plots into a grid
final_density_plot_hosp <- wrap_plots(density_plot_list_hosp, ncol = 3) +
    plot_layout(guides = "collect") &
    plot_annotation(
        title = "Density Plots of Average Weighted Delta Travel Time Scenarios for Hospital Only per Parish",
        theme = theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
        )
    )

# Display the combined density plot grid
print(final_density_plot_hosp)

# Overlayed density plots for each rainfall duration (1h, 3h, 6h) showing all precipitation intensities for hospital-only scenarios
overlayed_plot_list_hosp <- list()
hours_list <- c(3600, 10800, 21600)
color_palette <- c("blue", "green", "orange", "purple", "red")

for (hour in hours_list) {
    scenarios_for_hour <- grep(paste0("_", hour), scenario_columns_ordered_hosp, value = TRUE)
    combined_df <- data.frame()

    for (scenario_col in scenarios_for_hour) {
        mm_value <- sub(".*_(\\d+)_\\d+$", "\\1", scenario_col)
        temp_df <- data.frame(
            mm = as.numeric(mm_value),
            travel_times = parish_boundaries_hosp_df[[scenario_col]]
        )
        combined_df <- rbind(combined_df, temp_df)
    }

    p <- ggplot(combined_df, aes(x = travel_times, color = as.factor(mm))) +
        geom_density(fill = NA, alpha = 0.7) +
        theme_minimal() +
        ggtitle(paste(hour / 3600, "hour(s)")) +
        xlab("Average Weighted Delta Travel Time") +
        ylab("Density") +
        scale_color_manual(values = color_palette, name = "mm Values", breaks = c(20, 40, 60, 80, 100)) +
        theme(
            plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
            axis.title = element_text(size = 12),
            axis.text = element_text(size = 10)
        )

    overlayed_plot_list_hosp[[as.character(hour)]] <- p
}

# Combine overlayed density plots for all durations into a single figure
final_overlayed_plot_hosp <- wrap_plots(overlayed_plot_list_hosp, ncol = 1) +
    plot_annotation(
        title = "Overlayed Density Plots for Different mm by Hours for Hospital Only",
        theme = theme(
            plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
        )
    )

# Display the overlayed density plots
print(final_overlayed_plot_hosp)
```

# Mean Travel Time Delta Violin plots all/hospitals
```{r}
# Step 1: Rename scenario rasters for clarity (extract mm and duration for plot labels)
x_delta_all <- delta_tt_list_all
names(x_delta_all) <- sapply(names(x_delta_all), function(name) {
    mm <- sub(".*_(\\d+)_.*", "\\1", name)
    sec <- sub(".*_(\\d+)$", "\\1", name)
    hr <- as.character(as.numeric(sec) / 3600)
    paste0(mm, "mm / ", hr, "hr")
})

x_delta_hosp <- delta_tt_list_hosp
names(x_delta_hosp) <- sapply(names(x_delta_hosp), function(name) {
    mm <- sub(".*_(\\d+)_.*", "\\1", name)
    sec <- sub(".*_(\\d+)$", "\\1", name)
    hr <- as.character(as.numeric(sec) / 3600)
    paste0(mm, "mm / ", hr, "hr")
})

# Step 2: Extract pixel-level population-weighted delta travel time values for each scenario
extract_weighted_pixels <- function(raster_list, pop_raster, facility_type) {
  out_list <- lapply(names(raster_list), function(name) {
    delta_vals <- as.numeric(values(raster_list[[name]]))
    pop_vals <- as.numeric(values(pop_raster))

    # Extract scenario metadata (intensity and duration)
    intensity <- as.numeric(sub("mm /.*", "", name))
    hours <- sub(".* / ", "", name)

    df <- data.frame(
      scenario = name,
      delta_tt = delta_vals,
      pop = pop_vals,
      Intensity = factor(intensity, levels = c(20, 40, 60, 80, 100)),
      Duration = factor(hours, levels = c("1hr", "3hr", "6hr")),
      FacilityType = facility_type
    )

    # Remove NA and zero population pixels, then expand rows by population count
    df <- df[!is.na(df$delta_tt) & !is.na(df$pop) & df$pop > 0, ]
    df$pop <- round(df$pop)
    df <- tidyr::uncount(df, weights = pop)

    return(df)
  })

  do.call(rbind, out_list)
}

# Step 3: Extract weighted pixel data for both facility types
delta_tt_all_weighted <- extract_weighted_pixels(x_delta_all, pop_density_raster, "All Facilities")
delta_tt_hosp_weighted <- extract_weighted_pixels(x_delta_hosp, pop_density_raster, "Hospitals")

# Step 4: Combine all and hospital-only data, then sample up to 100,000 pixels per group for plotting
delta_tt_weighted_combined <- bind_rows(delta_tt_all_weighted, delta_tt_hosp_weighted)

set.seed(834)
delta_tt_weighted_sampled <- delta_tt_weighted_combined %>%
  group_by(Duration, Intensity, FacilityType) %>%
  group_modify(~ {
    if(nrow(.) <= 100000) return(.)
    else sample_n(., 100000)
  }) %>%
  ungroup()

# Step 5: Create violin plots comparing travel time change distributions by facility type, intensity, and duration
ggplot(delta_tt_weighted_sampled, aes(x = Intensity, y = delta_tt, fill = FacilityType)) +
    geom_violin(
        aes(group = interaction(FacilityType, Intensity)),
        position = position_dodge(width = 0.8),
        scale = "width",
        alpha = 0.5
    ) +
    stat_summary(
        aes(group = interaction(FacilityType, Intensity)),
        fun = mean,
        geom = "point",
        position = position_dodge(width = 0.8),
        shape = 18,
        size = 2,
        color = "red",
        show.legend = FALSE
    ) +
    stat_summary(
        aes(group = interaction(FacilityType, Intensity)),
        fun = median,
        geom = "point",
        position = position_dodge(width = 0.8),
        shape = 4,
        size = 2,
        color = "blue",
        show.legend = FALSE
    ) +
    facet_wrap(~ Duration, ncol = 1, strip.position = "top") +
    scale_fill_manual(
        values = c("All Facilities" = "#1f78b4", "Hospitals" = "#a6cee3"),
        name = "Facility Type"
    ) +
    scale_y_continuous(
        breaks = c(0, 15, 30, 45, 60, 90, 120, 180, 240, 300, 360),
        limits = c(0, 360)
    ) +
    geom_vline(
        xintercept = seq(1.5, length(unique(delta_tt_weighted_sampled$Intensity)) - 0.5, by = 1),
        linetype = "dashed",
        color = "grey50"
    ) +
    labs(
        x = "Precipitation Amount (mm)",
        y = "Change in Travel Time (minutes)",
        caption = "Violin plots show sampled pixel-level distributions.\nRed points indicate mean, blue crosses indicate median."
    ) +
    theme_minimal(base_size = 8) +
    theme(
        axis.text.x = element_text(size = 8),
        axis.title.x = element_text(size = 8, face = "bold"),
        axis.title.y = element_text(size = 8, face = "bold"),
        plot.title = element_text(size = 8, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 8, hjust = 0.5),
        plot.caption = element_text(size = 8),
        strip.text = element_text(size = 8, face = "bold"),
        panel.grid.major = element_line(color = "grey80"),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        panel.spacing.y = unit(0.2, "lines"),
        plot.margin = margin(t = 5, r = 10, b = 5, l = 10),
        legend.position = c(0.0, 1.055),
        legend.justification = c("left", "top"),
        legend.direction = "horizontal",
        legend.background = element_rect(fill = scales::alpha("white", 0.8), color = NA),
        legend.box.background = element_blank(),
        legend.key.size = unit(0.7, "lines"),
        legend.title = element_text(face = "bold", size = 8),
        legend.text = element_text(size = 8)
    )

# Step 6: Save the violin plot to PDF
ggsave(
    filename = here("Figures", "Pdf_Final", "Figure_5_violin_plot.pdf"),
    plot = last_plot(),
    width = 15.92,
    height = 18.47,
    units = "cm"
)
```

# Numerical summary of the delta travel time rasters
```{r}
# Step 1: Create summary tables for pixel-level population-weighted delta travel time
summary_table_all <- delta_tt_weighted_sampled %>%
    filter(FacilityType == "All Facilities") %>%
    group_by(Duration, Intensity) %>%
    summarise(
        Mean = round(mean(delta_tt, na.rm = TRUE), 1),
        Min = round(min(delta_tt, na.rm = TRUE), 1),
        Q25 = round(quantile(delta_tt, 0.25, na.rm = TRUE), 1),
        Median = round(median(delta_tt, na.rm = TRUE), 1),
        Q75 = round(quantile(delta_tt, 0.75, na.rm = TRUE), 1),
        Max = round(max(delta_tt, na.rm = TRUE), 1),
        .groups = "drop"
    ) %>%
    mutate(Facility = "All Facilities")

summary_table_hosp <- delta_tt_weighted_sampled %>%
    filter(FacilityType == "Hospitals") %>%
    group_by(Duration, Intensity) %>%
    summarise(
        Mean = round(mean(delta_tt, na.rm = TRUE), 1),
        Min = round(min(delta_tt, na.rm = TRUE), 1),
        Q25 = round(quantile(delta_tt, 0.25, na.rm = TRUE), 1),
        Median = round(median(delta_tt, na.rm = TRUE), 1),
        Q75 = round(quantile(delta_tt, 0.75, na.rm = TRUE), 1),
        Max = round(max(delta_tt, na.rm = TRUE), 1),
        .groups = "drop"
    ) %>%
    mutate(Facility = "Hospitals")

# Step 2: Combine and pivot summary tables for display
combined_long <- bind_rows(summary_table_all, summary_table_hosp)

combined_wide <- combined_long %>%
    pivot_wider(
        id_cols = c(Duration, Intensity),
        names_from = Facility,
        values_from = c(Mean, Min, Q25, Median, Q75, Max),
        names_glue = "{Facility}_{.value}"
    ) %>%
    arrange(Duration, Intensity) %>%
    dplyr::select(
        Duration, Intensity,
        `All Facilities_Mean`, `All Facilities_Min`, `All Facilities_Q25`, `All Facilities_Median`, `All Facilities_Q75`, `All Facilities_Max`,
        `Hospitals_Mean`, `Hospitals_Min`, `Hospitals_Q25`, `Hospitals_Median`, `Hospitals_Q75`, `Hospitals_Max`
    )

# Step 3: Clean column names for table output
colnames(combined_wide) <- c(
    "Duration", "Intensity",
    "Mean", "Min", "Q25", "Median", "Q75", "Max",   # All Facilities
    "Mean", "Min", "Q25", "Median", "Q75", "Max"    # Hospitals
)

# Step 4: Render styled summary table for HTML output
kable(combined_wide, format = "html", digits = 1, align = c("l", "l", rep("c", 12)),
            caption = "Flood Scenario Travel Time Summary by Facility Type (All vs Hospitals)") %>%
    add_header_above(c(" " = 2, 
                                         "All Facilities" = 6, 
                                         "Hospitals" = 6)) %>%
    kable_styling(full_width = FALSE, position = "center", font_size = 13) %>%
    column_spec(2, border_right = "1px solid", include_thead = TRUE) %>%   # after Intensity
    column_spec(8, border_right = "1px dashed", include_thead = TRUE)      # after All Facilities
```

# Export the summary table to a word document
```{r}
# 1. Rename columns for clarity and uniqueness (All Facilities vs Hospitals)
colnames(combined_wide) <- c(
    "Duration", "Intensity",
    "All_Mean", "All_Min", "All_Q25", "All_Median", "All_Q75", "All_Max",
    "Hosp_Mean", "Hosp_Min", "Hosp_Q25", "Hosp_Median", "Hosp_Q75", "Hosp_Max"
)

# 2. Build a formatted summary table using flextable for Word export
ft <- flextable(combined_wide) %>%
    set_header_labels(
        All_Mean = "Mean", All_Min = "Min", All_Q25 = "Q25", All_Median = "Median", All_Q75 = "Q75", All_Max = "Max",
        Hosp_Mean = "Mean", Hosp_Min = "Min", Hosp_Q25 = "Q25", Hosp_Median = "Median", Hosp_Q75 = "Q75", Hosp_Max = "Max"
    ) %>%

    # Add a header row to distinguish facility types
    add_header_row(
        values = c("Duration", "Intensity", "All Facilities", "Hospitals"),
        colwidths = c(1, 1, 6, 6)
    ) %>%

    theme_box() %>%
    align(align = "center", part = "all") %>%
    fontsize(size = 10, part = "all") %>%
    font(fontname = "Arial", part = "all") %>%
    padding(padding = 4, part = "all") %>%
    set_table_properties(layout = "autofit", width = .9) %>%

    # Add vertical borders to visually separate facility groups
    border(j = 2, border = fp_border(color = "black", width = 1), part = "all") %>%   # after Intensity
    border(j = 8, border = fp_border(color = "black", width = 1, style = "dashed"), part = "all") %>%  # after All_Max

    # Apply box styling to the table
    theme_box()

# 3. Export the formatted table to a Word document
doc <- read_docx() %>%
    body_add_par("Travel Time Summary under Flood Scenarios", style = "heading 1") %>%
    body_add_flextable(ft)

print(doc, target = here::here("Manuscripts", "flood_summary_table.docx"))
```

# Helper functions for top-10, skewness, and narrative analysis
```{r}
# ── Helper functions ──────────────────────────────────────────────────────────

#' Format a scenario column name to a human-readable label
#' e.g. "Average_Weighted_Delta_TT_100_3600" → "100 mm, 1 hour"
format_scenario <- function(col_name) {
    parts     <- regmatches(col_name, regexpr("\\d+_\\d+$", col_name))
    nums      <- as.numeric(unlist(strsplit(parts, "_")))
    hours_map <- c("3600" = 1, "10800" = 3, "21600" = 6)
    paste0(nums[1], " mm, ", hours_map[as.character(nums[2])], " hour")
}

#' Return Average_Weighted_Delta_TT_* columns sorted by intensity then duration
get_scenario_cols <- function(df) {
    cols      <- grep("^Average_Weighted_Delta_TT_", names(df), value = TRUE)
    intensity <- as.numeric(sub(".*_(\\d+)_(\\d+)$", "\\1", cols))
    duration  <- as.numeric(sub(".*_(\\d+)$", "\\1", cols))
    cols[order(intensity, duration)]
}

#' Build a long top-n table from a parish data frame
top_n_by_scenario <- function(df, cols, tier, n = 10) {
    purrr::map_dfr(cols, function(col) {
        df %>%
            dplyr::transmute(
                Scenario              = format_scenario(col),
                Parish                = PName2016,
                Population_Sum        = Population_Sum,
                Avg_Weighted_Delta_TT = .data[[col]]
            ) %>%
            dplyr::arrange(dplyr::desc(Avg_Weighted_Delta_TT)) %>%
            dplyr::slice_head(n = n) %>%
            dplyr::mutate(Rank = dplyr::row_number(), Tier = tier)
    })
}

#' Frequency table: how often each parish appears in the top-n across scenarios
frequency_top_n <- function(long_tbl) {
    long_tbl %>%
        dplyr::group_by(Parish, Tier) %>%
        dplyr::summarise(
            Frequency     = dplyr::n(),
            Median_Rank   = round(median(Rank), 1),
            Mean_Increase = round(mean(Avg_Weighted_Delta_TT), 1),
            .groups       = "drop"
        ) %>%
        dplyr::arrange(dplyr::desc(Frequency), Median_Rank)
}
```

# Top-10 parishes by scenario and frequency tables
```{r}
# ── Build long top-10 tables ─────────────────────────────────────────────────
top10_long_hosp <- top_n_by_scenario(
    parish_boundaries_hosp_df,
    get_scenario_cols(parish_boundaries_hosp_df),
    tier = "Hospitals only"
)

top10_long_all <- top_n_by_scenario(
    parish_boundaries_all_df,
    get_scenario_cols(parish_boundaries_all_df),
    tier = "All facilities"
)

# ── Render: Top 10 parishes by scenario, hospitals only ──────────────────────
top10_long_hosp %>%
    dplyr::select(Scenario, Rank, Parish, Avg_Weighted_Delta_TT, Population_Sum) %>%
    kable(
        format    = "html",
        digits    = 1,
        col.names = c("Scenario", "Rank", "Parish",
                       "Avg Weighted \u0394 TT (min)", "Population"),
        caption   = "Top 10 parishes by scenario \u2013 Hospitals only"
    ) %>%
    kable_styling(full_width = FALSE, position = "center", font_size = 13) %>%
    collapse_rows(columns = 1, valign = "top")

# ── Render: Top 10 parishes by scenario, all facilities ──────────────────────
top10_long_all %>%
    dplyr::select(Scenario, Rank, Parish, Avg_Weighted_Delta_TT, Population_Sum) %>%
    kable(
        format    = "html",
        digits    = 1,
        col.names = c("Scenario", "Rank", "Parish",
                       "Avg Weighted \u0394 TT (min)", "Population"),
        caption   = "Top 10 parishes by scenario \u2013 All facilities"
    ) %>%
    kable_styling(full_width = FALSE, position = "center", font_size = 13) %>%
    collapse_rows(columns = 1, valign = "top")

# ── Frequency tables ─────────────────────────────────────────────────────────
freq_hosp <- frequency_top_n(top10_long_hosp)
freq_all  <- frequency_top_n(top10_long_all)

freq_hosp %>%
    dplyr::select(-Tier) %>%
    kable(
        format    = "html",
        digits    = 1,
        col.names = c("Parish", "Frequency", "Median Rank",
                       "Mean \u0394 TT (min)"),
        caption   = "Parishes most frequently in top 10 across scenarios \u2013 Hospitals only"
    ) %>%
    kable_styling(full_width = FALSE, position = "center", font_size = 13)

freq_all %>%
    dplyr::select(-Tier) %>%
    kable(
        format    = "html",
        digits    = 1,
        col.names = c("Parish", "Frequency", "Median Rank",
                       "Mean \u0394 TT (min)"),
        caption   = "Parishes most frequently in top 10 across scenarios \u2013 All facilities"
    ) %>%
    kable_styling(full_width = FALSE, position = "center", font_size = 13)
```

# Skewness diagnostics and narrative inline objects
```{r}
# ── Parish-level distribution diagnostics (96 parishes) ─────────────────────
compute_diagnostics <- function(df, cols, tier) {
    purrr::map_dfr(cols, function(col) {
        x <- df[[col]]
        tibble::tibble(
            Tier           = tier,
            Scenario       = format_scenario(col),
            Mean           = round(mean(x, na.rm = TRUE), 1),
            Median         = round(median(x, na.rm = TRUE), 1),
            IQR            = round(IQR(x, na.rm = TRUE), 1),
            P90            = round(quantile(x, 0.90, na.rm = TRUE), 1),
            P95            = round(quantile(x, 0.95, na.rm = TRUE), 1),
            Max            = round(max(x, na.rm = TRUE), 1),
            tail_ratio_95  = round(max(x, na.rm = TRUE) /
                                   quantile(x, 0.95, na.rm = TRUE), 2),
            tail_ratio_med = round(max(x, na.rm = TRUE) /
                                   median(x, na.rm = TRUE), 2)
        )
    })
}

diag_hosp <- compute_diagnostics(
    parish_boundaries_hosp_df,
    get_scenario_cols(parish_boundaries_hosp_df),
    "Hospitals only"
)

diag_all <- compute_diagnostics(
    parish_boundaries_all_df,
    get_scenario_cols(parish_boundaries_all_df),
    "All facilities"
)

skewness_tbl <- dplyr::bind_rows(diag_hosp, diag_all)

# ── Render diagnostics table ─────────────────────────────────────────────────
skewness_tbl %>%
    kable(
        format  = "html",
        digits  = 2,
        caption = "Parish-level distribution diagnostics for population-weighted \u0394 travel time"
    ) %>%
    kable_styling(full_width = FALSE, position = "center", font_size = 13) %>%
    pack_rows(index = c("Hospitals only" = nrow(diag_hosp),
                         "All facilities" = nrow(diag_all)))

# ── Narrative inline objects ─────────────────────────────────────────────────
# Top 3 recurrent parishes (character vector: "Parish (freq)")
top3_recurrent_hosp <- freq_hosp %>%
    dplyr::slice_head(n = 3) %>%
    dplyr::mutate(label = paste0(Parish, " (", Frequency, "\u00d7)")) %>%
    dplyr::pull(label)

top3_recurrent_all <- freq_all %>%
    dplyr::slice_head(n = 3) %>%
    dplyr::mutate(label = paste0(Parish, " (", Frequency, "\u00d7)")) %>%
    dplyr::pull(label)

# Narrative statements: one per scenario per tier
make_narrative <- function(diag_tbl, tier_label) {
    diag_tbl %>%
        dplyr::mutate(label = paste0(
            "In ", tier_label, ", the maximum parish increase is ",
            Max, " min under ", Scenario,
            ", compared with a median of ", Median,
            " min and P95 of ", P95, " min."
        )) %>%
        dplyr::pull(label)
}

narrative_hosp <- make_narrative(diag_hosp, "hospitals only")
narrative_all  <- make_narrative(diag_all,  "all facilities")

narrative_all
narrative_hosp
```

# Export Supplementary Tables S3, S4, S5 to Word
```{r}
# ── Prepare data for flextable ───────────────────────────────────────────────
# S3: Top 10 parishes by scenario – Hospitals only
s3_df <- top10_long_hosp %>%
    dplyr::select(
        Scenario, Rank, Parish,
        Population = Population_Sum,
        `Avg Pop-Weighted Δ TT (min)` = Avg_Weighted_Delta_TT
    ) %>%
    dplyr::mutate(`Avg Pop-Weighted Δ TT (min)` = round(`Avg Pop-Weighted Δ TT (min)`, 1),
                  Population = round(Population, 0))

# S4: Top 10 parishes by scenario – All Facilities
s4_df <- top10_long_all %>%
    dplyr::select(
        Scenario, Rank, Parish,
        Population = Population_Sum,
        `Avg Pop-Weighted Δ TT (min)` = Avg_Weighted_Delta_TT
    ) %>%
    dplyr::mutate(`Avg Pop-Weighted Δ TT (min)` = round(`Avg Pop-Weighted Δ TT (min)`, 1),
                  Population = round(Population, 0))

# S5: Frequency across scenarios (both tiers)
s5_df <- dplyr::bind_rows(freq_hosp, freq_all) %>%
    dplyr::select(
        Parish, Tier, Frequency,
        `Median Rank`        = Median_Rank,
        `Mean Δ TT (min)`    = Mean_Increase
    ) %>%
    dplyr::arrange(Tier, dplyr::desc(Frequency))

# ── Build flextables ─────────────────────────────────────────────────────────
style_ft <- function(ft) {
    ft %>%
        font(fontname = "Arial", part = "all") %>%
        fontsize(size = 10, part = "all") %>%
        bold(part = "header") %>%
        align(align = "center", part = "all") %>%
        align(j = "Parish", align = "left", part = "body") %>%
        padding(padding = 4, part = "all") %>%
        set_table_properties(layout = "autofit", width = 0.9) %>%
        theme_box()
}

ft_s3 <- flextable(s3_df) %>%
    merge_v(j = "Scenario") %>%
    valign(j = "Scenario", valign = "top") %>%
    style_ft()

ft_s4 <- flextable(s4_df) %>%
    merge_v(j = "Scenario") %>%
    valign(j = "Scenario", valign = "top") %>%
    style_ft()

ft_s5 <- flextable(s5_df) %>%
    style_ft() %>%
    align(j = "Parish", align = "left", part = "body")

# ── Assemble Word document ───────────────────────────────────────────────────
doc_supp <- read_docx() %>%
    body_add_par("Supplementary Table S3 \u2013 Top 10 Parishes by Scenario (Hospitals Only)",
                 style = "heading 1") %>%
    body_add_flextable(ft_s3) %>%
    body_add_break(pos = "after") %>%

    body_add_par("Supplementary Table S4 \u2013 Top 10 Parishes by Scenario (All Facilities)",
                 style = "heading 1") %>%
    body_add_flextable(ft_s4) %>%
    body_add_break(pos = "after") %>%

    body_add_par("Supplementary Table S5 \u2013 Parishes Most Frequently Appearing Among the Top 10 Across Scenarios",
                 style = "heading 1") %>%
    body_add_flextable(ft_s5)

print(doc_supp, target = here::here("Manuscripts",
                                     "Supplementary_Tables_Flood_Accessibility.docx"))
```