# Load packages

```{r}
# Set up renv for package management
if (!requireNamespace("renv", quietly = TRUE)) {
    install.packages("renv")
} else {
    message("renv is already installed.")
}

# Load the renv package
library(renv)

# Check if renv is already active, activate if not
if (!renv::status()$synchronized) {
    renv::activate()
}

# To restore the environment (for a collaborator)
renv::restore()

# Load packages
library(sf)
library(dplyr)
library(foreach)
library(snow)
library(terra)
library(raster)
library(ggplot2)
library(gridExtra)
library(tidyterra)
library(cowplot)
library(viridis)
library(here)
```

# Load pop_impacted all facilities
```{r}
# Load baseline scenario with no flood
baseline_all <- read.csv(here("AccessMod/Output/Base/zonalStat__MLC0_all.csv"))

# Define the base directory for Flood_all
base_dir <- here("AccessMod/Output/Flood_all")

# Recursively search for all CSV files in the zonalStat folders
pop_coverage_all_files <- list.files(
    path = base_dir,
    pattern = "*.csv",
    full.names = TRUE,
    recursive = TRUE
)

# Filter files to include only those within zonalStat folders
pop_coverage_all_files <- pop_coverage_all_files[grepl("zonalStat", pop_coverage_all_files)]

# Read all CSV files into a list
pop_coverage_all <- lapply(pop_coverage_all_files, read.csv)

# Name each element in the list based on the folder structure and add _all to each name
names(pop_coverage_all) <- paste0(
        gsub(
                pattern = "zonalStat__|\\.csv", 
                replacement = "", 
                x = basename(pop_coverage_all_files)
        ), 
        "_all"
)

# Add baseline_all to pop_coverage_all
pop_coverage_all[["baseline_all"]] <- baseline_all

# Convert columns to factors for all MLC dataframes, including baseline
for (name in names(pop_coverage_all)) {
    if (grepl("MLC", name) || name == "baseline_all") {
        pop_coverage_all[[name]]$table.pname2016 <- as.factor(pop_coverage_all[[name]]$table.pname2016)
        pop_coverage_all[[name]]$table.time_m <- as.factor(pop_coverage_all[[name]]$table.time_m)
    }
}

# Load each as a separate dataframe object
for (i in 1:length(pop_coverage_all)) {
        assign(
                x = names(pop_coverage_all)[i],
                value = pop_coverage_all[[i]]
        )
}
```

# Population coverage difference calculation for all facilities
```{r}
# List to store differences
pop_coverage_all_diff <- list()

# Iterate through all flood scenarios in pop_coverage_all
pop_coverage_all_diff <- lapply(names(pop_coverage_all), function(scenario) {
  # Skip the baseline itself
  if (scenario == "baseline_all") return(NULL)
  
  # Get the flood scenario data
  flood_data <- pop_coverage_all[[scenario]]
  
  # Merge the flood scenario with the baseline by parish name and time threshold
  flood_diff <- left_join(flood_data, baseline_all, 
                          by = c("table.pname2016" = "table.pname2016", 
                                 "table.time_m" = "table.time_m"), 
                          suffix = c("_flood", "_baseline")) %>%
    mutate(
      popTotal_diff = table.popTotal_flood - table.popTotal_baseline,
      popTravelTime_diff = table.popTravelTime_flood - table.popTravelTime_baseline,
      popCoveredPercent_diff = table.popCoveredPercent_flood - table.popCoveredPercent_baseline
    ) %>%
    dplyr::select(table.pname2016, table.time_m, popTotal_diff, popTravelTime_diff, popCoveredPercent_diff)
  
  # Return the difference data
  return(flood_diff)
})

# Name each result by the scenario for easy identification
names(pop_coverage_all_diff) <- names(pop_coverage_all)
pop_coverage_all_diff <- pop_coverage_all_diff[!sapply(pop_coverage_all_diff, is.null)] # Remove NULL results
```

# Plot the percentage change in travel time coverage for all facilities 
```{r}
# Generate a list of plots for all scenarios
plot_list <- lapply(names(pop_coverage_all_diff), function(scenario_name) {
  ggplot(pop_coverage_all_diff[[scenario_name]], 
         aes(x = table.time_m, y = popCoveredPercent_diff, fill = as.factor(table.time_m))) +
    geom_bar(stat = "identity", position = "dodge") +
    facet_wrap(~ table.pname2016, scales = "fixed") +
    scale_y_continuous(limits = c(-100, 0)) +
    labs(
      title = paste("Percentage Change in Travel Time Coverage by Time Threshold for Each Parish (Scenario:", scenario_name, ")"),
      x = "Time Threshold (minutes)",
      y = "Percentage Change in Travel Time Coverage (%)",
      fill = "Time Threshold"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
      legend.position = "bottom",
      strip.text = element_text(size = 8),
      panel.background = element_rect(fill = "white", colour = "white"),
      plot.background = element_rect(fill = "white", colour = "white")
    )
})

# Name each plot in the list for easy reference
names(plot_list) <- names(pop_coverage_all_diff)

# Example: Display the plot for the first scenario
plot_list[[1]]

### [OPTIONAL] Save the plots as PNG files ###
# Create a new subfolder directly within Figures
# output_dir <- here("Figures", "pop_coverage_all_diff_plots")
# if (!dir.exists(output_dir)) {
#    dir.create(output_dir, recursive = TRUE)
# }

# Save all the plots as PNG files in the new subfolder
#for (scenario_name in names(plot_list)) {
#    ggsave(
#        filename = file.path(output_dir, paste0("pop_coverage_all_diff_", scenario_name, ".png")), 
#        plot = plot_list[[scenario_name]], 
#        width = 30.48, 
#        height = 20.32, 
#        units = "cm", 
#        dpi = 300
#    )
#}

# Create a plot list for each scenario
plot_list_per_scenario_all <- lapply(names(pop_coverage_all_diff), function(scenario_name) {
  # Get data for the current scenario
  scenario_data <- pop_coverage_all_diff[[scenario_name]]
  
  # Create a list of plots for each time category within the scenario
  time_categories <- unique(scenario_data$table.time_m)
  
  time_category_plots <- lapply(time_categories, function(time_category) {
    # Filter data for the specific time category
    filtered_data <- scenario_data %>%
      filter(table.time_m == time_category) %>%
      arrange(desc(popCoveredPercent_diff))  # Arrange by descending percentage difference
    
    # Create the bar plot
    ggplot(filtered_data, aes(x = reorder(table.pname2016, popCoveredPercent_diff), 
                              y = popCoveredPercent_diff, 
                              fill = popCoveredPercent_diff)) +
      geom_bar(stat = "identity") +
      scale_y_continuous(limits = c(-100, 0)) +  # Standardize y-axis
      scale_fill_gradient(
        limits = c(-100, 0),  # Ensure the legend matches the y-axis
        low = "red",          # -100 is red
        high = "blue"         # 0 is blue
      ) +
      labs(
        title = paste("Percentage Change in Travel Time Coverage (", time_category, " mins)"),
        x = "Parish",
        y = "Percentage Change (%)",
        fill = "Percentage Change"
      ) +
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
        axis.text.y = element_text(size = 8),
        legend.position = "right",
        plot.title = element_text(size = 12, hjust = 0.5),
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white")
      )
  })
  
  # Name the plots by the time category
  names(time_category_plots) <- paste0("Time_", time_categories)
  return(time_category_plots)
})

# Name the plot list by scenarios for easy reference
names(plot_list_per_scenario_all) <- names(pop_coverage_all_diff)

# Example: Access the plot for a specific scenario and time category
plot_list_per_scenario_all[["MLC_100_10800_all"]][["Time_15"]]

# [OPTIONAL] Save all the plots as PNG files in the new subfolder
output_dir <- here("Figures", "pop_coverage_all_diff_plots_per_scenario")
if (!dir.exists(output_dir)) {
   dir.create(output_dir, recursive = TRUE)
}

# Create a subfolder for each MLC scenario
for (scenario_name in names(plot_list_per_scenario_all)) {
  scenario_dir <- file.path(output_dir, scenario_name)
  if (!dir.exists(scenario_dir)) {
    dir.create(scenario_dir, recursive = TRUE)
  }
  
  # Save all the plots as PNG files in the new subfolder
  for (time_category in names(plot_list_per_scenario_all[[scenario_name]])) {
    ggsave(
      filename = file.path(scenario_dir, paste0("pop_coverage_all_diff_", scenario_name, "_", time_category, ".png")),
      plot = plot_list_per_scenario_all[[scenario_name]][[time_category]],
      width = 30.48,
      height = 20.32,
      units = "cm",
      dpi = 300
    )
  }
}
```

# Load pop_impacted hospitals only
```{r}
# Load baseline scenario with no flood
baseline_hosp <- read.csv(here("AccessMod/Output/Base/zonalStat__MLC0_hosp.csv"))

# Define the base directory for Flood_hosp
base_dir <- here("AccessMod/Output/Flood_hosp")

# Recursively search for all CSV files in the zonalStat folders
pop_coverage_hosp_files <- list.files(
    path = base_dir,
    pattern = "*.csv",
    full.names = TRUE,
    recursive = TRUE
)

# Filter files to include only those within zonalStat folders
pop_coverage_hosp_files <- pop_coverage_hosp_files[grepl("zonalStat", pop_coverage_hosp_files)]

# Read all CSV files into a list
pop_coverage_hosp <- lapply(pop_coverage_hosp_files, read.csv)

# Name each element in the list based on the folder structure and add _hosp to each name
names(pop_coverage_hosp) <- paste0(
    gsub(
        pattern = "zonalStat__|\\.csv", 
        replacement = "", 
        x = basename(pop_coverage_hosp_files)
    ), 
    "_hosp"
)

# Convert columns to factors for all MLC dataframes, including baseline
for (name in names(pop_coverage_hosp)) {
    if (grepl("MLC", name)) {
        pop_coverage_hosp[[name]]$table.pname2016 <- as.factor(pop_coverage_hosp[[name]]$table.pname2016)
        pop_coverage_hosp[[name]]$table.time_m <- as.factor(pop_coverage_hosp[[name]]$table.time_m)
    }
}

baseline_hosp$table.pname2016 <- as.factor(baseline_hosp$table.pname2016)
baseline_hosp$table.time_m <- as.factor(baseline_hosp$table.time_m)

# Load each as a separate dataframe object
for (i in 1:length(pop_coverage_hosp)) {
    assign(
        x = names(pop_coverage_hosp)[i],
        value = pop_coverage_hosp[[i]]
    )
}

# Remove redundant objects from the global environment
rm(base_dir, pop_coverage_hosp_files, i, name)
```

# Calculate Population Coverage Difference for the Hospitals-Only Scenario
```{r}
# List to store differences
pop_coverage_hosp_diff <- list()

# Iterate through all flood scenarios in pop_coverage_hosp
pop_coverage_hosp_diff <- lapply(names(pop_coverage_hosp), function(scenario) {
  # Skip the baseline itself
  if (scenario == "baseline_hosp") return(NULL)
  
  # Get the flood scenario data
  flood_data <- pop_coverage_hosp[[scenario]]
  
  # Merge the flood scenario with the baseline by parish name and time threshold
  flood_diff <- left_join(flood_data, baseline_hosp, 
                          by = c("table.pname2016" = "table.pname2016", 
                                 "table.time_m" = "table.time_m"), 
                          suffix = c("_flood", "_baseline")) %>%
    mutate(
      popTotal_diff = table.popTotal_flood - table.popTotal_baseline,
      popTravelTime_diff = table.popTravelTime_flood - table.popTravelTime_baseline,
      popCoveredPercent_diff = table.popCoveredPercent_flood - table.popCoveredPercent_baseline
    ) %>%
    dplyr::select(table.pname2016, table.time_m, popTotal_diff, popTravelTime_diff, popCoveredPercent_diff)
  
  # Return the difference data
  return(flood_diff)
})

# Name each result by the scenario for easy identification
names(pop_coverage_hosp_diff) <- names(pop_coverage_hosp)
pop_coverage_hosp_diff <- pop_coverage_hosp_diff[!sapply(pop_coverage_hosp_diff, is.null)] # Remove NULL results
```

# Plot the percentage change in travel time coverage for hospitals only
```{r}
# Generate a list of plots for all hosp-only scenarios
plot_list_hosp <- lapply(names(pop_coverage_hosp_diff), function(scenario_name) {
  ggplot(pop_coverage_hosp_diff[[scenario_name]], 
         aes(x = table.time_m, y = popCoveredPercent_diff, fill = as.factor(table.time_m))) +
    geom_bar(stat = "identity", position = "dodge") +
    facet_wrap(~ table.pname2016, scales = "fixed") +
    scale_y_continuous(limits = c(-100, 0)) +
    labs(
      title = paste("Percentage Change in Travel Time Coverage by Time Threshold for Each Parish (Scenario:", scenario_name, ")"),
      x = "Time Threshold (minutes)",
      y = "Percentage Change in Travel Time Coverage (%)",
      fill = "Time Threshold"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
      legend.position = "bottom",
      strip.text = element_text(size = 8),
      panel.background = element_rect(fill = "white", colour = "white"),
      plot.background = element_rect(fill = "white", colour = "white")
    )
})

# Name each plot in the list for easy reference
names(plot_list_hosp) <- names(pop_coverage_hosp_diff)

# Example: Display the plot for the first scenario
plot_list_hosp[[1]]

### [OPTIONAL] Save the plots as PNG files ###
# Create a new subfolder directly within Figures for hosp-only scenario
output_dir_hosp <- here("Figures", "pop_coverage_hosp_diff_plots")
if (!dir.exists(output_dir_hosp)) dir.create(output_dir_hosp, recursive = TRUE)

# Save all the plots as PNG files in the new subfolder
for (scenario_name in names(plot_list_hosp)) {
  ggsave(
    filename = file.path(output_dir_hosp, paste0("pop_coverage_hosp_diff_", scenario_name, ".png")), 
    plot = plot_list_hosp[[scenario_name]], 
    width = 30.48, 
    height = 20.32, 
    units = "cm", 
    dpi = 300
  )
}

# Create a plot list for each hosp-only scenario
plot_list_hosp_per_scenario <- lapply(names(pop_coverage_hosp_diff), function(scenario_name) {
  # Get data for the current scenario
  scenario_data <- pop_coverage_hosp_diff[[scenario_name]]
  
  # Create a list of plots for each time category within the scenario
  time_categories <- unique(scenario_data$table.time_m)
  
  time_category_plots <- lapply(time_categories, function(time_category) {
    # Filter data for the specific time category
    filtered_data <- scenario_data %>%
      filter(table.time_m == time_category) %>%
      arrange(desc(popCoveredPercent_diff))  # Arrange by descending percentage difference
    
    # Create the bar plot
    ggplot(filtered_data, aes(x = reorder(table.pname2016, popCoveredPercent_diff), 
                              y = popCoveredPercent_diff, 
                              fill = popCoveredPercent_diff)) +
      geom_bar(stat = "identity") +
      scale_y_continuous(limits = c(-100, 0)) +  # Standardize y-axis
      scale_fill_gradient(
        limits = c(-100, 0),  # Ensure the legend matches the y-axis
        low = "red",          # -100 is red
        high = "blue"         # 0 is blue
      ) +
      labs(
        title = paste("Percentage Change in Travel Time Coverage (", time_category, " mins)"),
        x = "Parish",
        y = "Percentage Change (%)",
        fill = "Percentage Change"
      ) +
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
        axis.text.y = element_text(size = 8),
        legend.position = "right",
        plot.title = element_text(size = 12, hjust = 0.5),
        panel.background = element_rect(fill = "white", colour = "white"),
        plot.background = element_rect(fill = "white", colour = "white")
      )
  })
  
  # Name the plots by the time category
  names(time_category_plots) <- paste0("Time_", time_categories)
  return(time_category_plots)
})

# Name the plot list by scenarios for easy reference
names(plot_list_hosp_per_scenario) <- names(pop_coverage_hosp_diff)

# Example: Access the plot for a specific scenario and time category
plot_list_hosp_per_scenario[["MLC_100_3600_hosp"]][["Time_30"]]

# [OPTIONAL] Save all the plots as PNG files in the new subfolder for hosp-only scenario
output_dir_hosp <- here("Figures", "pop_coverage_hosp_diff_plots_per_scenario")
if (!dir.exists(output_dir_hosp)) {
  dir.create(output_dir_hosp, recursive = TRUE)
}

# Create a subfolder for each MLC scenario
for (scenario_name in names(plot_list_hosp_per_scenario)) {
  scenario_dir <- file.path(output_dir_hosp, scenario_name)
  if (!dir.exists(scenario_dir)) {
   dir.create(scenario_dir, recursive = TRUE)
  }
  
  # Save all the plots as PNG files in the new subfolder
  for (time_category in names(plot_list_hosp_per_scenario[[scenario_name]])) {
   ggsave(
    filename = file.path(scenario_dir, paste0("pop_coverage_hosp_diff_", scenario_name, "_", time_category, ".png")),
    plot = plot_list_hosp_per_scenario[[scenario_name]][[time_category]],
    width = 30.48,
    height = 20.32,
    units = "cm",
    dpi = 300
   )
  }
}

```
