# Pre-processing of raw geospatial inputs for the Kampala study area.
# Crops, reprojects, and cleans all spatial datasets to a common extent
# and CRS defined by the Kampala parish boundaries.

```{r}
source(here::here("Code/utils/loading_utils.R"))

library(Spatial.align)
```

```{r}
# Load raw spatial datasets
parishes <- vect(here("Data/raw/Adm_boundaries/PARISHES_2016_UTM_36N.shp"))
dem30m <- rast(here("Data/raw/DEM/Uganda_SRTM30meters.tif"))
facilities_all <- vect(here("Data/raw/Facilities/UG_HF_all.shp"))
facilities_hosp <- vect(here("Data/raw/Facilities/UG_HF_hosp.shp"))
landcover10m <- rast(here("Data/raw/Landcover/ESA_WorldCover_10m_2021_v200_N00E030_Map.tif"))
osm_roads <- vect(here("Data/raw/OSM_roads/hotosm_uga_roads_lines.shp"))
water_lines <- vect(here("Data/raw/Water_lines/Kamp_rivers_streams_canals.shp"))

# Aggregate WorldPop age-sex population density layers into total population
worldpop_density <- list.files(here("Data/raw/Worldpop_density"), pattern = "\\.tif$", full.names = TRUE) %>%
  lapply(rast) %>%
  {Reduce(`+`, .)}

# Load WorldPop development indicator rasters (mean aggregates only)
worldpop_dev_indicators <- here("Data/raw/Worldpop_dev_indicators") %>%
  list.files(pattern = "MEAN\\.tif$", full.names = TRUE, recursive = TRUE) %>%
  lapply(rast)

# Load Plasmodium falciparum incidence rate raster (GBD 2023, reference year 2016)
malaria <- rast(here("Data/raw/Worldpop_dev_indicators/Malaria/2024_GBD2023_Global_Pf_Incidence_Rate_2016.tif")) 
malaria <- malaria[["Data"]]
```

# Filter and export Kampala parish boundaries
```{r}
parishes_kamp <- parishes[parishes$DName2016 == "KAMPALA",]
writeVector(parishes_kamp, here("Data/processed/Adm_boundaries/parishes_kampala.shp"), overwrite = TRUE)
```

# Align all datasets to the Kampala parish extent and CRS
```{r}
dem30m <- align(dem30m, parishes_kamp, show_plots = TRUE)
facilities_all <- align(facilities_all, parishes_kamp)
facilities_hosp <- align(facilities_hosp, parishes_kamp)
landcover10m <- align(landcover10m, parishes_kamp)
osm_roads <- align(osm_roads, parishes_kamp)
water_lines <- align(water_lines, parishes_kamp)
worldpop_density <- align(worldpop_density, parishes_kamp)
worldpop_dev_indicators <- lapply(worldpop_dev_indicators, function(x) align(x, parishes_kamp))
malaria = align(malaria, parishes_kamp)
```

# Reclassify landcover to ESA WorldCover standard classes
```{r}
reclass_matrix <- matrix(c(
  0, 19.999999, 10,
  20, 29.999999, 20,
  30, 39.999999, 30,
  40, 49.999999, 40,
  50, 59.999999, 50,
  60, 69.999999, 60,
  70, 79.999999, 70,
  80, 89.999999, 80,
  90, 94.999999, 90,
  95, 99.999999, 95,
  100, 999.999999, 100
), ncol = 3, byrow = TRUE)

landcover10m <- classify(landcover10m, reclass_matrix)
landcover10m <- as.factor(landcover10m)

levels(landcover10m)[[1]]$ID <- c(
  "10" = "Tree cover",
  "20" = "Shrubland",
  "30" = "Grassland",
  "40" = "Cropland",
  "50" = "Built-up",
  "60" = "Bare / sparse vegetation",
  "70" = "Snow and Ice",
  "80" = "Permanent water bodies",
  "90" = "Herbaceous wetland",
  "95" = "Mangroves",
  "100" = "Moss and lichen"
)
```

# Standardise facility attribute names and assign ordered factor levels
```{r}	
facilities_all <- facilities_all %>%
  rename_with(~ gsub(" ", "_", .))

facilities_all$Facility_t <- factor(
  facilities_all$Facility_t,
  levels = c(
    "Health Centre II",
    "Health Centre III",
    "Health Centre IV",
    "Hospital",
    "Regional Referral Hospital",
    "National Referral Hospital"
  )
)

facilities_hosp <- facilities_hosp %>%
  rename_with(~ gsub(" ", "_", .))

facilities_hosp$Facility_t <- factor(
  facilities_hosp$Facility_t,
  levels = c(
    "Hospital",
    "Regional Referral Hospital",
    "National Referral Hospital"
  )
)
```

# Retain major road classes and collapse link variants into parent types
```{r}
osm_roads <- osm_roads[osm_roads$highway %in% c("motorway", "motorway_link", "primary", "primary_link", 
                                                "secondary", "secondary_link", "tertiary", "tertiary_link", 
                                                "trunk", "trunk_link"), ]

osm_roads$highway <- factor(gsub("_link", "", osm_roads$highway))
```

# Retain rivers, streams, and canals from the hydrography dataset
```{r}
water_lines <- water_lines[water_lines$waterway %in% c("river", "stream", "canal"), ]
```

# Export all processed datasets
```{r}
processed_dir <- here("Data/processed")

writeRaster(dem30m, file.path(processed_dir, "DEM", "dem30m_kampala.tif"), overwrite = TRUE)
writeVector(facilities_all, file.path(processed_dir, "Facilities", "facilities_all_kampala.shp"), overwrite = TRUE)
writeVector(facilities_hosp, file.path(processed_dir, "Facilities", "facilities_hosp_kampala.shp"), overwrite = TRUE)
writeRaster(landcover10m, file.path(processed_dir, "Landcover", "landcover10m_kampala.tif"), overwrite = TRUE)
writeVector(osm_roads, file.path(processed_dir, "OSM_roads", "osm_roads_kampala.shp"), overwrite = TRUE)
writeVector(water_lines, file.path(processed_dir, "Water_lines", "water_lines_kamp.shp"), overwrite = TRUE)
writeRaster(worldpop_density, file.path(processed_dir, "Worldpop_density", "worldpop_density_kampala.tif"), overwrite = TRUE)

lapply(worldpop_dev_indicators, function(raster) {
    raster_name <- names(raster)
    writeRaster(raster, 
                            file.path(processed_dir, "Worldpop_dev_indicators", paste0(raster_name, "_kampala.tif")), 
                            overwrite = TRUE)
})

writeRaster(malaria, file.path(processed_dir, "Worldpop_dev_indicators", "malaria_kampala.tif"), overwrite = TRUE)
```
