# Preparation of input datasets for AccessMod 5 travel time modelling.
# Copies processed spatial data into the inAccessMod folder structure,
# applies CRS projection, and processes vector/raster inputs.

```{r}
source(here::here("Code/utils/loading_utils.R"))

library(inAccessMod)
library(Spatial.align)
```

# Initialise the inAccessMod project structure
```{r}
mainPath <- here("AccessMod/InAccessMod")
country <- "Uganda"
```

# Copy processed datasets into the AccessMod input directories
## Administrative boundaries
```{r}
inputPath_adm <- here("Data/processed/Parish_boundaries")
copy_input(mainPath, country, inputPath_adm)
```

## Population density
```{r}
inputPath_pop <- here("Data/processed/Worldpop_density/worldpop_density_kampala.tif")
copy_input(mainPath, country, inputPath_pop)
```

## Digital Elevation Model
```{r}
inputPath_DEM <- here("Data/processed/DEM/dem30m_kampala.tif")
copy_input(mainPath, country, inputPath_DEM)
```

## Landcover
```{r}
inputPath_LC <- here("Data/processed/Landcover/landcover10m_kampala.tif")
copy_input(mainPath, country, inputPath_LC)
```

## Health facilities
```{r}
inputPath_HF <- here("Data/processed/Facilities")
copy_input(mainPath, country, inputPath_HF)
```

## OSM road network
```{r}
inputPath_roads <- here("Data/processed/OSM_roads")
copy_input(mainPath, country, inputPath_roads)
```

# Flood extent polygons
```{r}
inputPath_flood <- here("Data/processed/Flood_extents/Flood_extents_full")
copy_input(mainPath, country, inputPath_flood)
```

# Hydrography
```{r}
inputPath_water <- here("Data/processed/Water_lines")
copy_input(mainPath, country, inputPath_water)
```

# Set projected CRS for all vector inputs
```{r}
set_projection(mainPath, country, mostRecent = TRUE, alwaysSet = TRUE, bestCRS = TRUE)
```

# Process vector inputs; rasters are handled separately below
```{r}
process_inputs(mainPath, country, mostRecent = TRUE, alwaysProcess = FALSE)

# Align DEM and landcover to the disaggregated population raster grid
dem <- rast(here("Data/processed/DEM/dem30m_kampala.tif"))
landcover <- rast(here("Data/processed/Landcover/landcover10m_kampala.tif"))

crs(dem) <- crs(pop_total_10m)
dem <- crop(dem, pop_total_10m)
dem <- resample(dem, pop_total_10m, method = "bilinear")

crs(landcover) <- crs(pop_total_10m)
landcover <- crop(landcover, pop_total_10m)
landcover <- resample(landcover, pop_total_10m, method = "near")
```

# Compile processed data and export rasters to AccessMod
```{r}
compile_processed_data(mainPath, country, mostRecent = FALSE)

# NOTE: Hardcoded timestamp directory. Update if AccessMod project is reinitialised.
writeRaster(dem, here("AccessMod/InAccessMod/Uganda/data/zToAccessMod/20250108115609/UG_DEM_kamp.tif"), overwrite = TRUE)
writeRaster(landcover, here("AccessMod/InAccessMod/Uganda/data/zToAccessMod/20250108115609/UG_Landcover_kamp.tif"), overwrite = TRUE)
```
