# Disaggregation of WorldPop population density to 10 m resolution.
# The original ~100 m grid is disaggregated by a factor of 10 in each
# dimension, with cell values divided by 100 to preserve total population.

```{r}
source(here::here("Code/utils/loading_utils.R"))
```

# Load population raster and study area boundary
```{r}
pop_raster_total <- rast(here("Data/processed/Worldpop_density/worldpop_density_kampala.tif"))
study_area <- vect(here("Data/processed/Adm_boundaries/parishes_kampala.shp"))
```

# Reproject study area to match the population raster CRS
```{r}
study_area <- project(study_area, crs(pop_raster_total))
```

# Disaggregate to 10 × 10 m resolution and adjust values
```{r}
# Each original cell is split into 10 × 10 = 100 sub-cells
pop_total_10m <- disagg(pop_raster_total, c(10, 10))

# Divide by disaggregation factor to conserve total population count
pop_total_10m <- pop_total_10m / 100
```

# Export disaggregated population raster
```{r}
# NOTE: Hardcoded timestamp directory. Update if AccessMod project is reinitialised.
writeRaster(pop_total_10m, here("AccessMod/InAccessMod/Uganda/data/zToAccessMod/20250108115609/pop_total_10m.tif"), overwrite = TRUE)

writeRaster(pop_total_10m, here("Data/processed/Worldpop_density/pop_total_10m.tif"), overwrite = TRUE)
```
