# Construction of merged landcover rasters for AccessMod travel time modelling.
# For each flood scenario, flood extent polygons are rasterised and merged
# with the baseline landcover raster (MLC0) to produce scenario-specific
# merged landcover layers.

```{r}
source(here::here("Code/utils/loading_utils.R"))
source(here::here("Code/utils/export_utils.R"))
source(here::here("Code/00_scenario_definitions.R"))

library(sf)
library(foreach)
library(snow)
library(raster)
library(gridExtra)
library(cowplot)
library(Spatial.align)
library(ggspatial)
library(scales)
```

# Separate flood extent polygons by scenario and export as individual shapefiles
```{r}
flood_full <- st_read(here("AccessMod/InAccessMod/Uganda/data/vNaturalPolygons/20250107104606/processed/20250107121314"))

for (i in 1:nrow(scenarios_all)) {
    scenario <- scenarios_all[i, ]
    rainfall <- scenario$rainfall_mm
    duration <- scenario$duration_hr
    
    filtered_data <- flood_full[flood_full$rainfall == rainfall & flood_full$duration == duration, ]
    
    if (nrow(filtered_data) > 0) {
        subfolder_name <- scenario$scenario_id
        dir_path <- here("Data", "raw", "Flood_extents", subfolder_name)
        if (!dir.exists(dir_path)) {
            dir.create(dir_path, recursive = TRUE)
        }
        
        shapefile_name <- paste0(subfolder_name, ".shp")
        st_write(filtered_data, file.path(dir_path, shapefile_name), append = FALSE)
    }
}

# Load all scenario-specific flood extent shapefiles
flood_list <- list.files(path = here("Data", "raw", "Flood_extents"), pattern = "\\.shp$", full.names = TRUE, recursive = TRUE) %>%
    lapply(st_read)

flood_extents_names <- basename(dirname(list.files(path = here("Data", "raw", "Flood_extents"), pattern = "\\.shp$", full.names = TRUE, recursive = TRUE)))

for (i in 1:length(flood_list)) {
    flood_list[[i]]$Scenario <- gsub("flood_", "", flood_extents_names[i])
}

names(flood_list) <- flood_extents_names

# Remove original rainfall/duration columns (now encoded in scenario metadata)
flood_list <- lapply(flood_list, function(x) {
    x$rainfall <- NULL
    x$duration <- NULL
    return(x)
})

```

# Merge flood extent rasters with base landcover for each scenario
```{r}
merged_landcovers <- list()

# Baseline merged landcover raster produced by AccessMod
MLC0 <- rast(here("Data/processed/Merged_landcovers/raster_land_cover_merged_MLC0.img"))
landcover <- rast(here("AccessMod/InAccessMod/Uganda/data/zToAccessMod/20250108115609/UG_Landcover_kamp.tif"))
crs(MLC0) <- crs(landcover)

for (i in 1:length(flood_list)) {
    landcov <- MLC0
    flood_extent <- flood_list[[i]] %>%
        mutate(class = case_when(
            threshold == 0.1 ~ 1,  # Flood depth class: 10 cm
            threshold == 0.2 ~ 2,  # Flood depth class: 20 cm
            threshold == 0.3 ~ 3,  # Flood depth class: 30 cm
            TRUE ~ NA_real_
        )) %>% 
        as(., "Spatial") %>%
        vect()

    # Rasterise flood polygons onto the base landcover grid
    flood_extent_rast <- terra::rasterize(flood_extent, landcov, "class", touches = TRUE)

    # Preserve NA structure from the base landcover
    flood_extent_rast <- mask(flood_extent_rast, landcov, inverse = FALSE) 

    # Overlay: flood classes take precedence over base landcover
    landcover_merge <- terra::cover(flood_extent_rast, landcov) 

    writeRaster(landcover_merge, here("Data", "processed", "Merged_landcovers", paste0("MLC_", flood_extents_names[i], ".tif")), overwrite = TRUE)
}
```

# Prepare merged landcover layers for visualisation
```{r}	
merged_landcovers <- list.files(path = here("Data", "processed", "Merged_landcovers"), pattern = "\\.tif$", full.names = TRUE, recursive = TRUE) %>%
    lapply(terra::rast)

merged_landcovers_names <- list.files(path = here("Data", "processed", "Merged_landcovers"), pattern = "\\.tif$", full.names = FALSE, recursive = TRUE)

# Sort by rainfall intensity then duration (extracted from filenames)
flood_depths <- as.numeric(sub(".*_flood_([0-9]+)_.*", "\\1", merged_landcovers_names))
durations <- as.numeric(sub(".*_([0-9]+)\\.tif$", "\\1", merged_landcovers_names))

flood_depths[is.na(flood_depths)] <- Inf
durations[is.na(durations)] <- Inf

merged_landcovers_names <- merged_landcovers_names[order(flood_depths, durations, merged_landcovers_names)]
merged_landcovers_names <- merged_landcovers_names[!grepl("full", merged_landcovers_names) & !is.na(merged_landcovers_names)]
merged_landcovers <- merged_landcovers[order(flood_depths, durations)]

for (i in 1:length(merged_landcovers)) {
    assign(gsub(".tif", "", merged_landcovers_names[i]), merged_landcovers[[i]])
}

# Legend definitions for combined landcover and flood depth classes
legend_colour <- data.frame(
    value = c(1, 2, 3, 10, 20, 30, 40, 50, 60, 80, 90, 1001, 1002, 1003, 1004, 1005),
    color = c("#169FFB", "#3C3CCF", "#170AA6", "#A8E16E", "#FFB400", "#FFFF1B", "#D7B5D8", "#FF0000", "#30123B", "#0312DC", "#15C44C", "#DE0004", "#CE0FA4", "#336CE6", "#34EF05", "#333333")
)

legend_names <- c(
    "Flood 10cm",
    "Flood 20cm",
    "Flood 30cm",
    "Tree Cover",
    "Shrubs",
    "Grass Land",
    "Agriculture",
    "Urban",
    "Bare",
    "Water Bodies",
    "Herbaceous wetland",
    "Trunk",
    "Primary",
    "Motorway",
    "Secondary",
    "Tertiary"
)

levels_df <- data.frame(
  ID = legend_colour$value,
  Category = legend_names
)

# Assign categorical levels for ggplot display
merged_landcovers <- lapply(merged_landcovers, function(r) {
    levels(r) <- levels_df
    r
})

merged_landcovers <- lapply(merged_landcovers, function(r) {
    as.factor(r)
})

names(merged_landcovers) <- scenarios_all$label_plot

color_mapping <- setNames(legend_colour$color, legend_colour$value)
```

# Faceted panel plot of all 15 merged landcover scenarios
```{r}
meta <- scenarios_all %>%
    mutate(name = label_plot,
           lab = label_plot)

ord <- meta$name
mlc_stack <- rast(merged_landcovers[ord])
names(mlc_stack) <- meta$lab

# Redefine colour palette and alpha mapping for the panel plot
legend_colour <- data.frame(
    value = c(1, 2, 3, 10, 20, 30, 40, 50, 60, 80, 90, 1001, 1002, 1003, 1004, 1005),
    color = c("#169FFB", "#3C3CCF", "#170AA6", "#A8E16E", "#FFB400", "#FFFF1B",
                        "#D7B5D8", "#FF0000", "#30123B", "#0312DC", "#15C44C",
                        "#DE0004", "#CE0FA4", "#336CE6", "#34EF05", "#333333")
)

legend_names <- c(
    "Flood 10cm", "Flood 20cm", "Flood 30cm",
    "Tree Cover", "Shrubs", "Grass Land", "Agriculture", "Urban", "Bare",
    "Water Bodies", "Herbaceous wetland",
    "Trunk", "Primary", "Motorway", "Secondary", "Tertiary"
)

pal   <- setNames(legend_colour$color, as.character(legend_colour$value))
labs  <- setNames(legend_names,            as.character(legend_colour$value))

# Flood classes (1–3) fully opaque; other classes semi-transparent
alpha_map <- setNames(
    ifelse(as.integer(names(pal)) %in% c(1, 2, 3), 1, 0.6),
    names(pal)
)

# Convert overlay vectors to sf if necessary
if (inherits(OSM_roads, "SpatVector"))  OSM_roads  <- sf::st_as_sf(OSM_roads)
if (inherits(water_lines, "SpatVector")) water_lines <- sf::st_as_sf(water_lines)

gridplot_mlc <-
        ggplot() +
        geom_spatraster(
                data = mlc_stack,
                aes(fill = as.factor(after_stat(value)), alpha = as.factor(after_stat(value)))
        ) +
        facet_wrap(~lyr, nrow = 5, ncol = 3) +
        geom_sf(data = OSM_roads,  color = "black", linewidth = 0.15, inherit.aes = FALSE) +
        geom_sf(data = water_lines, color = "blue",  linewidth = 0.12, linetype = 2, inherit.aes = FALSE) +
        scale_fill_manual(
                values = pal,
                breaks = names(pal),
                labels = labs[names(pal)],
                name   = "Merged land-cover, barriers and floods",
                na.translate = FALSE,
                na.value = "transparent",
                guide = guide_legend(
                        ncol = 2,
                        title.position = "top",
                        override.aes = list(alpha = unname(alpha_map[names(pal)]))
                )
        ) +
        scale_alpha_manual(values = alpha_map, guide = "none") +
        coord_sf(expand = FALSE) +
        theme_minimal(base_size = 8) +
        theme(
                plot.background = element_rect(fill = "white", color = NA),
                panel.background = element_rect(fill = "white", color = NA),
                axis.text = element_blank(),
                axis.title = element_blank(),
                panel.grid = element_blank(),
                strip.text = element_text(size = 8, face = "plain"),
                strip.background = element_rect(fill = "white", colour = NA),
                legend.position = "right",
                legend.title = element_text(size = 8, face = "bold"),
                legend.text  = element_text(size = 8),
                panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6)
        )

save_manuscript_pdf(gridplot_mlc, "merged_landcovers_w_floods", subdir = "QC", width = 15.92, height = 24.62)
```
