#Load libraries
```{r}
# Set up renv for reproducible package management
if (!requireNamespace("renv", quietly = TRUE)) {
  install.packages("renv")
} else {
  message("renv is already installed.")
}

# Load the renv package
library(renv)

# Activate renv if the environment is not synchronized
if (!renv::status()$synchronized) {
  renv::activate()
}

# Restore the environment from lockfile (for collaborators)
renv::restore()

# Load required libraries for spatial analysis, plotting, and data management
library(sf)         # Simple Features for vector data
library(dplyr)      # Data manipulation
library(foreach)    # Parallel processing (loops)
library(snow)       # Parallel computing support
library(terra)      # Raster data handling
library(raster)     # Legacy raster package (some functions)
library(ggplot2)    # Plotting
library(gridExtra)  # Arranging multiple plots
library(tidyterra)  # ggplot2 support for terra objects
library(cowplot)    # Advanced plot composition
library(viridis)    # Color palettes for plots
library(here)       # Reliable file paths
library(knitr)      # Dynamic report generation
```

# ALL SCENARIOS
## Flood scenarios Travel Time All
```{r}
# Load all flood scenario travel time rasters (.img files) for ALL facilities
# Define the base directory containing flood scenario outputs
base_dir <- here("AccessMod/Output/Flood_all")

# Find all scenario directories (MLC_*) within the base directory
mlc_dirs <- list.dirs(path = base_dir, full.names = TRUE, recursive = FALSE)

# Initialize a named list to store .img file paths for each scenario
scenario_img_files <- list()

# Loop through each scenario directory and find .img files in the expected subfolder
for (mlc_dir in mlc_dirs) {
  # Extract scenario name from folder name
  scenario_name <- basename(mlc_dir)
  
  # Path to the raster_travel_time_Floods_all_config subfolder
  raster_dir <- file.path(mlc_dir, "0-min/raster_travel_time_Floods_all_config")
  
  # If the raster folder exists, get all .img files
  if (dir.exists(raster_dir)) {
    img_files <- list.files(path = raster_dir, pattern = "\\.img$", full.names = TRUE, recursive = FALSE)
    
    # Add .img files to the list under the scenario name
    if (length(img_files) > 0) {
      scenario_img_files[[scenario_name]] <- img_files
    }
  }
}

# Print all scenarios and their corresponding .img files for verification
if (length(scenario_img_files) > 0) {
  cat("Flood travel time scenarios and their .img files:\n")
  print(scenario_img_files)
} else {
  cat("No .img files found in any raster_travel_time_Floods_all_config folder.\n")
}

# Load all flood scenario rasters into a list (each scenario may have multiple rasters)
x <- lapply(scenario_img_files, function(img_files) {
  lapply(img_files, rast)
})

# Assign meaningful names to each raster in the list for clarity and consistency
x_names <- c(
    "flood_tt_all_100_10800", "flood_tt_all_100_21600", "flood_tt_all_100_3600",
    "flood_tt_all_20_10800", "flood_tt_all_20_21600", "flood_tt_all_20_3600",
    "flood_tt_all_40_10800", "flood_tt_all_40_21600", "flood_tt_all_40_3600",
    "flood_tt_all_60_10800", "flood_tt_all_60_21600", "flood_tt_all_60_3600",
    "flood_tt_all_80_10800", "flood_tt_all_80_21600", "flood_tt_all_80_3600"
)

# Assign scenario names to the raster list
names(x) <- x_names

# Extract the first SpatRaster from each scenario list element
x_extracted <- lapply(x, function(obj) obj[[1]])

# Export each raster to the global environment for direct access by name
list2env(x_extracted, envir = .GlobalEnv)
```

## Plotting the travel time rasters
```{r}
# Extract the first SpatRaster from each scenario in the list 'x'
raster_list <- lapply(x, function(obj) obj[[1]])

# Combine the extracted SpatRasters into a single SpatRaster stack for plotting
stacked_rasters <- rast(raster_list)

# Define the color scheme for the travel time categories (using reversed viridis palette)
color_scheme <- rev(colorRampPalette(viridis(8))(8))

# Define the color breaks and corresponding colors for travel time intervals
color_data <- data.frame(
  from = c(0, 15, 30, 45, 60, 90, 120, 180),
  to = c(15, 30, 45, 60, 90, 120, 180, Inf),
  color = color_scheme
)

# Create a panel grid plot of all flood scenario travel time rasters
gridplot <- terra::panel(
  stacked_rasters,
  col = color_data,
  nr = 5, nc = 3, # Set grid layout: 5 rows, 3 columns
  axes = FALSE    # Hide axes for cleaner visualization
)

# Display the panel grid plot
print(gridplot)

# [OPTIONAL] Save the gridplot as a PNG file for publication or sharing
  # png(here("Figures/flood_tt_all_scenarios.png"), width = 40, height = 60, units = "cm", res = 1000)
  # print(gridplot)
  # dev.off()
```

## Travel time delta all
```{r}
# Load the baseline travel time raster for all facilities
base_tt_all <- rast(here("AccessMod/Output/Base/raster_travel_time_Base_all/raster_travel_time_Base_all.img"))

# Plot the baseline travel time raster with healthcare facilities and administrative boundaries overlay
# NOTE: Run 02_baseline_geography.qmd first to load and align landcover and facility data

w = 159.2 / 25.4  # Convert width from mm to inches for PDF output
h = 123.1 / 25.4  # Convert height from mm to inches for PDF output

pdf(here("Figures", "Pdf_final", "Figure_2a_baseline_tt.pdf"),
  width = w, height = h,
  version = "1.4",      # Required for alpha transparency
  useDingbats = FALSE)

# 1. Plot semi-transparent landcover background (outside Kampala AOI)
plot(
  landcover_complete,
  col = adjustcolor(filtered_landcover_colors, alpha.f = 0.15),
  legend = FALSE,
  axes = FALSE,
  box = FALSE
)

# 2. Overlay the baseline travel time raster (all facilities)
base_tt_all %>%
  plot(
    col = colorRampPalette(turbo(8))(8),  # Turbo color palette for strong contrast
    breaks = c(0, 15, 30, 45, 60, 90, 120, 180, Inf),  # Travel time intervals (minutes)
    nr = 5, nc = 3,  # Grid layout for legend (5 rows, 3 columns)
    axes = FALSE,    # Hide axes
    box = TRUE,      # Show plot box
    legend = FALSE,  # Hide default legend
    cex.main = 8/12, # Adjust title font size
    add = TRUE       # Add to existing plot
  ) %>%
  { 
    grid(nx = NULL, ny = NULL, col = "lightgray", lty = "dashed", lwd = 0.5)
    # Add hospitals inside AOI (solid symbols)
    points(hospitals, pch = 23, col = "black", bg = "blue", cex = 1)
    # Add health centers inside AOI (solid symbols)
    points(health_centres, pch = 21, col = "black", bg = "green", cex = 1)
    # Add hospitals outside AOI (semi-transparent symbols)
    points(hospitals_complete, pch = 23, col = adjustcolor("black", alpha.f = 0.5), bg = adjustcolor("blue", alpha.f = 0.4), cex = 1)
    # Add health centers outside AOI (semi-transparent symbols)
    points(health_centers_complete, pch = 21, col = adjustcolor("black", alpha.f = 0.5), bg = adjustcolor("green", alpha.f = 0.4), cex = 1)
  } %>%
  {
    # Add travel time legend (bottom left)
    legend(
      x = "bottomleft",
      inset = c(0.20, 0.01),
      legend = c("0-15 min", "15-30 min", "30-45 min", "45-60 min", 
             "60-90 min", "90-120 min", "120-180 min", ">180 min"),
      fill = colorRampPalette(viridis::turbo(8))(8),
      bg = "white",
      cex = 7/12,
      x.intersp = 0.1,
      y.intersp = 0.6,
      title = "Travel Time Categories (walking)",
      ncol = 2
    )
    # Add healthcare facility legend (bottom left, above travel time legend)
    legend(
      x = "bottomleft",
      inset = c(0.20, 0.17),
      legend = c("Hospitals", "Health Centers"),
      pch = c(23, 21),
      col = c("black", "black"),
      pt.bg = c("blue", "green"),
      pt.cex = c(1, 1),
      title = "Healthcare Facilities",
      bg = "white",
      cex = 7/12,
      x.intersp = 0.5,
      y.intersp = 0.6
    )
  }

# 3. Overlay administrative boundaries (parishes)
plot(Adm_boundaries, add = TRUE, col = NA, border = "black", lwd = 0.5)

# 4. Add scale bar and north arrow for map orientation
sbar(2000, xy = "bottomright", type = "bar", divs = 4, below = "km", cex = 7/12, lonlat = FALSE, scaleby = 1000)
north(type = 2, cex = 7/12)

dev.off()
```

# Create delta travel time rasters
```{r}
# Create delta travel time rasters for all facilities under each flood scenario
delta_tt_all_20_3600 <- flood_tt_all_20_3600 - base_tt_all
delta_tt_all_20_10800 <- flood_tt_all_20_10800 - base_tt_all
delta_tt_all_20_21600 <- flood_tt_all_20_21600 - base_tt_all

delta_tt_all_40_3600 <- flood_tt_all_40_3600 - base_tt_all
delta_tt_all_40_10800 <- flood_tt_all_40_10800 - base_tt_all
delta_tt_all_40_21600 <- flood_tt_all_40_21600 - base_tt_all

delta_tt_all_60_3600 <- flood_tt_all_60_3600 - base_tt_all
delta_tt_all_60_10800 <- flood_tt_all_60_10800 - base_tt_all
delta_tt_all_60_21600 <- flood_tt_all_60_21600 - base_tt_all

delta_tt_all_80_3600 <- flood_tt_all_80_3600 - base_tt_all
delta_tt_all_80_10800 <- flood_tt_all_80_10800 - base_tt_all
delta_tt_all_80_21600 <- flood_tt_all_80_21600 - base_tt_all

delta_tt_all_100_3600 <- flood_tt_all_100_3600 - base_tt_all
delta_tt_all_100_10800 <- flood_tt_all_100_10800 - base_tt_all
delta_tt_all_100_21600 <- flood_tt_all_100_21600 - base_tt_all

# Combine all delta travel time rasters into a named list for easy access and plotting
x_delta_all <- c(
  delta_tt_all_20_3600, delta_tt_all_20_10800, delta_tt_all_20_21600,
  delta_tt_all_40_3600, delta_tt_all_40_10800, delta_tt_all_40_21600,
  delta_tt_all_60_3600, delta_tt_all_60_10800, delta_tt_all_60_21600,
  delta_tt_all_80_3600, delta_tt_all_80_10800, delta_tt_all_80_21600,
  delta_tt_all_100_3600, delta_tt_all_100_10800, delta_tt_all_100_21600
)

# Assign descriptive scenario names to each delta raster
names(x_delta_all) <- c(
  "20mm / 1hr", "20mm / 3hr", "20mm / 6hr",
  "40mm / 1hr", "40mm / 3hr", "40mm / 6hr",
  "60mm / 1hr", "60mm / 3hr", "60mm / 6hr",
  "80mm / 1hr", "80mm / 3hr", "80mm / 6hr",
  "100mm / 1hr", "100mm / 3hr", "100mm / 6hr"
)

# Utility function to save delta travel time rasters to disk
save_delta_rasters <- function(raster_data, folder_name = "processed/delta_travel") {
  # Define the output directory inside Data/
  processed_dir <- here("Data", folder_name)
  
  # Create the directory if it does not exist
  if (!dir.exists(processed_dir)) {
    dir.create(processed_dir, recursive = TRUE, showWarnings = FALSE)
  }
  
  # Print the output path for verification
  message("Saving rasters to: ", processed_dir)
  
  # Save either a single SpatRaster or a list of SpatRasters
  if (inherits(raster_data, "SpatRaster")) {
    # Save multilayer SpatRaster as a single file
    file_name <- deparse(substitute(raster_data))
    writeRaster(raster_data, filename = file.path(processed_dir, paste0(file_name, ".tif")), overwrite = TRUE)
  } else if (inherits(raster_data, "list") && inherits(raster_data[[1]], "SpatRaster")) {
    # Save each SpatRaster in the list as a separate file
    for (i in seq_along(raster_data)) {
      file_path <- file.path(processed_dir, paste0(names(raster_data)[i], ".tif"))
      message("Saving: ", file_path)  # Print file being saved
      writeRaster(raster_data[[i]], filename = file_path, overwrite = TRUE)
    }
  } else {
    stop("Input must be a SpatRaster or a list of SpatRasters")
  }

  message("All rasters saved successfully.")
}

# Call the function to save all delta travel time rasters (uncomment to run)
# save_delta_rasters(x_delta_all, "processed/delta_travel")
```

## Plotting the delta travel time rasters
```{r}
# Plot all delta travel time rasters for ALL facilities using terra::panel
gridplot_delta_tt <- panel(x_delta_all,
  col = data.frame(
    from = c(0, 0.01, 15, 30, 45, 60, 90, 120, 180),  # Lower bounds for delta travel time categories
    to = c(0.01, 15, 30, 45, 60, 90, 120, 180, Inf),   # Upper bounds for delta travel time categories
    color = colorRampPalette(turbo(9))(9)              # Turbo color palette for visualization
  ),
  nr = 5, nc = 3,            # Arrange plots in a 5x3 grid
  axes = FALSE,              # Hide axes for cleaner appearance
  box = TRUE,                # Draw box around each plot
  grid = TRUE,               # Add gridlines between plots
  main = names(x_delta_all), # Use scenario names as plot titles
  loc.main = "bottom",       # Position titles at the bottom of each plot
)

# Add a custom legend for delta travel time categories
legend("topright", legend = c("0-15", "15-30", "..."),
     fill = colorRampPalette(turbo(9))(9), 
     title = "Time Difference (mins)")

# Add a main title to the entire grid plot
mtext("Delta Travel Time Accessibility Analysis for\n All Public/PFNP Facilities under different Flooding Scenarios", side = 3, line = 3, cex = 1.5)

# Plot all delta travel time rasters for ALL facilities using ggplot2's autoplot from tidyterra
gridplot_delta_tt = autoplot(x_delta_all, nrow = 5, ncol = 3) +
  scale_fill_stepsn(
    colours = colorRampPalette(turbo(9))(9),  # Turbo color palette with 9 steps
    name = "Delta Travel Time\nCategories (walking)",  # Legend title
    breaks = c(0, 15, 30, 45, 60, 90, 120, 180),       # Breaks for delta travel time categories
    labels = c("0-15 min", "15-30 min", "30-45 min", 
           "45-60 min", "60-90 min", "90-120 min", 
           "120-180 min", ">180 min"),             # Legend labels
    na.value = "transparent",                          # Make NA values transparent
    limits = c(0, 180),                                # Set limits for color scale
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",                # Place legend on the right
    legend.title = element_text(size = 12),   # Legend title font size
    legend.text = element_text(size = 10),    # Legend label font size
    legend.key.width = unit(1.5, "cm"),       # Legend key width
    legend.key.height = unit(0.5, "cm"),      # Legend key height
    legend.spacing.x = unit(1.5, 'cm'),       # Spacing between legend items
    axis.text = element_blank(),              # Remove axis text
    axis.title = element_blank(),             # Remove axis titles
    panel.grid.major = element_line(color = "gray80", size = 0.5),  # Major gridlines
    panel.grid.minor = element_line(color = "gray90", size = 0.25), # Minor gridlines
    strip.background = element_rect(fill = "white")                 # Facet label background
  ) +
  guides(
    fill = guide_coloursteps(title.position = "top", nrow = 3, byrow = FALSE)  # Arrange legend items in 3 rows
  ) +
  ggtitle("Delta Travel Time Accessibility Analysis for All Public/PFNP Facilities under different Flooding Scenarios")


# [OPTIONAL] Save the gridplot as a PNG file
  # png(here("Figures/delta_tt_all_scenarios.png"), width = 40, height = 60, units = "cm", res = 1000)
  # print(gridplot_delta_tt)
  # dev.off()

# Plot selected delta travel time scenarios only (20mm x 6 hours and 100mm x 1 hour)
x_selected <- c(delta_tt_all_20_21600, delta_tt_all_100_3600)

names(x_selected) <- c("20mm / 6hr", "100mm / 1hr")

# Plot selected scenarios using ggplot2's autoplot from tidyterra
autoplot(x_selected, nrow = 1, ncol = 2) +
  scale_fill_stepsn(
    colours = viridisLite::turbo(8),  # Turbo color palette with 8 steps
    name = "Delta Travel Time Categories (walking)",  # Legend title
    breaks = c(0, 15, 30, 45, 60, 90, 120, 180),      # Breaks for delta travel time categories
    labels = c("0-15 min", "15-30 min", "30-45 min", "45-60 min", 
           "60-90 min", "90-120 min", "120-180 min", ">180 min"),  # Legend labels
    na.value = "transparent"  # Make NA values transparent
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",                # Place legend at the bottom
    legend.title = element_text(size = 30),    # Legend title font size
    legend.text = element_text(size = 30),     # Legend label font size
    legend.key.width = unit(2, "cm"),          # Legend key width
    legend.key.height = unit(1, "cm"),         # Legend key height
    legend.spacing.x = unit(2, 'cm'),          # Spacing between legend items
    axis.text = element_blank(),               # Remove axis text
    axis.title = element_blank(),              # Remove axis titles
    panel.grid.major = element_line(color = "gray80", size = 0.5),  # Major gridlines
    panel.grid.minor = element_line(color = "gray90", size = 0.25), # Minor gridlines
    strip.background = element_rect(fill = "white"),                # Facet label background
    plot.title = element_text(size = 35),       # Plot title font size
    strip.text = element_text(size = 32)        # Scenario name font size
    ) +
    guides(
    fill = guide_legend(title.position = "top", nrow = 4, ncol = 2, byrow = FALSE),  # Arrange legend items in 2 columns
    ) +
  ggtitle("Delta Travel Time Accessibility")
```

# HOSPITAL SCENARIO
## Flood scenarios Travel Time Hospitals
```{r}
# Load all flood scenario travel time rasters (.img files) for HOSPITALS only
# Define the base directory containing flood scenario outputs for hospitals
base_dir <- here("AccessMod/Output/Flood_hosp")

# Find all scenario directories (MLC_*) within the base directory
mlc_dirs <- list.dirs(path = base_dir, full.names = TRUE, recursive = FALSE)

# Initialize a named list to store .img file paths for each scenario
scenario_img_files <- list()

# Loop through each scenario directory and find .img files in the expected subfolder
for (mlc_dir in mlc_dirs) {
  # Extract scenario name from folder name
  scenario_name <- basename(mlc_dir)
  
  # Path to the raster_travel_time_Floods_hosp_config subfolder
  raster_dir <- file.path(mlc_dir, "0-min/raster_travel_time_Floods_hosp_config")
  
  # If the raster folder exists, get all .img files
  if (dir.exists(raster_dir)) {
    img_files <- list.files(path = raster_dir, pattern = "\\.img$", full.names = TRUE, recursive = FALSE)
    
    # Add .img files to the list under the scenario name
    if (length(img_files) > 0) {
      scenario_img_files[[scenario_name]] <- img_files
    }
  }
}

# Print all scenarios and their corresponding .img files for verification
if (length(scenario_img_files) > 0) {
  cat("Flood hospital travel time scenarios and their .img files:\n")
  print(scenario_img_files)
} else {
  cat("No .img files found in any raster_travel_time_Floods_hosp_config folder.\n")
}

# Load all flood scenario rasters into a list (each scenario may have multiple rasters)
x_hosp <- lapply(scenario_img_files, function(img_files) {
  lapply(img_files, rast)
})

# Assign meaningful names to each raster in the list for clarity and consistency
x_hosp_names <- c(
  "flood_tt_hosp_100_10800", "flood_tt_hosp_100_21600", "flood_tt_hosp_100_3600",
  "flood_tt_hosp_20_10800", "flood_tt_hosp_20_21600", "flood_tt_hosp_20_3600",
  "flood_tt_hosp_40_10800", "flood_tt_hosp_40_21600", "flood_tt_hosp_40_3600",
  "flood_tt_hosp_60_10800", "flood_tt_hosp_60_21600", "flood_tt_hosp_60_3600",
  "flood_tt_hosp_80_10800", "flood_tt_hosp_80_21600", "flood_tt_hosp_80_3600"
)

# Assign scenario names to the raster list
names(x_hosp) <- x_hosp_names

# Extract the first SpatRaster from each scenario list element
x_hosp_extracted <- lapply(x_hosp, function(obj) obj[[1]])

# Export each raster to the global environment for direct access by name
list2env(x_hosp_extracted, envir = .GlobalEnv)
```

## Plotting the travel time rasters
```{r}
# Extract the first SpatRaster from each scenario in the x_hosp list
raster_list_hosp <- lapply(x_hosp, function(obj) obj[[1]])

# Combine the extracted SpatRasters into a single SpatRaster stack for visualization
stacked_rasters_hosp <- rast(raster_list_hosp)

# Define the color scheme for hospital travel time categories (reversed viridis palette)
color_scheme_hosp <- rev(colorRampPalette(viridis(8))(8))

# Specify color breaks and corresponding colors for travel time intervals (minutes)
color_data_hosp <- data.frame(
  from = c(0, 15, 30, 45, 60, 90, 120, 180),
  to = c(15, 30, 45, 60, 90, 120, 180, Inf),
  color = color_scheme_hosp
)

# Create a panel grid plot of all hospital flood scenario travel time rasters
gridplot_hosp <- terra::panel(
  stacked_rasters_hosp,
  col = color_data_hosp,
  nr = 5, nc = 3, # Set grid layout: 5 rows, 3 columns
  axes = FALSE    # Hide axes for cleaner visualization
)

# Display the panel grid plot
print(gridplot_hosp)

# [OPTIONAL] Save the gridplot as a PNG file for publication or sharing
  # png(here("Figures/flood_tt_hosp_scenarios.png"), width = 40, height = 60, units = "cm", res = 1000)
  # print(gridplot_hosp)
  # dev.off()
````

## Travel time delta Hospitals only
```{r}
# Load the baseline hospital travel time raster (for all hospitals)
base_tt_hosp <- rast(here("AccessMod/Output/Base/raster_travel_time_Base_hosp/raster_travel_time_Base_hosp.img"))

# Save the baseline hospital accessibility map as a PDF ("Figure 2b")
w = 159.2 / 25.4  # Convert width from mm to inches for PDF output
h = 123.1 / 25.4  # Convert height from mm to inches for PDF output
pdf(here("Figures", "Pdf_final", "Figure_2b_baseline_tt_hosp.pdf"),
  width = w, height = h,
  version = "1.4",      # Required for alpha transparency
  useDingbats = FALSE
)

# 1. Plot semi-transparent landcover background (outside Kampala AOI)
plot(
  landcover_complete,
  col = adjustcolor(filtered_landcover_colors, alpha.f = 0.15),
  legend = FALSE,
  axes = FALSE,
  box = FALSE
)

# 2. Overlay the baseline travel time raster (to hospitals only)
base_tt_hosp %>%
  plot(
    col = colorRampPalette(turbo(8))(8),  # Turbo color palette for strong contrast
    breaks = c(0, 15, 30, 45, 60, 90, 120, 180, Inf),  # Travel time intervals (minutes)
    nr = 5, nc = 3,  # Grid layout for legend (5 rows, 3 columns)
    axes = FALSE,    # Hide axes
    box = TRUE,      # Show plot box
    legend = FALSE,  # Hide default legend
    cex.main = 8/12, # Adjust title font size
    add = TRUE
  ) %>%
  {
    grid(nx = NULL, ny = NULL, col = "lightgray", lty = "dashed", lwd = 0.5)
    # Add hospitals outside AOI (semi-transparent symbols)
    points(hospitals_complete,
      pch = 23,
      col = adjustcolor("black", alpha.f = 0.5),
      bg  = adjustcolor("blue", alpha.f = 0.4),
      cex = 1
    )
    # Add hospitals inside AOI (solid symbols)
    points(hospitals, pch = 23, col = "black", bg = "blue", cex = 1)
  } %>%
  {
    # Add travel time legend (bottom left)
    legend(
      x = "bottomleft",
      inset = c(0.20, 0.01),
      legend = c("0-15 min", "15-30 min", "30-45 min", "45-60 min", 
                 "60-90 min", "90-120 min", "120-180 min", ">180 min"),
      fill = colorRampPalette(turbo(8))(8),
      bg = "white",
      cex = 7/12,  # Adjust legend text size
      x.intersp = 0.1,
      y.intersp = 0.6,
      title = "Travel Time Categories (walking)",
      ncol = 2
    )
    # Add healthcare facility legend (bottom left, above travel time legend)
    legend(
      x = "bottomleft",
      inset = c(0.20, 0.17),
      legend = c("Hospitals"),
      pch = c(23),
      col = "black",
      pt.bg = c("blue"),
      pt.cex = 1,
      title = "Healthcare Facilities",
      bg = "white",
      cex = 7/12,  # Adjust legend text size
      x.intersp = 0.5,
      y.intersp = 0.6
    )
  }

# 3. Overlay administrative boundaries (parishes)
plot(Adm_boundaries, add = TRUE, col = NA, border = "black", lwd = 0.5)

# 4. Add scale bar and north arrow for map orientation
sbar(2000, xy = "bottomright", type = "bar", divs = 4, below = "km", cex = 7/12, lonlat = FALSE, scaleby = 1000)
north(type = 2, cex = 7/12)

dev.off()
```

# Create delta travel time rasters
```{r}
# Calculate delta travel time rasters for each hospital flood scenario (difference from baseline)
delta_tt_hosp_20_3600 <- flood_tt_hosp_20_3600 - base_tt_hosp
delta_tt_hosp_20_10800 <- flood_tt_hosp_20_10800 - base_tt_hosp
delta_tt_hosp_20_21600 <- flood_tt_hosp_20_21600 - base_tt_hosp

delta_tt_hosp_40_3600 <- flood_tt_hosp_40_3600 - base_tt_hosp
delta_tt_hosp_40_10800 <- flood_tt_hosp_40_10800 - base_tt_hosp
delta_tt_hosp_40_21600 <- flood_tt_hosp_40_21600 - base_tt_hosp

delta_tt_hosp_60_3600 <- flood_tt_hosp_60_3600 - base_tt_hosp
delta_tt_hosp_60_10800 <- flood_tt_hosp_60_10800 - base_tt_hosp
delta_tt_hosp_60_21600 <- flood_tt_hosp_60_21600 - base_tt_hosp

delta_tt_hosp_80_3600 <- flood_tt_hosp_80_3600 - base_tt_hosp
delta_tt_hosp_80_10800 <- flood_tt_hosp_80_10800 - base_tt_hosp
delta_tt_hosp_80_21600 <- flood_tt_hosp_80_21600 - base_tt_hosp

delta_tt_hosp_100_3600 <- flood_tt_hosp_100_3600 - base_tt_hosp
delta_tt_hosp_100_10800 <- flood_tt_hosp_100_10800 - base_tt_hosp
delta_tt_hosp_100_21600 <- flood_tt_hosp_100_21600 - base_tt_hosp

# Combine all hospital delta travel time rasters into a named list for plotting and export
x_delta_hosp <- c(
  delta_tt_hosp_20_3600, delta_tt_hosp_20_10800, delta_tt_hosp_20_21600,
  delta_tt_hosp_40_3600, delta_tt_hosp_40_10800, delta_tt_hosp_40_21600,
  delta_tt_hosp_60_3600, delta_tt_hosp_60_10800, delta_tt_hosp_60_21600,
  delta_tt_hosp_80_3600, delta_tt_hosp_80_10800, delta_tt_hosp_80_21600,
  delta_tt_hosp_100_3600, delta_tt_hosp_100_10800, delta_tt_hosp_100_21600
)

# Assign descriptive scenario names to each hospital delta raster
names(x_delta_hosp) <- c(
  "20mm / 1hr", "20mm / 3hr", "20mm / 6hr",
  "40mm / 1hr", "40mm / 3hr", "40mm / 6hr",
  "60mm / 1hr", "60mm / 3hr", "60mm / 6hr",
  "80mm / 1hr", "80mm / 3hr", "80mm / 6hr",
  "100mm / 1hr", "100mm / 3hr", "100mm / 6hr"
)

# Save all hospital delta travel time rasters to disk for further analysis or sharing
save_delta_rasters(x_delta_hosp, "processed/delta_travel")
```

## Plotting the delta travel time rasters
```{r}
# Plot all hospital delta travel time rasters using ggplot2's autoplot() from tidyterra
gridplot_delta_tt_hosp = autoplot(x_delta_hosp, nrow = 5, ncol = 3) +
  scale_fill_stepsn(
    colours = colorRampPalette(turbo(9))(9),  # Turbo color palette with 9 steps for clear visualization
    name = "Delta Travel Time\nCategories (walking)",  # Legend title for delta travel time
    breaks = c(0, 15, 30, 45, 60, 90, 120, 180),  # Delta travel time category breaks (minutes)
    labels = c("0-15 min", "15-30 min", "30-45 min", 
           "45-60 min", "60-90 min", "90-120 min", 
           "120-180 min", ">180 min"),  # Labels for legend categories
    na.value = "transparent",  # Make NA values transparent for cleaner plots
    limits = c(0, 180),  # Set color scale limits to match breaks
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",  # Place legend on the right side
    legend.title = element_text(size = 12),  # Legend title font size
    legend.text = element_text(size = 10),  # Legend label font size
    legend.key.width = unit(1.5, "cm"),  # Legend key width
    legend.key.height = unit(0.5, "cm"),  # Legend key height
    legend.spacing.x = unit(1.5, 'cm'),  # Spacing between legend items
    axis.text = element_blank(),  # Remove axis text for clarity
    axis.title = element_blank(),  # Remove axis titles for clarity
    panel.grid.major = element_line(color = "gray80", size = 0.5),  # Major gridlines for panel background
    panel.grid.minor = element_line(color = "gray90", size = 0.25),  # Minor gridlines for panel background
    strip.background = element_rect(fill = "white")  # White background for facet labels
  ) +
  guides(
    fill = guide_coloursteps(title.position = "top", nrow = 3, byrow = FALSE)  # Arrange legend items in 3 rows
  ) +
  ggtitle("Delta Travel Time Accessibility Analysis for Public/PNFP Hospitals under different Flooding Scenarios")

# [OPTIONAL] Save the gridplot as a PNG file for publication or sharing
  # png(here("Figures/delta_tt_hosp_scenarios.png"), width = 40, height = 60, units = "cm", res = 300)
  # print(gridplot_delta_tt_hosp)
  # dev.off()
```

# Figure 2 — Combined two-panel figure
```{r}
# Two-panel PDF: panel a (all facilities) on the left,
#                panel b (hospitals only) on the right.
# Shared legend in panel a only. No legend in panel b.
# Output: Figure_2_combined.pdf (318.4 x 123.1 mm, double-column)

w_panel <- 159.2 / 25.4   # single-panel width  in inches
h_panel <- 123.1 / 25.4   # single-panel height in inches

pdf(here("Figures", "Pdf_final", "Figure_2_combined.pdf"),
    width       = 2 * w_panel,
    height      = h_panel,
    version     = "1.4",
    useDingbats = FALSE)

# Two equal columns, zero margins (cex controlled per-call throughout)
par(mfrow = c(1, 2),
    mai  = c(0, 0, 0, 0),
    oma  = c(0, 0, 0, 0),
    xaxs = "i", yaxs = "i")

# ================================================================
# PANEL a  —  Baseline travel time to ALL facilities
# ================================================================

# 1. Semi-transparent landcover background (outside AOI)
plot(landcover_complete,
     col    = adjustcolor(filtered_landcover_colors, alpha.f = 0.15),
     legend = FALSE, axes = FALSE, box = FALSE)

# 2. Travel time raster (all facilities)
plot(base_tt_all,
     col    = colorRampPalette(turbo(8))(8),
     breaks = c(0, 15, 30, 45, 60, 90, 120, 180, Inf),
     axes   = FALSE, box = TRUE, legend = FALSE, add = TRUE)

# 3. Light reference grid
grid(nx = NULL, ny = NULL, col = "lightgray", lty = "dashed", lwd = 0.5)

# 4. Healthcare facilities
points(hospitals_complete,
       pch = 23, col = adjustcolor("black", alpha.f = 0.5),
       bg  = adjustcolor("blue",  alpha.f = 0.4), cex = 1)
points(health_centers_complete,
       pch = 21, col = adjustcolor("black", alpha.f = 0.5),
       bg  = adjustcolor("green", alpha.f = 0.4), cex = 1)
points(hospitals,      pch = 23, col = "black", bg = "blue",  cex = 1)
points(health_centres, pch = 21, col = "black", bg = "green", cex = 1)

# 5. Parish / administrative boundaries
plot(Adm_boundaries, add = TRUE, col = NA, border = "black", lwd = 0.5)

# Panel label — inside frame, upper-left, bold
usr_a <- par("usr")
text(usr_a[1] + 0.03 * (usr_a[2] - usr_a[1]),
     usr_a[4] - 0.04 * (usr_a[4] - usr_a[3]),
     labels = "a", adj = c(0, 1), cex = 9/12, font = 2, col = "black")

# Custom scale bar — Figure 1 design (inch-anchored)
pin_a  <- par("pin")
uin_xa <- (usr_a[2] - usr_a[1]) / pin_a[1]
uin_ya <- (usr_a[4] - usr_a[3]) / pin_a[2]
sb_h_a  <- 0.055 * uin_ya
sb_x2_a <- usr_a[2] - 0.08 * uin_xa
sb_x1_a <- sb_x2_a - 2000
sb_xm_a <- sb_x1_a + 1000
sb_y0_a <- usr_a[3] + 0.70 * uin_ya
sb_y1_a <- sb_y0_a + sb_h_a
rect(sb_x1_a, sb_y0_a, sb_xm_a, sb_y1_a, col = "black",  border = NA)
rect(sb_xm_a, sb_y0_a, sb_x2_a, sb_y1_a, col = "grey88", border = NA)
rect(sb_x1_a, sb_y0_a, sb_x2_a, sb_y1_a, col = NA, border = "black", lwd = 0.7)
segments(c(sb_x1_a, sb_xm_a, sb_x2_a), rep(sb_y1_a, 3),
         c(sb_x1_a, sb_xm_a, sb_x2_a), rep(sb_y1_a + sb_h_a, 3),
         col = "black", lwd = 0.7)
lbl_ya <- sb_y1_a + sb_h_a + 0.018 * uin_ya
text(c(sb_x1_a, sb_xm_a, sb_x2_a), rep(lbl_ya, 3),
     labels = c("0", "1", "2"), adj = c(0.5, 0), cex = 6/12, col = "black")
text(sb_x2_a + 0.035 * uin_xa, lbl_ya,
     labels = "km", adj = c(0, 0), cex = 5.5/12, col = "black")

north(type = 2, cex = 7/12)

# 8. SHARED LEGEND — Travel time categories
legend(
  x = "bottomleft", inset = c(0.20, 0.01),
  legend = c("0-15 min", "15-30 min", "30-45 min", "45-60 min",
             "60-90 min", "90-120 min", "120-180 min", ">180 min"),
  fill    = colorRampPalette(turbo(8))(8),
  bty     = "o", box.lwd = 0.5, box.col = "grey70",
  bg      = adjustcolor("white", alpha.f = 0.85),
  cex     = 7/12, x.intersp = 0.1, y.intersp = 0.6,
  title   = "Travel Time Categories (walking)", ncol = 2
)

# 9. SHARED LEGEND — Map elements (facilities + parish boundaries)
legend(
  x = "bottomleft", inset = c(0.20, 0.17),
  legend = c("Hospitals", "Health centres", "Parish boundaries"),
  pch    = c(23,      21,      NA),
  lty    = c(NA,      NA,       1),
  lwd    = c(NA,      NA,     0.5),
  col    = c("black", "black", "black"),
  pt.bg  = c("blue",  "green",  NA),
  pt.cex = 1,
  title  = "Healthcare Facilities",
  bty    = "o", box.lwd = 0.5, box.col = "grey70",
  bg     = adjustcolor("white", alpha.f = 0.85),
  cex    = 7/12, x.intersp = 0.5, y.intersp = 0.6
)

# ================================================================
# PANEL b  —  Baseline travel time to HOSPITALS ONLY
# No legend — shared legend is in panel a
# ================================================================

# 1. Semi-transparent landcover background (outside AOI)
plot(landcover_complete,
     col    = adjustcolor(filtered_landcover_colors, alpha.f = 0.15),
     legend = FALSE, axes = FALSE, box = FALSE)

# 2. Travel time raster (hospitals only)
plot(base_tt_hosp,
     col    = colorRampPalette(turbo(8))(8),
     breaks = c(0, 15, 30, 45, 60, 90, 120, 180, Inf),
     axes   = FALSE, box = TRUE, legend = FALSE, add = TRUE)

# 3. Light reference grid
grid(nx = NULL, ny = NULL, col = "lightgray", lty = "dashed", lwd = 0.5)

# 4. Hospitals only
points(hospitals_complete,
       pch = 23, col = adjustcolor("black", alpha.f = 0.5),
       bg  = adjustcolor("blue",  alpha.f = 0.4), cex = 1)
points(hospitals, pch = 23, col = "black", bg = "blue", cex = 1)

# 5. Parish / administrative boundaries
plot(Adm_boundaries, add = TRUE, col = NA, border = "black", lwd = 0.5)

# Panel label — inside frame, upper-left, bold
usr_b <- par("usr")
text(usr_b[1] + 0.03 * (usr_b[2] - usr_b[1]),
     usr_b[4] - 0.04 * (usr_b[4] - usr_b[3]),
     labels = "b", adj = c(0, 1), cex = 9/12, font = 2, col = "black")

# Custom scale bar — Figure 1 design (inch-anchored)
pin_b  <- par("pin")
uin_xb <- (usr_b[2] - usr_b[1]) / pin_b[1]
uin_yb <- (usr_b[4] - usr_b[3]) / pin_b[2]
sb_h_b  <- 0.055 * uin_yb
sb_x2_b <- usr_b[2] - 0.08 * uin_xb
sb_x1_b <- sb_x2_b - 2000
sb_xm_b <- sb_x1_b + 1000
sb_y0_b <- usr_b[3] + 0.70 * uin_yb
sb_y1_b <- sb_y0_b + sb_h_b
rect(sb_x1_b, sb_y0_b, sb_xm_b, sb_y1_b, col = "black",  border = NA)
rect(sb_xm_b, sb_y0_b, sb_x2_b, sb_y1_b, col = "grey88", border = NA)
rect(sb_x1_b, sb_y0_b, sb_x2_b, sb_y1_b, col = NA, border = "black", lwd = 0.7)
segments(c(sb_x1_b, sb_xm_b, sb_x2_b), rep(sb_y1_b, 3),
         c(sb_x1_b, sb_xm_b, sb_x2_b), rep(sb_y1_b + sb_h_b, 3),
         col = "black", lwd = 0.7)
lbl_yb <- sb_y1_b + sb_h_b + 0.018 * uin_yb
text(c(sb_x1_b, sb_xm_b, sb_x2_b), rep(lbl_yb, 3),
     labels = c("0", "1", "2"), adj = c(0.5, 0), cex = 6/12, col = "black")
text(sb_x2_b + 0.035 * uin_xb, lbl_yb,
     labels = "km", adj = c(0, 0), cex = 5.5/12, col = "black")

north(type = 2, cex = 7/12)

# ----------------------------------------------------------------
dev.off()
```
